### 2017-01-14 paper/ms.tex
manuscript files for Entropy journal

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================

% LaTeX Class File and Rendering Mode (choose one)
% You will need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================

\documentclass[journal,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, administrativesciences, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences, arts, atmosphere, atoms, axioms, batteries, behavioralsciences, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsciences, buildings, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, cosmetics, crystals, data, dentistryjournal, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, inorganics, insects, ijerph, ijfs, ijms, ijns, ijgi, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, joi, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, lubricants, machines, marinedrugs, materials, mathematics, medicalsciences, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, sinusitis, socialsciences, societies, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, universe, vaccines, veterinarysciences, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but could be replaced by using one of the class options:
% article, review, communication, commentary, bookreview, correction, addendum, editorial, changes, supfile, casereport, comment, conceptpaper, conferencereport, meetingreport, discussion, essay, letter, newbookreceived, opinion, projectreport, reply, retraction, shortnote, technicalnote, creative, datadescriptor (for journal Data), briefreport, hypothesis, interestingimage
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
% Please insert a blank line is before and after all equation and eqnarray environments to ensure proper line numbering when option submit is chosen
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use the optioin "dvipdfm", with LaTeX and dvi2pdf only.

%=================================================================
\setcounter{page}{1}
\lastpage{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2016}
%\externaleditor{Academic Editor: xx}
\history{Received: xx / Accepted: xx / Published: xx}

\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{microtype}

\DeclareUnicodeCharacter{00A0}{ }
\newcommand{\xx}{\boldsymbol{\theta}}
\newcommand{\dx}{d\boldsymbol{\theta}}
\newcommand{\data}{\boldsymbol{D}}
\newcommand{\II}{\boldsymbol{I}}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================

% Add packages and commands to include here
% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are loaded by the MDPI class.
%\usepackage{graphicx}
%\usepackage{subfigure,psfig}

%=================================================================
%% Please use the following mathematics environments:
% \theoremstyle{mdpi}
% \newcounter{thm}
% \setcounter{thm}{0}
% \newcounter{ex}
% \setcounter{ex}{0}
% \newcounter{re}
% \setcounter{re}{0}
%
% \newtheorem{Theorem}[thm]{Theorem}
% \newtheorem{Lemma}[thm]{Lemma}
% \newtheorem{Corollary}[thm]{Corollary}
% \newtheorem{Proposition}[thm]{Proposition}
%
% \theoremstyle{mdpidefinition}
% \newtheorem{Characterization}[thm]{Characterization}
% \newtheorem{Property}[thm]{Property}
% \newtheorem{Problem}[thm]{Problem}
% \newtheorem{Example}[ex]{Example}
% \newtheorem{ExamplesandDefinitions}[ex]{Examples and Definitions}
% \newtheorem{Remark}[re]{Remark}
% \newtheorem{Definition}[thm]{Definition}
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================

% Full title of the paper (Capitalized)
\Title{Calculating Entropies With Nested Sampling}

% Authors (Add full first names)
\Author{Brendon J. Brewer$^{1,}$*}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{
$^{1}$ Department of Statistics, The University of Auckland, Private Bag 92019,
Auckland 1142, New Zealand}

%\contributed{$^\dagger$ These authors contributed equally to this work.}

% Contact information of the corresponding author (Add [2] after \corres if there are more than one corresponding author.)
\corres{{\tt bj.brewer@auckland.ac.nz}}

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{}

% Keywords: add 3 to 10 keywords
\keyword{nested sampling; information theory; monte carlo}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{}
%\MSC{}
%\JEL{}

% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\dagger$ in the end of the title of this article.
%\conference{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For journal Data:

%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

\acknowledgments{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\authorcontributions{Author Contributions}

%%%%%%%%%%%%%%%%%%%%%%%%%%

\conflictofinterests{Conflicts of Interest}
The authors declare no conflicts of interest.

%=================================================================
% References: Variant A
%=================================================================
% Back Matter (References and Notes)
%----------------------------------------------------------
% Style and layout of the references
\bibliographystyle{mdpi}
\makeatletter
\renewcommand\@biblabel[1]{#1. }
\makeatother

%=================================================================
% References:  Variant B
%=================================================================
% Use the following option to include external BibTeX files:
\bibliography{lite}
\bibliographystyle{mdpi}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\abbreviations{Abbreviations/Nomenclature}
%
%Main text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2017-01-14 cpp/RNG.cpp
Copied DNest4's RNG class

@@ -1,0 +1,53 @@
+#include "RNG.h"
+#include <cmath>
+
+namespace InfoNest
+{
+
+RNG::RNG()
+:uniform(0., 1.)
+,normal(0., 1.)
+{
+
+}
+
+RNG::RNG(unsigned int seed)
+:uniform(0., 1.)
+,normal(0., 1.)
+{
+    set_seed(seed);
+}
+
+void RNG::set_seed(unsigned int seed)
+{
+    twister.seed(seed);
+}
+
+double RNG::rand()
+{
+    return uniform(twister);
+}
+
+double RNG::randn()
+{
+    return normal(twister);
+}
+
+double RNG::randt2()
+{
+    // t-distribution with 2 degrees of freedom (less ridiculous than Cauchy)
+    return this->randn()/sqrt(-log(this->rand()));
+}
+
+double RNG::randh()
+{
+    return pow(10.0, 1.5 - 3*std::abs(this->randt2()))*this->randn();
+}
+
+int RNG::rand_int(int N)
+{
+    return static_cast<int>(floor(N*this->rand()));
+}
+
+} // namespace InfoNest
+

### 2017-01-14 cpp/RNG.h

@@ -1,0 +1,53 @@
+#ifndef InfoNest_RNG
+#define InfoNest_RNG
+
+#include <random>
+
+namespace InfoNest
+{
+
+/*
+* An object of this class represents a random number generator 
+*/
+class RNG
+{
+    private:
+        // 64-bit Mersenne Twister
+        std::mt19937_64 twister;
+
+        // For uniform distribution
+        std::uniform_real_distribution<double> uniform;
+
+        // For normal distribution
+        std::normal_distribution<double> normal;
+
+    public:
+        // Constructor
+        RNG();
+
+        // Constructor that also sets seed
+        RNG(unsigned int seed);
+
+        // Set the seed (obviously)
+        void set_seed(unsigned int seed);
+
+        // Uniform(0, 1)
+        double rand();
+
+        // Normal(0, 1)
+        double randn();
+
+        // t(location=0, scale=1, nu=2)
+        double randt2();
+
+        // My favourite heavy-tailed distribution
+        double randh();
+
+        // Integer from {0, 1, 2, ..., N-1}
+        int rand_int(int N);
+}; // class RNG
+
+} // namespace InfoNest
+
+#endif
+

### 2017-01-14 cpp/Rep.h
Starting Rep template class

@@ -1,0 +1,56 @@
+#ifndef InfoNest_Rep
+#define InfoNest_Rep
+
+#include <stdlib.h>
+
+namespace InfoNest
+{
+
+/*
+* An object of this class represents a single `rep`
+* of the algorithm. The template parameter `particle`
+* determines what problem you're working on.
+*/
+template<class Particle>
+class Rep
+{
+    private:
+        // Nested Sampling parameters
+        const size_t num_particles;
+        const size_t mcmc_steps;
+        const double depth;
+
+    public:        
+        // Constructor. You need to pass in the NS parameters
+        Rep(size_t num_particles, size_t mcmc_steps, double depth);
+
+
+
+
+};
+
+
+/******************************/
+/* IMPLEMENTATION STARTS HERE */
+/******************************/
+
+template<class Particle>
+Rep<Particle>::Rep(size_t num_particles, size_t mcmc_steps, double depth)
+:num_particles(num_particles)
+,mcmc_steps(mcmc_steps)
+,depth(depth)
+{
+
+}
+
+
+
+
+
+
+
+
+} // namespace InfoNest
+
+#endif
+

### 2017-01-14 cpp/Rep.h
Implemented initialise and compute_distances

@@ -1,7 +1,10 @@
 #ifndef InfoNest_Rep
 #define InfoNest_Rep
 
+// Includes
 #include <stdlib.h>
+#include <vector>
+#include "RNG.h"
 
 namespace InfoNest
 {
@@ -20,12 +23,27 @@ class Rep
         const size_t mcmc_steps;
         const double depth;
 
+        // The reference particle
+        Particle reference_particle;
+
+        // The other particles
+        std::vector<Particle> particles;
+
+        // Distances
+        std::vector<double> distances;
+
+        // Iteration counter
+        unsigned int iteration;
+
+        // Member function to compute all the distances
+        void compute_distances();
+
     public:        
         // Constructor. You need to pass in the NS parameters
         Rep(size_t num_particles, size_t mcmc_steps, double depth);
 
-
-
+        // Initialise all the particles
+        void initialise(RNG& rng);
 
 };
 
@@ -39,16 +57,38 @@ Rep<Particle>::Rep(size_t num_particles, size_t mcmc_steps, double depth)
 :num_particles(num_particles)
 ,mcmc_steps(mcmc_steps)
 ,depth(depth)
+,particles(num_particles)
+,distances(num_particles)
+,iteration(0)
 {
 
 }
 
 
+template<class Particle>
+void Rep<Particle>::initialise(RNG& rng)
+{
+    // Generate the reference particle from the distribution
+    reference_particle.generate(rng);
+
+    // Generate the other particles
+    for(auto& p: particles)
+        p.generate(rng);
+
+    // Compute all the distances
+    compute_distances();
+
+    // Set iteration number
+    iteration = 1;
+}
 
 
-
-
-
+template<class Particle>
+void Rep<Particle>::compute_distances()
+{
+    for(size_t i=0; i<num_particles; ++i)
+        distances[i] = Particle::distance(reference_particle, particles[i]);
+}
 
 } // namespace InfoNest
 

### 2017-01-14 cpp/main.cpp
Implemented but needs testing

@@ -5,13 +5,31 @@
 #include "RNG.h"
 #include "Examples/Normal.h"
 
+void copyright();
+
 int main()
 {
     // Create random number generator, seeded with time
     InfoNest::RNG rng(time(0));
 
-    // Create a Rep.
-    InfoNest::Rep<InfoNest::Normal> rep(10, 1000, 25.0, rng);
+    // Total number of reps to do
+    int num_reps = 10000;
+
+    // Clear output file
+    std::fstream fout("output.txt", std::ios::out);
+    fout.close();
+
+    for(int i=0; i<num_reps; ++i)
+    {
+        // Create a Rep.
+        InfoNest::Rep<InfoNest::Normal> rep(10, 1000, 25.0, rng);
+
+        // Initialise and execute it.
+        rep.initialise();
+        rep.execute();
+
+        std::cout<<"Completed "<<(i+1)<<" reps."<<std::endl;
+    }
 
     return 0;
 }

### 2017-01-14 cpp/main.cpp
All working

@@ -5,10 +5,13 @@
 #include "RNG.h"
 #include "Examples/Normal.h"
 
-void copyright();
-
 int main()
 {
+    // Run parameters
+    size_t num_particles = 10;
+    size_t mcmc_steps = 1000;
+    double depth = 25.0;
+
     // Create random number generator, seeded with time
     InfoNest::RNG rng(time(0));
 
@@ -19,16 +22,25 @@ int main()
     std::fstream fout("output.txt", std::ios::out);
     fout.close();
 
+    // Write run data to a file
+    fout.open("run_data.txt", std::ios::out);
+    fout << num_particles << "\t# num_particles." << std::endl;
+    fout.close();
+
     for(int i=0; i<num_reps; ++i)
     {
         // Create a Rep.
-        InfoNest::Rep<InfoNest::Normal> rep(10, 1000, 25.0, rng);
+        InfoNest::Rep<InfoNest::Normal> rep(num_particles,
+                                            mcmc_steps,
+                                            depth,
+                                            rng);
 
         // Initialise and execute it.
         rep.initialise();
         rep.execute();
 
-        std::cout<<"Completed "<<(i+1)<<" reps."<<std::endl;
+        // Print a message.
+        std::cout << "Completed " << (i+1) << " reps." << std::endl;
     }
 
     return 0;

### 2017-01-14 cpp/postprocess.py

@@ -1,0 +1,45 @@
+__all__ = ["postprocess"]
+
+# Imports
+import numpy as np
+import matplotlib.pyplot as plt
+
+def postprocess(num_particles=1, tol=1E-3):
+    """
+    Load the output file and estimate the entropy.
+    """
+
+    # Load output file
+    output = np.loadtxt("output.txt")
+    num_particles = np.loadtxt("run_data.txt")
+
+    # Find the starts of the runs
+    start_indices = np.nonzero(output[:,0] == 1)[0]
+
+    reps = len(start_indices)
+    counts = np.empty(reps)
+    for i in range(0, reps):
+        start_index = start_indices[i]
+        if i < reps - 1:
+            run = output[start_index : start_indices[i+1], :]
+        else:
+            run = output[start_index : , :]
+        print("Processed run {k}/{N}.".format(k=i+1, N=reps))
+        counts[i] = (run[:,1] > tol).sum()
+
+    print("mean(counts) = {mc}.".format(mc=counts.mean()))
+    print("std (counts) = {sd}.".format(sd=counts.std()))
+
+    # Estimate of the differential entropy
+    # log(probability) ~= log(2 * tol * density)
+    H = counts.mean() / num_particles + np.log(2 * tol)
+    stderr = (counts.std() / num_particles) / np.sqrt(reps)
+
+    print("H = {H} +- {stderr} nats.".format(H=H, stderr=stderr))
+
+#    plt.hist(mean_depth, 100, color=[0.2, 0.2, 0.2])
+#    plt.show()
+
+if __name__ == "__main__":
+    postprocess()
+

### 2017-01-14 paper/ms.tex
Wrote down the Shannon entropy and the product rule

@@
 \lastpage{x}
-\doinum{10.3390/------}
-\pubvolume{xx}
-\pubyear{2016}
+\doinum{}       %\doinum{10.3390/------}
+\pubvolume{}    % \pubvolume{xx}
+\pubyear{2017}
 %\externaleditor{Academic Editor: xx}
-\history{Received: xx / Accepted: xx / Published: xx}
+\history{In preparation}%\history{Received: xx / Accepted: xx / Published: xx}
 
@@
 \DeclareUnicodeCharacter{00A0}{ }
-\newcommand{\xx}{\boldsymbol{\theta}}
-\newcommand{\dx}{d\boldsymbol{\theta}}
-\newcommand{\data}{\boldsymbol{D}}
-\newcommand{\II}{\boldsymbol{I}}
+\newcommand{\x}{\boldsymbol{x}}
+\renewcommand{\d}{\boldsymbol{d}}
 
@@
 
+If an unknown quantity $\x$ has a discrete probability distribution $p(\x)$,
+the entropy is
+\begin{align}
+H_{\x} &= -\sum_{\x} p(\x) \log p(\x)
+\end{align}
+where the sum is over all of the possible values of $\x$ under consideration.
+This probability distribution is implicitly conditional on some prior
+information, which is have omitted from the notation.
+
+\section{Entropies in Bayesian inference}
+
+In a standard situation where Bayesian inference would be used,
+the unknown quantities are the parameters $\x$ and the data $\d$
+(which subsequently becomes known). The prior information leads the
+user to specify a prior distribution $p(\x)$ for the unknown parameters,
+along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
+of how the data is related to the parameters. By the product rule, this yields
+a joint prior
+\begin{align}
+p(\x, \d) &= p(\x)p(\d | \x)
+\end{align}
+
 \acknowledgments{Acknowledgements}

### 2017-01-14 paper/ms.tex
Bibtex file
Conditional entropy

@@
 
+\subsection{The relevance of data}
+
+The degree to which the question ``what is the value of $\x$?'' is
+answered by the question ``what is the value of $\d$?'' is given by the
+conditional entropy
+\begin{align}
+H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
+\end{align}
+which is the expected value of the entropy of the posterior, averaged over
+all possible datasets which might be observed.
+
+\section{Nested Sampling}
+Nested Sampling \citep[NS][]{skilling2006nested}
+
 \acknowledgments{Acknowledgements}
@@
 % Style and layout of the references
-\bibliographystyle{mdpi}
+%\bibliographystyle{mdpi}
 \makeatletter
@@
 % Use the following option to include external BibTeX files:
-\bibliography{lite}
 \bibliographystyle{mdpi}
+\bibliography{references}
 

### 2017-01-14 cpp/main.cpp
Faster running and loading

@@ -8,7 +8,7 @@
 int main()
 {
     // Run parameters
-    size_t num_particles = 10;
+    size_t num_particles = 1;
     size_t mcmc_steps = 1000;
     double depth = 25.0;
 

### 2017-01-14 cpp/postprocess.py

@@ -1,4 +1,4 @@
-__all__ = ["postprocess"]
+__all__ = ["postprocess", "my_loadtxt"]
 
 # Imports
 import numpy as np
@@ -10,7 +10,7 @@ def postprocess(num_particles=1, tol=1E-3):
     """
 
     # Load output file
-    output = np.loadtxt("output.txt")
+    output = my_loadtxt("output.txt")
     num_particles = np.loadtxt("run_data.txt")
 
     # Find the starts of the runs
@@ -39,6 +39,48 @@ def postprocess(num_particles=1, tol=1E-3):
 
 #    plt.hist(mean_depth, 100, color=[0.2, 0.2, 0.2])
 #    plt.show()
+
+
+def my_loadtxt(filename, single_precision=False):
+    """
+    Load rectangular arrays from a file, faster than numpy.loadtxt does it.
+    """
+    # Open the file
+    f = open(filename, "r")
+
+    # Storage
+    results = []
+
+    # Rows and columns
+    nrow = 0
+    ncol = None
+
+    while(True):
+        # Read the line and split by whitespace
+        line = f.readline()
+        cells = line.split()
+
+        # Quit when you see a different number of columns
+        if ncol is not None and len(cells) != ncol:
+            break
+
+        # Non-comment lines
+        if cells[0][0] != "#":
+            # If it's the first one, get the number of columns
+            if ncol is None:
+                ncol = len(cells)
+
+            # Otherwise, include in results
+            if single_precision:
+                results.append(np.array([float(cell) for cell in cells],\
+                                                          dtype="float32"))
+            else:
+                results.append(np.array([float(cell) for cell in cells]))
+            nrow += 1
+
+    results = np.vstack(results)
+    return results
+
 
 if __name__ == "__main__":
     postprocess()

### 2017-01-15 paper/ms.tex
Walter reference

@@
 
+\subsection{The sequence of $X$ values}
+When $N=1$, the sequence of $X$ values is a Poisson process.
+
+\citet{Walter2015} showed how this property can be used to construct a
+version of NS that produces unbiased estimates of the marginal likelihood.
+However, it can be used to construct an unbiased estimator of
+a log-probability, which is more relevant to information theoretic
+quantities.
+
+\subsection{}
+
 \acknowledgments{Acknowledgements}

### 2017-01-15 paper/ms.tex
Some sections, to remind myself of what to discuss

--- 2017-01-14 paper/ms.tex
@@
 \DeclareUnicodeCharacter{00A0}{ }
+
+\definecolor{orange}{rgb}{1, 0.5, 0}
+
+\renewcommand{\d}{\boldsymbol{d}}
+\newcommand{\todo}{\color{orange} \bf}
 \newcommand{\x}{\boldsymbol{x}}
-\renewcommand{\d}{\boldsymbol{d}}
 
@@
 
+The
+
+\subsection{Mutual information}
+
 \section{Nested Sampling}
@@
 
+\subsection{The sequence of $X$ values}
+When $N=1$, the sequence of $X$ values is a Poisson process.
+
+\citet{Walter2015} showed how this view of the NS sequence of points
+can be used to construct a
+version of NS that produces unbiased estimates of the marginal likelihood.
+However, it can also be used to construct an unbiased estimator of
+a log-probability, which is more relevant to information theoretic
+quantities.
+
+\section{The algorithm}
+
+\section{Example 1: A trivial bivariate gaussian}
+
+\section{Example 2: Measuring the period of an oscillating signal}
+
+The multimodality of the posterior here raises an interesting issue. Is
+the question we really want answered ``what is the value of $T$
+{\bf precisely}?'', to which the mutual information relates?
+Most practicing scientists would not feel particularly informed to learn
+that the vast majority of possibilities had been ruled out, if the
+posterior still consisted of several widely separated modes!
+
+Perhaps, in some applications, a more appropriate question is
+``what is the value of $T$ to within $\pm$ 10\%''
+{\todo or something like that.}
+
+\section{Example 3: TBD}
+
 \acknowledgments{Acknowledgements}
+It is a pleasure to thank the following people for interesting and helpful
+conversations about this topic: Ruth Angus (Flatiron Institute),
+Ewan Cameron (Oxford), David Hogg (NYU), Kevin Knuth (SUNY Albany),
+Iain Murray (Edinburgh).
 

### 2017-01-15 paper/ms.tex
Why it's hard
Distinguish the two kinds of conditional entropy

@@
 
-The
+This is difficult to compute because the expression
+for the posterior distribution
+\begin{align}
+p(\x | \d) &= \frac{p(\x)p(\d | \x)}{p(\d)}
+\end{align}
+contains the marginal likelihood $p(\d)$.
+
+It is important to distinguish between $H_{\x | \d}$ and the
+entropy of $\x$ given a particular value of $\d$, which might be written
+$H_{\x | \d=\d_0}$.
+The former measures the degree to which one question answers
+another, so is a function of two questions. The latter measures the
+degree to which a {\em statement} answered a question, and is a function
+of a question and a statement.
 
 \subsection{Mutual information}
+
+The mutual information is
+\begin{align}
+I_{\x, \d} &= \sum_{\x} \sum_{\d}
+\end{align}
 

### 2017-01-15 paper/ms.tex
Mutual information formula
More on mutual information

@@
 of how the data is related to the parameters. By the product rule, this yields
-a joint prior
+a {\em joint prior}
 \begin{align}
@@
 \end{align}
+
+The entropy of this joint prior describes the degree to which the question
+``what is the value of the pair $(\x, \d)$?'' remains unanswered.
 
@@
 \subsection{Mutual information}
+The mutual information is another way of describing the relevance of the
+data to the parameters.
 
-The mutual information is
+Its definition is
 \begin{align}
-I_{\x, \d} &= \sum_{\x} \sum_{\d}
+I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
+                       \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
+           &= H_{\x} + H_{\d} - H_{\x, \d}.
 \end{align}
+For present purposes, it is more conveniently written as
+\begin{align}
+I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
+              \log\left[\frac{p(\d | \x)}{p(\d)}\right].
+\end{align}
+As with the conditional entropy, we see that the difficulty appears in the
+form of the log of the marginal likelihood,
+$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.
+
+For experimental design purposes, maximising either $I_{\x, \d}$ or
+$H_{\x | \d}$ will produce the same result. {\todo Mention reference priors?}
 

### 2017-01-15 cpp/tests/sinusoid.cpp
Check generate as well

@@ -30,8 +30,19 @@ int main()
         {
             s.print(fout);
             fout << std::endl;
-            std::cout << "Done " << (i+1) << " steps." << std::endl;
+            std::cout << "Done " << (i+1) << " MCMC steps." << std::endl;
         }
+    }
+
+    // Now do independent generating
+    steps /= thin;
+    for(int i=0; i<steps; ++i)
+    {
+        s.generate(rng);
+
+        s.print(fout);
+        fout << std::endl;
+        std::cout << "Done " << (i+1) << " generates." << std::endl;
     }
 
     fout.close();

### 2017-01-15 cpp/tests/SamplingTest.hpp
Moved test function to a template function

@@ -1,0 +1,60 @@
+#ifndef InfoNest_SamplingTest
+#define InfoNest_SamplingTest
+
+// Includes
+#include <iostream>
+#include <fstream>
+#include <cmath>
+#include "RNG.h"
+
+// Do MCMC sampling and iid sampling on a given example.
+template<class Example>
+void sampling_test()
+{
+    // Make an RNG
+    InfoNest::RNG rng(time(0));
+
+    // Make and initialise a Sinusoid
+    Example s;
+    s.generate(rng);
+
+    // Output file
+    std::fstream fout("output.txt", std::ios::out);
+
+    // Do Metropolis steps
+    int steps = 1000000;
+    int thin = 100;
+
+    for(int i=0; i<steps; ++i)
+    {
+        auto proposal = s;
+        double logH = proposal.perturb(rng);
+
+        if(rng.rand() <= exp(logH))
+            s = proposal;
+
+        if((i+1) % thin == 0)
+        {
+            s.print(fout);
+            fout << std::endl;
+            std::cout << "Done " << (i+1) << " MCMC steps." << std::endl;
+        }
+    }
+
+    // Now do independent generating
+    steps /= thin;
+    for(int i=0; i<steps; ++i)
+    {
+        s.generate(rng);
+
+        s.print(fout);
+        fout << std::endl;
+        std::cout << "Done " << (i+1) << " generates." << std::endl;
+    }
+
+    fout.close();
+
+} // void sampling_test()
+
+#endif
+

### 2017-01-15 cpp/tests/sinusoid.cpp

@@ -1,51 +1,9 @@
-#include <iostream>
-#include <fstream>
+#include "SamplingTest.hpp"
 #include "../examples/Sinusoid.h"
 
 int main()
 {
-    // Make an RNG
-    InfoNest::RNG rng(time(0));
-
-    // Make and initialise a Sinusoid
-    InfoNest::Sinusoid s;
-    s.generate(rng);
-
-    // Output file
-    std::fstream fout("output.txt", std::ios::out);
-
-    // Do Metropolis steps
-    int steps = 1000000;
-    int thin = 100;
-
-    for(int i=0; i<steps; ++i)
-    {
-        auto proposal = s;
-        double logH = proposal.perturb(rng);
-
-        if(rng.rand() <= exp(logH))
-            s = proposal;
-
-        if((i+1) % thin == 0)
-        {
-            s.print(fout);
-            fout << std::endl;
-            std::cout << "Done " << (i+1) << " MCMC steps." << std::endl;
-        }
-    }
-
-    // Now do independent generating
-    steps /= thin;
-    for(int i=0; i<steps; ++i)
-    {
-        s.generate(rng);
-
-        s.print(fout);
-        fout << std::endl;
-        std::cout << "Done " << (i+1) << " generates." << std::endl;
-    }
-
-    fout.close();
+    sampling_test<InfoNest::Sinusoid>();
 
     return 0;
 }

### 2017-01-15 cpp/examples/Sinusoid.cpp
Work in terms of log10 period

@@ -16,7 +16,7 @@ Sinusoid::Sinusoid()
 void Sinusoid::generate(RNG& rng)
 {
     A = -log(1.0 - rng.rand());
-    T = exp(log(1E-3 * t_range) + log(1E3) * rng.rand());
+    log10_period = log10(1E-3 * t_range) + 3.0 * rng.rand();
     phi = 2.0 * M_PI * rng.rand();
 
     for(double& nn : n)
@@ -29,7 +29,7 @@ void Sinusoid::generate(RNG& rng)
 void Sinusoid::assemble()
 {
     double t;
-    double omega = 2*M_PI/T; // Angular frequency
+    double omega = 2*M_PI/pow(10.0, log10_period); // Angular frequency
     for(size_t i=0; i<N; ++i)
     {
         t = t_min + i * dt;
@@ -54,10 +54,8 @@ double Sinusoid::perturb(RNG& rng)
         }
         else if(which == 1)
         {
-            T = log(T);
-            T += log(1E3) * rng.randh();
-            wrap(T, log(1E-3*t_range), log(t_range));
-            T = exp(T);
+            log10_period += 3.0 * rng.randh();
+            wrap(log10_period, log10(1E-3 * t_range), log10(t_range));
         }
         else
         {
@@ -93,9 +91,15 @@ double Sinusoid::perturb(RNG& rng)
 
 void Sinusoid::print(std::ostream& out)
 {
-    out<<A<<' '<<T<<' '<<phi<<' ';
+    out<<A<<' '<<log10_period<<' '<<phi<<' ';
     for(double yy: y)
         out<<yy<<' ';
+}
+
+double Sinusoid::distance(const Sinusoid& s1, const Sinusoid& s2)
+{
+    // For H_T
+    return std::abs(s2.log10_period - s1.log10_period);
 }
 
 } // namespace InfoNest

### 2017-01-15 cpp/examples/Sinusoid.h

@@ -27,8 +27,8 @@ class Sinusoid
         static constexpr double sigma = 1.0;
 
     private:
-        // Amplitude, period, and phase
-        double A, T, phi;
+        // Amplitude, log-period, and phase
+        double A, log10_period, phi;
 
         // Latent coordinates for noise
         std::vector<double> n;

### 2017-01-15 cpp/postprocess.py
Changed printed messages

@@ -27,16 +27,16 @@ def postprocess(num_particles=1, tol=1E-3):
         print("Processed run {k}/{N}.".format(k=i+1, N=reps))
         counts[i] = (run[:,1] > tol).sum()
 
-    print("mean(counts) = {mc}.".format(mc=counts.mean()))
-    print("std (counts) = {sd}.".format(sd=counts.std()))
+    print("mean(depth) = {mc}.".format(mc=counts.mean() / num_particles))
+    print("std (depth) = {sd}.".format(sd=counts.std()  / num_particles))
+    print("sem (depth) = {sem}.".format(\
+            sem=counts.std() / num_particles/np.sqrt(reps)))
 
-    # Estimate of the differential entropy
-    # log(probability) ~= log(2 * tol * density)
-    H = counts.mean() / num_particles + np.log(2 * tol)
-    stderr = (counts.std() / num_particles) / np.sqrt(reps)
-
-    print("H = {H} +- {stderr} nats.".format(H=H, stderr=stderr))
-
+#    # Estimate of the differential entropy
+#    # log(probability) ~= log(2 * tol * density)
+#    H = counts.mean() / num_particles + np.log(2 * tol)
+#    stderr = (counts.std() / num_particles) / np.sqrt(reps)
+#    print("H = {H} +- {stderr} nats.".format(H=H, stderr=stderr))
 #    plt.hist(mean_depth, 100, color=[0.2, 0.2, 0.2])
 #    plt.show()
 

### 2017-01-15 cpp/postprocess.py
Don't use the last run if it's incomplete

@@ -17,15 +17,26 @@ def postprocess(num_particles=1, tol=1E-3):
     start_indices = np.nonzero(output[:,0] == 1)[0]
 
     reps = len(start_indices)
-    counts = np.empty(reps)
+    counts = []
+    last_run_size = None
+
     for i in range(0, reps):
         start_index = start_indices[i]
         if i < reps - 1:
             run = output[start_index : start_indices[i+1], :]
         else:
             run = output[start_index : , :]
+
+        # Don't count the last rep if it's incomplete
+        if last_run_size != None and run.shape[0] != last_run_size:
+            break
+        last_run_size = run.shape[0]
+
+        counts.append((run[:,1] > tol).sum())
         print("Processed run {k}/{N}.".format(k=i+1, N=reps))
-        counts[i] = (run[:,1] > tol).sum()
+
+    # Convert to numpy array
+    counts = np.array(counts)
 
     print("mean(depth) = {mc}.".format(mc=counts.mean() / num_particles))
     print("std (depth) = {sd}.".format(sd=counts.std()  / num_particles))

### 2017-01-16 cpp/postprocess.py
Message about the last run

@@ -38,10 +38,13 @@ def postprocess(num_particles=1, tol=1E-3):
     # Convert to numpy array
     counts = np.array(counts)
 
+    if len(counts) < reps:
+        print("Skipping run {reps} as it is incomplete.".format(reps=reps))
+
     print("mean(depth) = {mc}.".format(mc=counts.mean() / num_particles))
     print("std (depth) = {sd}.".format(sd=counts.std()  / num_particles))
     print("sem (depth) = {sem}.".format(\
-            sem=counts.std() / num_particles/np.sqrt(reps)))
+            sem=counts.std() / num_particles/np.sqrt(len(counts))))
 
 #    # Estimate of the differential entropy
 #    # log(probability) ~= log(2 * tol * density)

### 2017-01-16 cpp/examples/Sinusoid.cpp
Three distance functions

@@ -96,10 +96,41 @@ void Sinusoid::print(std::ostream& out)
         out<<yy<<' ';
 }
 
+double Sinusoid::distance1(const Sinusoid& s1, const Sinusoid& s2)
+{
+    // For H_{x} where x = log10_period.
+    return std::abs(s2.log10_period - s1.log10_period);
+}
+
+double Sinusoid::distance2(const Sinusoid& s1, const Sinusoid& s2)
+{
+    // For H_{data}
+    double rms = 0.0;
+
+    for(size_t i=0; i<N; ++i)
+        rms += pow(s2.y[i] - s1.y[i], 2);
+    rms = sqrt(rms / N);
+
+    return rms;
+}
+
+double Sinusoid::distance3(const Sinusoid& s1, const Sinusoid& s2)
+{
+    // For H_{x, data} where x = log10_period.
+    double rms = 0.0;
+
+    rms += pow(s2.log10_period - s1.log10_period, 2);
+    for(size_t i=0; i<N; ++i)
+        rms += pow(s2.y[i] - s1.y[i], 2);
+    rms = sqrt(rms / (N + 1));
+
+    return rms;
+}
+
+
 double Sinusoid::distance(const Sinusoid& s1, const Sinusoid& s2)
 {
-    // For H_T
-    return std::abs(s2.log10_period - s1.log10_period);
+    return distance3(s1, s2);
 }
 
 } // namespace InfoNest

### 2017-01-16 cpp/examples/Sinusoid.h

@@ -52,9 +52,14 @@ class Sinusoid
         // Printing to stream
         void print(std::ostream& out);
 
-        // Distance function
-        static double distance(const Sinusoid& s1,
-                               const Sinusoid& s2);
+    public:
+        // Distance function (the one that actually gets used)
+        static double distance(const Sinusoid& s1, const Sinusoid& s2);
+
+        // A few options to use for `distance`
+        static double distance1(const Sinusoid& s1, const Sinusoid& s2);
+        static double distance2(const Sinusoid& s1, const Sinusoid& s2);
+        static double distance3(const Sinusoid& s1, const Sinusoid& s2);
 };
 
 } // namespace InfoNest

### 2017-01-16 cpp/Rep.hpp
Count accepts

@@ -184,6 +184,9 @@ void Rep<Particle>::replace(int which)
     double proposal_distance;
     double logA, alpha;
 
+    // Count acceptances
+    int accepts = 0;
+
     // Do the MCMC
     for(size_t i=0; i<mcmc_steps; ++i)
     {
@@ -200,8 +203,14 @@ void Rep<Particle>::replace(int which)
         {
             particles[which] = proposal;
             distances[which] = proposal_distance;
+
+            ++accepts;
         }
     }
+
+//    std::cout<<"Iteration "<<iteration<<". ";
+//    std::cout<<"Acceptance fraction = "<<accepts<<" / "<<mcmc_steps<<".";
+//    std::cout<<std::endl;
 }
 
 

### 2017-01-16 cpp/examples/Sinusoid.cpp
The three distance fucntions

@@ -117,20 +117,15 @@ double Sinusoid::distance2(const Sinusoid& s1, const Sinusoid& s2)
 double Sinusoid::distance3(const Sinusoid& s1, const Sinusoid& s2)
 {
     // For H_{x, data} where x = log10_period.
-    double rms = 0.0;
-
-    rms += pow(s2.log10_period - s1.log10_period, 2);
-    for(size_t i=0; i<N; ++i)
-        rms += pow(s2.y[i] - s1.y[i], 2);
-    rms = sqrt(rms / (N + 1));
-
-    return rms;
+    double d1 = distance1(s1, s2);
+    double d2 = distance2(s1, s2);
+    return (d2 > d1) ? (d2) : (d1);
 }
 
 
 double Sinusoid::distance(const Sinusoid& s1, const Sinusoid& s2)
 {
-    return distance3(s1, s2);
+    return distance1(s1, s2);
 }
 
 } // namespace InfoNest

### 2017-01-17 paper/ms.tex
Work in terms of log10 period
Some results
Numbers and centre for e-research
Some notes and numbers

@@
 
-\documentclass[journal,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
+\documentclass[entropy,article,submit,oneauthor,pdftex,10pt,a4paper]{mdpi}
 %--------------------
@@
 \definecolor{orange}{rgb}{1, 0.5, 0}
+\definecolor{green}{rgb}{0, 0.5, 0}
 
@@
 \newcommand{\todo}{\color{orange} \bf}
+\newcommand{\query}{\color{green} \bf}
 \newcommand{\x}{\boldsymbol{x}}
@@
 
+The priors are
+\begin{align}
+A           &\sim \textnormal{Exponential}(1)  \\
+\log_{10} T &\sim \textnormal{Uniform}(-3, 0)  \\
+\phi        &\sim \textnormal{Uniform}(0, 2\pi)
+\end{align}
+
+$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
+These results were computed using a tolerance of 0.001 for every quantity.
+Results:
+
+% These were produced with a random number seed of 0
+% on commit 25301bb
+% With 10 particles, 1000 mcmc steps per NS iteration, and 1000 reps.
+
+\begin{align}
+H_{\x}     &= 7.345 \pm 0.027 \\  % (max depth was 30)
+H_{\d}     &= 
+H_{\x, \d} &=    
+\end{align}
+
+{\query The multimodality of the posterior in this problem (sometimes) might
+affect mixing sufficiently that the NS gets the incorrect answer. Be
+careful! And maybe use an example where this is unlikely.}
+
 \section{Example 3: TBD}
@@
 Ewan Cameron (Oxford), David Hogg (NYU), Kevin Knuth (SUNY Albany),
-Iain Murray (Edinburgh).
+Iain Murray (Edinburgh). This work was supported by Centre for eResearch
+at the University of Auckland.
 

### 2017-01-17 paper/ms.tex
Numbers
Another number
Calculated mutual info and it is wrong. Need to figure out why

@@
 \begin{align}
-H_{\x}     &= 7.345 \pm 0.027 \\  % (max depth was 30)
-H_{\d}     &= 
-H_{\x, \d} &=    
+H_{\x}     &= 7.345  \pm 0.027 \\  % max depth was 30
+H_{\d}     &= 703.96 \pm 0.31  \\  % max depth was 1000
+H_{\x, \d} &= 712.27 \pm 0.31      % max depth was 1000
 \end{align}
@@
 careful! And maybe use an example where this is unlikely.}
+
+Converting these to differential entropies, we obtain
+\begin{align}
+H'_{\x}     &= 1.131  \pm 0.027 \\  % max depth was 30
+H'_{\d}     &= 697.75 \pm 0.31  \\  % max depth was 1000
+H'_{\x, \d} &= 699.84 \pm 0.31      % max depth was 1000
+\end{align}
+and therefore a mutual information of {\query a negative value! Oh dear!
+Something is wrong.}
 

### 2017-01-17 cpp/examples/Normal.cpp
Gutted Normal class, preparing to make it slightly more complex

@@ -4,6 +4,7 @@ namespace InfoNest
 {
 
 Normal::Normal()
+:xs(100)
 {
 
 }
@@ -11,46 +12,46 @@ Normal::Normal()
 
 void Normal::generate(RNG& rng)
 {
-    n1 = rng.randn();
-    n2 = rng.randn();
-    assemble();
+    mu = 10.0 * rng.randn();
+    for(double& x: xs)
+        x = mu + rng.randn();
 }
 
 double Normal::perturb(RNG& rng)
 {
     double logH = 0.0;
 
-    // Choose a coordinate to perturb
-    int which = rng.rand_int(2);
-    double& n = (which == 0)?(n1):(n2);
+    int proposal_type = rng.rand_int(4);
 
-    logH -= -0.5*n*n;
-    n += rng.randh();
-    logH += -0.5*n*n;
+    if(proposal_type == 0)
+    {
 
-    assemble();
+    }
+    else if(proposal_type == 1)
+    {
+
+    }
+    else if(proposal_type == 2)
+    {
+
+    }
+    else
+    {
+
+    }
+
 
     return logH;
 }
 
-void Normal::assemble()
-{
-    x = 10 * n1;
-    y = x + n2;
-}
-
 void Normal::print(std::ostream& out)
 {
-    out<<x<<' '<<y<<' ';
+    out<<mu<<' ';
 }
 
 double Normal::distance(const Normal& normal1, const Normal& normal2)
 {
-    // Currently set up for H[p(x, y)]
-    double d1 = std::abs(normal1.x - normal2.x);
-    double d2 = std::abs(normal1.y - normal2.y);
-
-    return (d2 > d1)?(d2):(d1);
+    return 0.0;
 }
 
 } // namespace InfoNest

### 2017-01-17 cpp/examples/Normal.h

@@ -10,39 +10,16 @@ namespace InfoNest
 
 /*
 * A 2D gaussian example.
-* p(x)     ~ N(0, 10^2)
-* p(y | x) ~ N(x, 1)
-* ==>
-* p(y)     ~ N(0, sd=sqrt(101))
-
-* H[p(x)]    = 3.72152363.            (differential entropy)
-* H[p(y)]    = 3.72649879.            (differential entropy)
-* H[p(x, y)] = 5.14046216.            (differential entropy)
-* I[x; y]    = 2.3075603.             (mutual information)
-*/
-
-/*
-* Numerical results:
-* NB: these algorithm computes log-probabilities of meeting the tolerance,
-*     so a correction factor of log(2E-3) (log-volume)
-*     has been applied to make them match the above by converting probability
-*     to density. For H[p(x,y)], the correction factor is 2*log(2E-3).
-*  H[p(x)]   = 3.761 +- 0.040
-*  H[p(y)]   = 3.776 +- 0.048
-*  H[p(x,y)] = 5.167 +- 0.055
-*  I[x; y]   = 2.370 +- 0.083 (computed from the above).
+* p(mu)     ~ N(0, 10^2)
+* p(x | mu) ~ N(x, 1),    and there are 100 of them
 */
 
 class Normal
 {
     private:
         // The two actual quantities
-        double x;
-        double y;
-
-        // Latent coordinates
-        double n1, n2;
-        void assemble();
+        double mu;
+        std::vector<double> xs;
 
     public:
         // Do-nothing constructor
@@ -58,7 +35,8 @@ class Normal
         void print(std::ostream& out);
 
         // Distance function
-        static double distance(const Normal& normal1, const Normal& normal2);
+        static double distance(const Normal& normal1,
+                               const Normal& normal2);
 };
 
 } // namespace InfoNest

### 2017-01-17 cpp/examples/Sinusoid.cpp
Renamed choice to proposal_type

@@ -79,9 +79,9 @@ double Sinusoid::perturb(RNG& rng)
 {
     double logH = 0.0;
 
-    int choice = rng.rand_int(4);
+    int proposal_type = rng.rand_int(4);
 
-    if(choice == 0)
+    if(proposal_type == 0)
     {
         // Perturb parameters, changing data along with it
         std::vector<double> mu_old = mu;
@@ -98,7 +98,7 @@ double Sinusoid::perturb(RNG& rng)
 
         calculate_logl();
     }
-    else if(choice == 1)
+    else if(proposal_type == 1)
     {
         // Perturb parameters, keeping data constant
         // (aka Metropolis step of the posterior!)
@@ -109,7 +109,7 @@ double Sinusoid::perturb(RNG& rng)
 
         logH += logl;        
     }
-    else if(choice == 2)
+    else if(proposal_type == 2)
     {
         // Just change one datum
         int which = rng.rand_int(N);

### 2017-01-17 cpp/examples/Sinusoid.cpp
Need to recalculate logl after *every* type of proposal :-/

@@ -95,8 +95,6 @@ double Sinusoid::perturb(RNG& rng)
             n = (y[i] - mu_old[i]) / sigma;
             y[i] = mu[i] + sigma * n;
         }
-
-        calculate_logl();
     }
     else if(proposal_type == 1)
     {
@@ -117,8 +115,6 @@ double Sinusoid::perturb(RNG& rng)
         logH -= -0.5*pow((y[which] - mu[which])/sigma, 2);
         y[which] += sigma * rng.randh();
         logH += -0.5*pow((y[which] - mu[which])/sigma, 2);
-
-        calculate_logl();
     }
     else
     {
@@ -130,10 +126,9 @@ double Sinusoid::perturb(RNG& rng)
             which = rng.rand_int(N);
             y[which] = mu[which] + sigma * rng.randn();
         }
-
-        calculate_logl();
     }
 
+    calculate_logl();
     return logH;
 }
 

### 2017-01-17 cpp/Rep.hpp
Be verbose

@@ -20,6 +20,9 @@ template<class Particle>
 class Rep
 {
     private:
+        // Identity
+        const size_t id;
+
         // Nested Sampling parameters
         const size_t num_particles;
         const size_t mcmc_steps;
@@ -59,7 +62,11 @@ class Rep
     public:        
         // Constructor. You need to pass in the NS parameters
         // and an RNG.
-        Rep(size_t num_particles, size_t mcmc_steps, double depth, RNG& rng);
+        Rep(size_t id,
+            size_t num_particles,
+            size_t mcmc_steps,
+            double depth,
+            RNG& rng);
 
         // Destructor. Closes output file.
         ~Rep();
@@ -77,9 +84,11 @@ class Rep
 /******************************/
 
 template<class Particle>
-Rep<Particle>::Rep(size_t num_particles, size_t mcmc_steps, double depth,
+Rep<Particle>::Rep(size_t id,
+                   size_t num_particles, size_t mcmc_steps, double depth,
                    RNG& rng)
-:num_particles(num_particles)
+:id(id)
+,num_particles(num_particles)
 ,mcmc_steps(mcmc_steps)
 ,depth(depth)
 ,rng(rng)
@@ -208,9 +217,10 @@ void Rep<Particle>::replace(int which)
         }
     }
 
-//    std::cout<<"Iteration "<<iteration<<". ";
-//    std::cout<<"Acceptance fraction = "<<accepts<<" / "<<mcmc_steps<<".";
-//    std::cout<<std::endl;
+    std::cout<<"Rep "<<id<<", ";
+    std::cout<<"iteration "<<iteration<<". ";
+    std::cout<<"Acceptance fraction = "<<accepts<<" / "<<mcmc_steps<<".";
+    std::cout<<std::endl;
 }
 
 

### 2017-01-17 cpp/main.cpp

@@ -30,10 +30,11 @@ int main()
     for(int i=0; i<num_reps; ++i)
     {
         // Create a Rep.
-        InfoNest::Rep<InfoNest::Normal> rep(num_particles,
-                                            mcmc_steps,
-                                            depth,
-                                            rng);
+        InfoNest::Rep<InfoNest::Normal> rep(i+1,
+                                              num_particles,
+                                              mcmc_steps,
+                                              depth,
+                                              rng);
 
         // Initialise and execute it.
         rep.initialise();

### 2017-01-17 cpp/examples/Normal.cpp
Working on aspects of the model, including log likelihood implemented

@@ -15,6 +15,17 @@ void Normal::generate(RNG& rng)
     mu = 10.0 * rng.randn();
     for(double& x: xs)
         x = mu + rng.randn();
+
+    calculate_logl();
+}
+
+void Normal::calculate_logl()
+{
+    logl = 0.0;
+
+    double C = -0.5*log(2*M_PI);
+    for(double x: xs)
+        logl += C - 0.5*pow(x - mu, 2);
 }
 
 double Normal::perturb(RNG& rng)
@@ -25,21 +36,42 @@ double Normal::perturb(RNG& rng)
 
     if(proposal_type == 0)
     {
+        // Perturb parameters, changing data along with it
+        double old_mu = mu;
+
+        logH -= -0.5*pow(mu/10.0, 2);
+        mu += 10.0 * rng.randh();
+        logH += -0.5*pow(mu/10.0, 2);
+
+        double delta_mu = mu - old_mu;
+        for(double& x: xs)
+            x += delta_mu;
 
     }
     else if(proposal_type == 1)
     {
-
+        // Perturb parameters, keeping data constant
+        // (aka Metropolis step of the posterior!)
     }
     else if(proposal_type == 2)
     {
-
+        // Change one data point
+        int which = rng.rand_int(xs.size());
+//        logH -= -0.5*pow(
     }
     else
     {
-
+        // Regenerate one or many data points
+        int reps = (int)pow((double)xs.size(), rng.rand());
+        int which;
+        for(int i=0; i<reps; ++i)
+        {
+            which = rng.rand_int(xs.size());
+            xs[which] = mu + rng.randn();
+        }
     }
 
+    calculate_logl();
 
     return logH;
 }

### 2017-01-17 cpp/examples/Normal.h

@@ -17,9 +17,15 @@ namespace InfoNest
 class Normal
 {
     private:
-        // The two actual quantities
+        // An unknown parameter
         double mu;
+
+        // Data
         std::vector<double> xs;
+
+        // Log-likelihood (useful for one of the proposals)
+        double logl;
+        void calculate_logl();
 
     public:
         // Do-nothing constructor

### 2017-01-17 cpp/examples/Normal.cpp
Finished rewrite (not verified)

@@ -28,6 +28,17 @@ void Normal::calculate_logl()
         logl += C - 0.5*pow(x - mu, 2);
 }
 
+double Normal::perturb_mu(RNG& rng)
+{
+    double logH = 0.0;
+
+    logH -= -0.5*pow(mu/10.0, 2);
+    mu += 10.0 * rng.randh();
+    logH += -0.5*pow(mu/10.0, 2);
+
+    return logH;
+}
+
 double Normal::perturb(RNG& rng)
 {
     double logH = 0.0;
@@ -39,25 +50,35 @@ double Normal::perturb(RNG& rng)
         // Perturb parameters, changing data along with it
         double old_mu = mu;
 
-        logH -= -0.5*pow(mu/10.0, 2);
-        mu += 10.0 * rng.randh();
-        logH += -0.5*pow(mu/10.0, 2);
+        logH += perturb_mu(rng);
 
         double delta_mu = mu - old_mu;
         for(double& x: xs)
             x += delta_mu;
 
+        calculate_logl();
     }
     else if(proposal_type == 1)
     {
         // Perturb parameters, keeping data constant
         // (aka Metropolis step of the posterior!)
+        logH -= logl;
+
+        logH += perturb_mu(rng);
+        
+        calculate_logl();
+        logH += logl;
     }
     else if(proposal_type == 2)
     {
         // Change one data point
         int which = rng.rand_int(xs.size());
-//        logH -= -0.5*pow(
+
+        logH -= -0.5*pow(xs[which] - mu, 2);
+        xs[which] += rng.randh();
+        logH += -0.5*pow(xs[which] - mu, 2);
+
+        calculate_logl();
     }
     else
     {
@@ -69,9 +90,9 @@ double Normal::perturb(RNG& rng)
             which = rng.rand_int(xs.size());
             xs[which] = mu + rng.randn();
         }
-    }
 
-    calculate_logl();
+        calculate_logl();
+    }
 
     return logH;
 }
@@ -83,7 +104,7 @@ void Normal::print(std::ostream& out)
 
 double Normal::distance(const Normal& normal1, const Normal& normal2)
 {
-    return 0.0;
+    return std::abs(normal1.mu - normal2.mu);
 }
 
 } // namespace InfoNest

### 2017-01-17 cpp/examples/Normal.h

@@ -27,6 +27,9 @@ class Normal
         double logl;
         void calculate_logl();
 
+        // Helper for perturb
+        double perturb_mu(RNG& rng);
+
     public:
         // Do-nothing constructor
         Normal();

### 2017-01-17 cpp/tests/normal.cpp
test for normal

@@ -1,0 +1,10 @@
+#include "SamplingTest.hpp"
+#include "../examples/Normal.h"
+
+int main()
+{
+    sampling_test<InfoNest::Normal>();
+
+    return 0;
+}
+

### 2017-01-17 cpp/examples/Normal.cpp
Print xs as well

@@ -100,6 +100,8 @@ double Normal::perturb(RNG& rng)
 void Normal::print(std::ostream& out)
 {
     out<<mu<<' ';
+    for(double x: xs)
+        out<<x<<' ';
 }
 
 double Normal::distance(const Normal& normal1, const Normal& normal2)

### 2017-01-17 cpp/main.cpp
Slightly more depth

@@ -10,7 +10,7 @@ int main()
     // Run parameters
     size_t num_particles = 1;
     size_t mcmc_steps = 1000;
-    double depth = 25.0;
+    double depth = 30.0;
 
     // Create random number generator, seeded with time
     InfoNest::RNG rng(time(0));

### 2017-01-17 cpp/examples/Normal.cpp
Other distance functions

@@ -104,9 +104,28 @@ void Normal::print(std::ostream& out)
         out<<x<<' ';
 }
 
-double Normal::distance(const Normal& normal1, const Normal& normal2)
+double Normal::parameter_distance(const Normal& normal1,
+                                  const Normal& normal2)
 {
     return std::abs(normal1.mu - normal2.mu);
+}
+
+double Normal::data_distance(const Normal& normal1,
+                             const Normal& normal2)
+{
+    double rms = 0.0;
+
+    size_t n = normal1.xs.size();
+    for(size_t i=0; i<n; ++i)
+        rms += pow(normal1.xs[i] - normal2.xs[i], 2);
+    rms = sqrt(rms / n);
+
+    return rms;
+}
+
+double Normal::distance(const Normal& normal1, const Normal& normal2)
+{
+    return parameter_distance(normal1, normal2);    
 }
 
 } // namespace InfoNest

### 2017-01-17 cpp/examples/Normal.h

@@ -12,6 +12,17 @@ namespace InfoNest
 * A 2D gaussian example.
 * p(mu)     ~ N(0, 10^2)
 * p(x | mu) ~ N(x, 1),    and there are 100 of them
+*
+* Differential entropies (true, analytic), in nats.
+* H(mu)     =   3.72152
+* H(x | mu) =   141.894
+* H(mu, x)  =   145.615
+* H(mu | x) = -0.883697
+* I[x, y]   =   4.60522
+*
+* Numerical results (based on prob of meeting tolerance, not density!)
+* H'(mu)    = 9.976 +- 0.032
+* 
 */
 
 class Normal
@@ -29,6 +40,12 @@ class Normal
 
         // Helper for perturb
         double perturb_mu(RNG& rng);
+
+        // Helpers for distance function
+        static double parameter_distance(const Normal& normal1,
+                                         const Normal& normal2);
+        static double data_distance(const Normal& normal1,
+                                    const Normal& normal2);
 
     public:
         // Do-nothing constructor

### 2017-01-17 cpp/examples/Normal.cpp
joint distance function

@@ -123,6 +123,13 @@ double Normal::data_distance(const Normal& normal1,
     return rms;
 }
 
+double Normal::joint_distance(const Normal& normal1, const Normal& normal2)
+{
+    double d1 = parameter_distance(normal1, normal2);
+    double d2 = data_distance(normal1, normal2);
+    return (d1 > d2) ? (d1) : (d2);
+}
+
 double Normal::distance(const Normal& normal1, const Normal& normal2)
 {
     return parameter_distance(normal1, normal2);    

### 2017-01-17 cpp/examples/Normal.h

@@ -46,6 +46,8 @@ class Normal
                                          const Normal& normal2);
         static double data_distance(const Normal& normal1,
                                     const Normal& normal2);
+        static double joint_distance(const Normal& normal1,
+                                     const Normal& normal2);
 
     public:
         // Do-nothing constructor

### 2017-01-17 paper/ms.tex
Working out differential entropy complications.

@@
 \section{Nested Sampling}
+
 Nested Sampling \citep[NS][]{skilling2006nested}
@@
 \subsection{The sequence of $X$ values}
+
 When $N=1$, the sequence of $X$ values is a Poisson process.
@@
 
-\section{Example 1: A trivial bivariate gaussian}
+\subsection{Discrete vs. differential entropies}
+
+Continuous hypothesis spaces are commonly used in practice. Suppose
+$\x$ has probability density $f(\x)$. The algorithm actually estimates
+\begin{align}
+Q &= \int f(\x) \int f(\x')
+            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
+                        \, d\x' \, d\x.
+\end{align}
+If we actually want the differential
+entropy
+\begin{align}
+h_{\x} &= -\int f(x) \log f(x) \, dx
+\end{align}
+we can use the fact that density equals mass divided
+by volume. Assume $\epsilon$ is small, so that
+\begin{align}
+P(d(\x', \x) < \epsilon | \x)
+    &\approx
+    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
+    &= f(\x) \, V
+\end{align}
+where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.
+
+\section{Example 1: Measuring the mean of a gaussian}
 
@@
 posterior still consisted of several widely separated modes!
-
 Perhaps, in some applications, a more appropriate question is
@@
 \end{align}
-
 $A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
@@
 and therefore a mutual information of {\query a negative value! Oh dear!
-Something is wrong.}
+Something is wrong. Update: well I found a bug after this, so it might
+change. :)}
 

### 2017-01-19 paper/ms.tex
Paragraph about volume of an n-ball

@@
 
+If the distance function $d(\x', \x)$ is chosen to be Cartesian,
+the constraint $d(\x', \x) < \epsilon$ corresponds to a ball in the space
+of possible $\x'$ values.
+The log-volume of a ball of radius $\epsilon$ in $n$ dimensions is
+\begin{align}
+\log V(\epsilon; n) &= \frac{n}{2}\log \pi
+                        - \log \Gamma\left(\frac{n}{2} + 1\right)
+                        + n \log \epsilon,
+\end{align}
+and provides the conversion factor between the differential entropy and
+the entropy actually computed from the algorithm.
+
 \section{Example 1: Measuring the mean of a gaussian}

### 2017-01-19 paper/figures/demo.py
A lovely figure

@@ -1,0 +1,51 @@
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+
+"""
+Generate some figures and perhaps a movie.
+"""
+
+def density(x, y):
+    """
+    A 2D joint density used for the demo.
+    """
+    fx   = 1.0/np.sqrt(3*2*np.pi)*np.exp(-0.5*x**2 / 3 **2)
+    fy_x = 1.0/np.sqrt(2*2*np.pi)*np.exp(-0.5*(y - 0.1*x*np.abs(x))**2 / 2**2)
+    return fx * fy_x
+
+def generate(num=1):
+    """
+    Generate from the above density.
+    """
+    x = 3*rng.randn(num)
+    y = 0.1*x*np.abs(x) + 2*rng.randn(num)
+    return {"x": x, "y": y}
+
+
+if __name__ == "__main__":
+    # RNG seed
+    rng.seed(0)
+
+    # A grid
+    x = np.linspace(-10.0, 10.0, 1001)
+    y = np.linspace(-10.0, 10.0, 1001)
+    [x, y] = np.meshgrid(x, y[::-1])
+
+    # Evaluate the pdf
+    f = density(x, y)
+
+    # Generate from the pdf
+    samples = generate(100)
+
+    # Plot the pdf and the samples
+    plt.imshow(density(x, y), cmap="viridis", interpolation="nearest",
+               extent=[x.min(), x.max(), y.min(), y.max()])
+    plt.xlabel("$x$", fontsize=20)
+    plt.ylabel("$y$", fontsize=20)
+    plt.hold(True)
+    plt.plot(samples["x"], samples["y"], "w.", alpha=0.6, markersize=10)
+    plt.axis([x.min(), x.max(), y.min(), y.max()])
+    plt.savefig("figure.png")
+    plt.show()
+

### 2017-01-19 cpp/examples/Normal.h
A number that looks pretty good!

@@ -24,7 +24,7 @@ namespace InfoNest
 * Numerical results (based on prob of meeting tolerance, not density!)
 * H'(mu)    = 9.976  +- 0.032
 * H'(x)     = 928.58 +- 0.47
-*
+* H'(mu, x) = 934.96 +- 0.80
 * log(volume of n-ball) = n/2 * log(pi) - gammaln(n/2 + 1) + n*log(tol)
 *   For mu, it's -6.21461. For x, it's -782.017.
 * 

### 2017-01-19 paper/figures/demo.py
Contours and colours

@@ -39,12 +39,13 @@ if __name__ == "__main__":
     samples = generate(100)
 
     # Plot the pdf and the samples
-    plt.imshow(density(x, y), cmap="viridis", interpolation="nearest",
+    plt.imshow(density(x, y), cmap="Blues", interpolation="nearest",
                extent=[x.min(), x.max(), y.min(), y.max()])
     plt.xlabel("$x$", fontsize=20)
     plt.ylabel("$y$", fontsize=20)
     plt.hold(True)
-    plt.plot(samples["x"], samples["y"], "w.", alpha=0.6, markersize=10)
+    plt.contour(x, y, density(x, y), 50, colors=["k"], alpha=0.1)
+    plt.plot(samples["x"], samples["y"], "r.", alpha=0.3, markersize=10)
     plt.axis([x.min(), x.max(), y.min(), y.max()])
     plt.savefig("figure.png")
     plt.show()

### 2017-01-22 paper/ms.tex
Wrote a bit about the gaussian problem
results
Differential entropy of p(d) for normal example

@@
 
+Consider the basic problem of inferring a quantity $\mu$ from
+100 observations $\x = \{x_1, ..., x_{100}\}$ whose probability distribution
+(conditional on $\mu$) is
+\begin{align}
+p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
+\end{align}
+If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
+\begin{align}
+\mu | \x &\sim \textnormal{Normal}\left(
+                                       \frac{1}{100}\sum_{i=1}^{100} x_i,
+                                       \left[\frac{10}{\sqrt{10001}}\right]^2
+                                       \right).
+\end{align}
+
+\begin{align}
+H_{\d} &=% 945.42 \pm 1.02 \\ % 1000 mcmc steps
+         % 929.26 \pm 1.00 \\ % 2000 mcmc steps
+         % 927.40 \pm 0.98 \\ % 3000 mcmc steps
+         % 927.26 \pm 1.00 \\ % 4000 mcmc steps
+          928.21 \pm 1.00    % 5000 mcmc steps
+\end{align}
+The log-volume of a 100-dimensional ball of radius $10^{-3}$ is
+-782.017. Therefore, the measured differential entropy of
+$p(\d)$ is $928.21 - 782.017 = 146.19 \pm 1.00$, consistent with the
+true value of $146.499$.
+
 \section{Example 2: Measuring the period of an oscillating signal}

### 2017-02-02 paper/ms.tex
updated numbers - challenging

@@
 % These were produced with a random number seed of 0
-% on commit 25301bb
-% With 10 particles, 1000 mcmc steps per NS iteration, and 1000 reps.
+% on commit 947317948d04255ea0baff7a7d3e357b0c2f7fcb
+% With 10 particles, 1000 mcmc steps per NS iteration.
 
-\begin{align}
-H_{\x}     &= 7.345  \pm 0.027 \\  % max depth was 30
-H_{\d}     &= 703.96 \pm 0.31  \\  % max depth was 1000
-H_{\x, \d} &= 712.27 \pm 0.31      % max depth was 1000
+\begin{align}                    % (num_particles, mcmc_steps, max_depth, reps)
+H_{\x}    &= 7.328 \pm 0.025 \\% (100, 1000, 30.0,  122)
+H_{\d}    &= 960.4 \pm 3.7   \\% (100, 1000, 1200.0, 99)
+H_{\x,\d} &= 968.6 \pm 3.4   \\% (100, 1000, 1200.0, 99)
 \end{align}
@@
 
-Converting these to differential entropies, we obtain
+To convert these to differential entropies, we need to add the appropriate
+log-volume to each. These are $-6.215$, $-782.02$, and
+$-788.23$ respectively.
 \begin{align}
-H'_{\x}     &= 1.131  \pm 0.027 \\  % max depth was 30
-H'_{\d}     &= 697.75 \pm 0.31  \\  % max depth was 1000
-H'_{\x, \d} &= 699.84 \pm 0.31      % max depth was 1000
+H'_{\x}     &= 1.113 \pm 0.025 \\
+H'_{\d}     &= 178.4 \pm 3.7   \\
+H'_{\x, \d} &= 180.4 \pm 3.4
 \end{align}

### 2017-02-04 paper/ms.tex
Changed x to theta, and also wrote in a few more relations
Updated example to make it more computable

@@
 \newcommand{\query}{\color{green} \bf}
-\newcommand{\x}{\boldsymbol{x}}
+\newcommand{\x}{\boldsymbol{\theta}}
 
@@
 
-In a standard situation where Bayesian inference would be used,
-the unknown quantities are the parameters $\x$ and the data $\d$
-(which subsequently becomes known). The prior information leads the
+Bayesian inference is the use of probability theory to
+describe uncertainty, often about unknown quantities
+(``parameters'') $\x$. Some data $\d$, initially unknown but
+thought to be relevant to $\x$, is obtained.
+The prior information leads the
 user to specify a prior distribution $p(\x)$ for the unknown parameters,
 along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
-of how the data is related to the parameters. By the product rule, this yields
-a {\em joint prior}
+of how the data is related to the parameters
+(i.e., if the parameters where known, what data would likely be observed?). By the product rule, this yields a {\em joint prior}
 \begin{align}
@@
 \end{align}
+which is the starting point for inference.
 
@@
 all possible datasets which might be observed.
-
 This is difficult to compute because the expression
@@
 The mutual information is another way of describing the relevance of the
-data to the parameters.
-
-Its definition is
+data to the parameters. Its definition, and relation to other quantities, is
 \begin{align}
@@
                        \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
-           &= H_{\x} + H_{\d} - H_{\x, \d}.
+           &= H_{\x} + H_{\d} - H_{\x, \d}\\
+           &= H_{\d} - H_{\d | \x}\\
+           &= H_{\x} - H_{\x | \d}.
 \end{align}
@@
 A           &\sim \textnormal{Exponential}(1)  \\
-\log_{10} T &\sim \textnormal{Uniform}(-3, 0)  \\
+\log_{10} T &\sim \textnormal{Uniform}(-1, 0)  \\
 \phi        &\sim \textnormal{Uniform}(0, 2\pi)
@@
 $A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
+The prior standard deviation of the noise, $\sigma$, was set to 0.1.
 These results were computed using a tolerance of 0.001 for every quantity.
@@
 \begin{align}                    % (num_particles, mcmc_steps, max_depth, reps)
-H_{\x}    &= 7.328 \pm 0.025 \\% (100, 1000, 30.0,  122)
-H_{\d}    &= 960.4 \pm 3.7   \\% (100, 1000, 1200.0, 99)
-H_{\x,\d} &= 968.6 \pm 3.4   \\% (100, 1000, 1200.0, 99)
+H_{\x}    &= 6.223 \pm 0.021 \\% (100, 1000, 20.0,  147)
+H_{\d}    &=    \\% (100, 1000, 1200.0, 99)
+H_{\x,\d} &=    \\% (100, 1000, 1200.0, 99)
 \end{align}

### 2017-02-04 cpp/examples/Sinusoid.cpp

@@ -16,7 +16,7 @@ Sinusoid::Sinusoid()
 void Sinusoid::generate(RNG& rng)
 {
     A = -log(1.0 - rng.rand());
-    log10_period = log10(1E-3 * t_range) + 3.0 * rng.rand();
+    log10_period = log10(0.1*t_range) + 1.0*rng.rand();
     phi = 2.0 * M_PI * rng.rand();
 
     calculate_mu();
@@ -53,8 +53,8 @@ double Sinusoid::perturb_parameters(RNG& rng)
     }
     else if(which == 1)
     {
-        log10_period += 3.0 * rng.randh();
-        wrap(log10_period, log10(1E-3 * t_range), log10(t_range));
+        log10_period += 1.0*rng.randh();
+        wrap(log10_period, log10(0.1*t_range), log10(t_range));
     }
     else
     {

### 2017-02-04 cpp/examples/Sinusoid.h

@@ -24,7 +24,7 @@ class Sinusoid
         static constexpr double dt = t_range / (N - 1);
 
         // Noise sd
-        static constexpr double sigma = 1.0;
+        static constexpr double sigma = 0.1;
 
     private:
         // Amplitude, log-period, and phase

### 2017-02-04 paper/ms.tex
impressive integration skillz

@@
 \newcommand{\x}{\boldsymbol{\theta}}
+\newcommand{\n}{\boldsymbol{\eta}}
 
@@
 
+The computational resources needed to compute these quantities are quite large,
+though the algorithm can be parallelised trivially. However, the problem
+being solved is not trivial. The mutual information of $\x$ and
+$\d$ is
+\begin{align}
+I_{\x, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
+\end{align}
+However, $\x$ is one of several parameters. The others, $A$, and $\phi$,
+were implicitly integrated out. Denoting the nuisance parameters by
+$\n$, the mutual information that has been calculated is
+\begin{align}
+I_{\x, \d} &= \int p(\x, \d)
+  \log\frac{\int p(\x, \n, \d) \, d\n }
+           {p(\x) \int p(\x', \n)p(\d | \x', \n) \, d\x' \, d\n}
+  \, d\x \, d\d.
+\end{align}
+It should not be considered surprising that this is a challenging task.
+
 \section{Example 3: TBD}

### 2017-02-04 paper/ms.tex
New numbers

@@
 % These were produced with a random number seed of 0
-% on commit 947317948d04255ea0baff7a7d3e357b0c2f7fcb
-% With 10 particles, 1000 mcmc steps per NS iteration.
+% on commit 44525f13edd0d29de43e703db17c594fc0d5045c
 
@@
 H_{\x}    &= 6.223 \pm 0.021 \\% (100, 1000, 20.0,  147)
-H_{\d}    &=    \\% (100, 1000, 1200.0, 99)
-H_{\x,\d} &=    \\% (100, 1000, 1200.0, 99)
+H_{\d}    &= 805.1 \pm 9.7   \\% (100, 1000, 1000.0, 52)
+H_{\x,\d} &= 873   \pm 25    \\% (100, 1000, 1200.0, 12)
 \end{align}
@@
 $-788.23$ respectively.
-\begin{align}
-H'_{\x}     &= 1.113 \pm 0.025 \\
-H'_{\d}     &= 178.4 \pm 3.7   \\
-H'_{\x, \d} &= 180.4 \pm 3.4
-\end{align}
+%\begin{align}
+%H'_{\x}     &= 1.113 \pm 0.025 \\
+%H'_{\d}     &= 178.4 \pm 3.7   \\
+%H'_{\x, \d} &= 180.4 \pm 3.4
+%\end{align}
 and therefore a mutual information of {\query a negative value! Oh dear!

### 2017-02-04 paper/ms.tex
Waxing about precisional questions

@@
 
+\appendix
+\section{Precisional questions}
+
+The central issue about the value of a parameter $\x$ asks
+``what is the value of $\x$, precisely?''. However, in practice we often
+don't need or want to know $\x$ to arbitrary precision. Using the central
+issue can also lead to counterintuitive results if you don't keep its
+definition in mind. For example, suppose $\x$ could take any integer value
+from 1 to 1 billion. If we learned that the last digit of $\x$ was a 9, we
+will have ruled out nine tenths of the possibilities, and therefore obtained
+a lot of information about the central issue. However, in many applications
+this ``information'' would be basically useless.
+
+The reason is that we often want to ask a slightly different question.
+For example, suppose we want to know the value of $\x$ to within a tolerance
+of $\pm 5$. The following statements would resolve the issue:
+\begin{itemize}
+\item $\x \in \{1, ..., 10\}$ and anything that implies it
+\item $\x \in \{2, ..., 11\}$ and anything that implies it
+...
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-02-04 paper/ms.tex
Working on the appendix

@@
 definition in mind. For example, suppose $\x$ could take any integer value
-from 1 to 1 billion. If we learned that the last digit of $\x$ was a 9, we
+from 1 to 1 billion. If we learned the last digit of $\x$, we
 will have ruled out nine tenths of the possibilities, and therefore obtained
-a lot of information about the central issue. However, in many applications
-this ``information'' would be basically useless.
+a lot of information about the central issue. However,
+this information might be useless {\em for practical purposes}.
+If we learned the final
+digit was a 9, $\x$ could still be 9, or 19, or 29, any number ending in 9
+up to 999,999,999.
 
-The reason is that we often want to ask a slightly different question.
+In practice, we may
+want to ask a question that is different from the central issue.
 For example, suppose we want to know the value of $\x$ to within a tolerance
-of $\pm 5$. The following statements would resolve the issue:
+of $\pm 5$. Any of the following statements would resolve the issue:
 \begin{itemize}
-\item $\x \in \{1, ..., 10\}$ and anything that implies it
-\item $\x \in \{2, ..., 11\}$ and anything that implies it
+\item $\x \in \{1, ..., 11\}$ and anything that implies it
+\item $\x \in \{2, ..., 12\}$ and anything that implies it
+\item $\x \in \{3, ..., 13\}$ and anything that implies it\\
 ...
+\item $\x \in \{999,999,990, ..., 1,000,000,000\}$ and anything that implies it.
 \end{itemize}

### 2017-02-04 paper/ms.tex
Mention common random number technique

@@
 
+The mutual information is equal to $H_{\x} + H_{\d} - H_{\x, \d}$. The
+uncertainties in each of these three quantities is quite large, and would lead
+to a large uncertainty on $I_{\x, \d}$ if computed naively. To reduce this
+uncertainty, we computed the three entropies using a common random number
+generator for the reference pairs. This is an example of the
+{\em common random numbers} technique of variance reduction.
+
 The computational resources needed to compute these quantities are quite large,

### 2017-02-06 cpp/main.cpp
demonstrate on p(d) for Normal example

@@ -10,7 +10,7 @@ int main()
 {
     // What example are we using? The class and the distance function.
     typedef InfoNest::Normal TheExample;
-    const auto& dist_func = TheExample::parameter_distance;
+    const auto& dist_func = TheExample::data_distance;
 
     // Create random number generators
     unsigned long seed1 = 0;
@@ -19,9 +19,9 @@ int main()
     InfoNest::RNG rng2(seed2);
 
     // Define run parameters
-    constexpr double depth         = 30.0;
+    constexpr double depth         = 1200.0;
     constexpr size_t num_reps      = 1000;
-    constexpr size_t num_particles = 100;
+    constexpr size_t num_particles = 1;
     constexpr size_t mcmc_steps    = 1000;
 
     // Do the run.

### 2017-02-06 paper/ms.tex
fixed up partition questions

@@
 issue can also lead to counterintuitive results if you don't keep its
-definition in mind. For example, suppose $\x$ could take any integer value
-from 1 to 1 billion. If we learned the last digit of $\x$, we
+definition in mind. For example, suppose $x$ could take any integer value
+from 1 to 1 billion. If we learned the last digit of $x$, we
 will have ruled out nine tenths of the possibilities, and therefore obtained
@@
 If we learned the final
-digit was a 9, $\x$ could still be 9, or 19, or 29, any number ending in 9
+digit was a 9, $x$ could still be 9, or 19, or 29, any number ending in 9
 up to 999,999,999.
@@
 want to ask a question that is different from the central issue.
-For example, suppose we want to know the value of $\x$ to within a tolerance
-of $\pm 5$. Any of the following statements would resolve the issue:
+For example, suppose we want to know the value of $x$ to within a tolerance
+of $\pm 1$. Any of the following statements would resolve the issue:
 \begin{itemize}
-\item $\x \in \{1, ..., 11\}$ and anything that implies it
-\item $\x \in \{2, ..., 12\}$ and anything that implies it
-\item $\x \in \{3, ..., 13\}$ and anything that implies it\\
-...
-\item $\x \in \{999,999,990, ..., 1,000,000,000\}$ and anything that implies it.
+\item $x \in \{1, 2, 3\}$, $x \notin \{1, 2, 3\}$,
+                or anything that implies one of these
+\item $x \in \{2, 3, 4\}$, $x \notin \{2, 3, 4\}$,
+                or anything that implies one of these ...
+\item ... $x \in \{999999998, 999999999, 1000000000\}$,
+            $x \notin \{999999998, 999999999, 1000000000\}$,
+                or anything that implies one of these
 \end{itemize}

### 2017-02-06 cpp/main.cpp
Probably a sensible default

@@ -21,8 +21,8 @@ int main()
     // Define run parameters
     constexpr double depth         = 1200.0;
     constexpr size_t num_reps      = 1000;
-    constexpr size_t num_particles = 1;
-    constexpr size_t mcmc_steps    = 1000;
+    constexpr size_t num_particles = 10;
+    constexpr size_t mcmc_steps    = 10000;
 
     // Do the run.
     InfoNest::execute<TheExample>(rng1, rng2, depth, num_reps, num_particles,

### 2017-02-08 paper/ms.tex
Final number
An abstract

@@
 % Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{}
+\abstract{The Shannon entropy, and related quantities such as mutual
+information, can be used to quantify uncertainty and relevance.
+However, in practice, it can be difficult to compute these quantities
+for arbitrary probability distributions, particularly if some probability
+mass functions or densities cannot be evaluated. This paper introduces a
+computational approach to evaluating entropies of probability distributions
+that can be sampled. The method is based on Nested Sampling, and is related to Approximate Bayesian Computation. I demonstrate the method
+on a simple gaussian example where the key quantities are known analytically,
+and an experimental design example about scheduling observations in order
+to measure the period of an oscillating signal.}
 
 % Keywords: add 3 to 10 keywords
-\keyword{nested sampling; information theory; monte carlo}
+\keyword{information theory; entropy; mutual information; monte carlo;
+nested sampling; bayesian inference}
 
@@
 
+The result was
 \begin{align}
-H_{\d} &=% 945.42 \pm 1.02 \\ % 1000 mcmc steps
-         % 929.26 \pm 1.00 \\ % 2000 mcmc steps
-         % 927.40 \pm 0.98 \\ % 3000 mcmc steps
-         % 927.26 \pm 1.00 \\ % 4000 mcmc steps
-          928.21 \pm 1.00    % 5000 mcmc steps
+H_{\d} &= 928.80 \pm 0.38.
 \end{align}
@@
 -782.017. Therefore, the measured differential entropy of
-$p(\d)$ is $928.21 - 782.017 = 146.19 \pm 1.00$, consistent with the
+$p(\d)$ is $928.80 - 782.017 = 146.79 \pm 0.38$, consistent with the
 true value of $146.499$.

### 2017-02-08 paper/ms.tex
Bit of work on the intro

@@
 \section{Introduction}
+\noindent
+If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
+the Shannon entropy \citep{shannon} is defined as
+\begin{align}
+H_{x} &= -\sum_{x} p(x) \log p(x)
+\end{align}
+where the sum is over all of the possible values of $x$ under consideration.
+Throughout this paper, I use the compact
+`overloaded' notation for probability distributions, writing $p(x)$ instead
+of the more cumbersome $P(X=x)$.
+This probability distribution is implicitly conditional on some prior
+information, which has been omitted from the notation.
 
-If an unknown quantity $\x$ has a discrete probability distribution $p(\x)$,
-the entropy is
-\begin{align}
-H_{\x} &= -\sum_{\x} p(\x) \log p(\x)
-\end{align}
-where the sum is over all of the possible values of $\x$ under consideration.
-This probability distribution is implicitly conditional on some prior
-information, which is have omitted from the notation.
+\citet{knuth} has shown that the entropy quantifies the degree to which the
+question ``which value of $x$ is true?'' remains unanswered by the prior
+information.
 
@@
 \end{align}
-which is the starting point for inference.
+which is the starting point for inference \citep{caticha}.
 

### 2017-02-08 paper/ms.tex
New priors

@@
 \subsection{The relevance of data}
-
 The degree to which the question ``what is the value of $\x$?'' is
@@
 
-The priors are
+The signal has the form
 \begin{align}
-A           &\sim \textnormal{Exponential}(1)  \\
+y(t) &= A \sin \left(\frac{2\pi t}{T} + \phi\right)
+\end{align}
+where $A$ is the amplitude, $T$ is the period, and $\phi$ is the phase.
+I assumed the following priors:
+\begin{align}
+\log_e{A}   &\sim \textnormal{Normal}\left(0, 0.1^2\right)  \\
 \log_{10} T &\sim \textnormal{Uniform}(-1, 0)  \\
@@
 \end{align}
+and the following conditional prior for the data:
+\begin{align}
+d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
+\end{align}
+That is, the data is just the signal $y(t)$ observed at particular times
+$\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
+
 $A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.

### 2017-02-08 cpp/examples/Sinusoid.cpp

@@ -15,8 +15,8 @@ Sinusoid::Sinusoid()
 
 void Sinusoid::generate(RNG& rng)
 {
-    A = -log(1.0 - rng.rand());
-    log10_period = log10(0.1*t_range) + 1.0*rng.rand();
+    A = exp(0.1*rng.randn());
+    log10_period = log10(1E-3*t_range) + log10(1E3)*rng.rand();
     phi = 2.0 * M_PI * rng.rand();
 
     calculate_mu();
@@ -46,15 +46,18 @@ double Sinusoid::perturb_parameters(RNG& rng)
     int which = rng.rand_int(3);
     if(which == 0)
     {
-        A = 1.0 - exp(-A);
-        A += rng.randh();
-        wrap(A, 0.0, 1.0);
-        A = -log(1.0 - A);
+        A = log(A);
+
+        logH -= -0.5*pow(A/0.1, 2);
+        A += 0.1*rng.randh();
+        logH += -0.5*pow(A/0.1, 2);
+
+        A = exp(A);
     }
     else if(which == 1)
     {
-        log10_period += 1.0*rng.randh();
-        wrap(log10_period, log10(0.1*t_range), log10(t_range));
+        log10_period += log10(1E3)*rng.randh();
+        wrap(log10_period, log10(1E-3*t_range), log10(t_range));
     }
     else
     {

### 2017-02-09 paper/ms.tex
Some sinusoidal results

@@
 
-$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
-The prior standard deviation of the noise, $\sigma$, was set to 0.1.
+Letting $\tau = \log_{10} T$, and treating
+$A$ and $\phi$ are nuisance parameters, I calculated the entropies
+of $\tau$, $\d$, and $(\tau, \d)$.
 These results were computed using a tolerance of 0.001 for every quantity.
-Results:
-
-% These were produced with a random number seed of 0
-% on commit 44525f13edd0d29de43e703db17c594fc0d5045c
-
-\begin{align}                  % (num_particles, mcmc_steps, max_depth, reps)
-H_{\x}    &= 6.223 \pm 0.021 \\% (100, 1000, 20.0,  147)
-H_{\d}    &= 805.1 \pm 9.7   \\% (100, 1000, 1000.0, 52)
-H_{\x,\d} &= 873   \pm 25    \\% (100, 1000, 1200.0, 12)
+The results were
+\begin{align}
+H_{\tau}    &= 7.337 \pm 0.028 \\
+H_{\d}      &= 904.9 \pm 5.4   \\ % Interim result only
+H_{\tau,\d} &= 900.4 \pm 4.7      % Interim result only
 \end{align}
@@
 affect mixing sufficiently that the NS gets the incorrect answer. Be
-careful! And maybe use an example where this is unlikely.}
+careful! And maybe use an example where this is unlikely, or double
+check it using brute force.}
 
 To convert these to differential entropies, we need to add the appropriate
-log-volume to each. These are $-6.215$, $-782.02$, and
-$-788.23$ respectively.
-%\begin{align}
-%H'_{\x}     &= 1.113 \pm 0.025 \\
+log-volume to each. These are $-6.215$, $-790.32$, and
+$-796.53$ respectively, so the differential entropies are:
+\begin{align}
+H'_{\tau}     &= 1.122 \pm 0.028 % \\
 %H'_{\d}       &= 178.4 \pm 3.7   \\
-%H'_{\x, \d} &= 180.4 \pm 3.4
-%\end{align}
-and therefore a mutual information of {\query a negative value! Oh dear!
-Something is wrong. Update: well I found a bug after this, so it might
-change. :)}
-
-The mutual information is equal to $H_{\x} + H_{\d} - H_{\x, \d}$. The
+%H'_{\tau, \d} &= 180.4 \pm 3.4
+\end{align}
+The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
 uncertainties in each of these three quantities is quite large, and would lead
-to a large uncertainty on $I_{\x, \d}$ if computed naively. To reduce this
+to a large uncertainty on $I_{\tau, \d}$ if computed naively. To reduce this
 uncertainty, we computed the three entropies using a common random number
@@
 {\em common random numbers} technique of variance reduction.
+The resulting mutual information is
+$I_{\tau, \d} = 12.1 \pm 3.1$ nats {\todo (interim result only)}.
 

### 2017-02-09 paper/ms.tex
Interim results

@@
 H_{\tau}    &= 7.337 \pm 0.028 \\
-H_{\d}      &= 904.9 \pm 5.4   \\ % Interim result only
-H_{\tau,\d} &= 900.4 \pm 4.7      % Interim result only
+H_{\d}      &= 905.9 \pm 1.9   \\ % Interim result only
+H_{\tau,\d} &= 902.2 \pm 1.7      % Interim result only
 \end{align}
@@
 H'_{\tau}     &= 1.122 \pm 0.028 % \\
-%H'_{\d}       &= 178.4 \pm 3.7   \\
-%H'_{\tau, \d} &= 180.4 \pm 3.4
+H'_{\d}       &= 115.5 \pm 1.9   \\
+H'_{\tau, \d} &= 105.7 \pm 1.7
 \end{align}
@@
 The resulting mutual information is
-$I_{\tau, \d} = 12.1 \pm 3.1$ nats {\todo (interim result only)}.
+$I_{\tau, \d} = 11.1 \pm 1.1$ nats {\todo (interim result only)}.
 

### 2017-02-09 cpp/examples/Sinusoid.cpp
Precompute timestamps

@@ -7,10 +7,18 @@ namespace InfoNest
 {
 
 Sinusoid::Sinusoid()
-:mu(N)
+:t(N)
+,mu(N)
 ,y(N)
 {
+    for(size_t i=0; i<N; ++i)
+    {
+        // For evenly spaced points
+        t[i] = t_min + i * dt;
 
+        // For unevenly spaced points
+        // t[i] = t_min + pow((i + 0.5)/N, 3)*t_range;
+    }
 }
 
 void Sinusoid::generate(RNG& rng)
@@ -29,13 +37,9 @@ void Sinusoid::generate(RNG& rng)
 
 void Sinusoid::calculate_mu()
 {
-    double t;
     double omega = 2*M_PI/pow(10.0, log10_period); // Angular frequency
     for(size_t i=0; i<N; ++i)
-    {
-        t = t_min + i * dt;
-        mu[i] = A * sin(omega*t + phi);
-    }
+        mu[i] = A * sin(omega*t[i] + phi);
 }
 
 double Sinusoid::perturb_parameters(RNG& rng)

### 2017-02-09 cpp/examples/Sinusoid.h

@@ -31,6 +31,7 @@ class Sinusoid
         double A, log10_period, phi;
 
         // Model curve
+        std::vector<double> t;
         std::vector<double> mu;
 
         // The data

### 2017-02-11 paper/ms.tex
Added check result
More numbers

@@
 H_{\tau}    &= 7.337 \pm 0.028 \\
-H_{\d}      &= 905.9 \pm 1.9   \\ % Interim result only
+H_{\d}      &= 905.9 \pm 1.9   \\ % Interim result. Check gave 903.3 +- 10.9
 H_{\tau,\d} &= 902.2 \pm 1.7      % Interim result only
@@
 \begin{align}
-H'_{\tau}     &= 1.122 \pm 0.028 % \\
+H'_{\tau}     &= 1.122 \pm 0.028 \\
 H'_{\d}       &= 115.5 \pm 1.9   \\
@@
 It should not be considered surprising that this is a challenging task.
+
+\subsection{Uneven observations}
+The results for the uneven observing schedule are
+\begin{align}
+I_{\x, \d} &= 12.09 \pm 0.82
+\end{align}
+which appears to be higher than under the even observing schedule,
+though the uncertainties
+are not small enough to be certain of this.
 

### 2017-02-13 paper/ms.tex
Some final numbers
New results

@@
 H_{\tau}    &= 7.337 \pm 0.028 \\
-H_{\d}      &= 905.9 \pm 1.9   \\ % Interim result. Check gave 903.3 +- 10.9
-H_{\tau,\d} &= 902.2 \pm 1.7      % Interim result only
+H_{\d}      &= 903.8 \pm 1.1   \\
+H_{\tau,\d} &= 900.8 \pm 1.1      % Interim result only
 \end{align}
@@
 The resulting mutual information is
-$I_{\tau, \d} = 11.1 \pm 1.1$ nats {\todo (interim result only)}.
+$I_{\tau, \d} = 10.35 \pm 0.61$ nats.
 
@@
 though the algorithm can be parallelised trivially. However, the problem
-being solved is not trivial. The mutual information of $\x$ and
+being solved is not trivial. The mutual information of $\tau$ and
 $\d$ is
 \begin{align}
-I_{\x, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
+I_{\tau, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
 \end{align}
@@
 \subsection{Uneven observations}
-The results for the uneven observing schedule are
+The results for the uneven observing schedule are as follows:
 \begin{align}
-I_{\x, \d} &= 12.09 \pm 0.82
+H'_{\tau}     &= 1.122 \pm 0.028 \\
+H'_{\d}       &= 106.8 \pm 1.1   \\
+H'_{\tau, \d} &= 97.0 \pm 1.0
+I_{\tau, \d} &= 10.9 \pm 0.4
 \end{align}

### 2017-02-13 paper/ms.tex
Wrote a tiny bit of NS prelude stuff

@@
 
-When $N=1$, the sequence of $X$ values is a Poisson process.
-
-\citet{Walter2015} showed how this view of the NS sequence of points
+When $N=1$, the sequence of $\ln X$ values is a Poisson process with rate
+1. \citet{Walter2015} showed how this view of the NS sequence of points
 can be used to construct a
@@
 quantities.
+
+Given the definition of the NS algorithm,
+the probability distribution for the proportion $t$ of the prior mass
+retained on a given iteration is
+\begin{align}
+t_i &\sim \textnormal{Beta}(N, 1),
+\end{align}
+and the probability distribution for the sequence of $X$-values is
+obtained from
+\begin{align}
+X_i &= \prod_{k=1}^i t_i.
+\end{align}
 

### 2017-02-22 cpp/results/mutual_info.py
Mutual information calculation

@@ -1,0 +1,17 @@
+# A little script for mutual information calculations
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Name of the problem
+name = "groupdiff"
+results1 = np.loadtxt(name + "1.txt")
+results2 = np.loadtxt(name + "2.txt")
+results3 = np.loadtxt(name + "3.txt")
+
+L = min([len(results1), len(results2), len(results3)])
+I = results1[0:L] + results2[0:L] - results3[0:L]
+
+plt.plot(I)
+print("I = " + str(I.mean()) + " +- " + str(I.std()/np.sqrt(L)))
+plt.show()
+

### 2017-02-22 paper/ms.tex
Some NS talk

@@
 
-Nested Sampling \citep[NS][]{skilling2006nested}
+Nested Sampling \citep[NS][]{skilling2006nested} is an algorithm whose aim is
+to calculate the marginal likelihood
+\begin{align}
+Z &= \int \pi(\x) L(\x) \, d\x
+\end{align}
+where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
+and $L$ is the likelihood function. It does this evolving a population of
+$N$ particles in the parameter space.
+
+Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
+some threshold $\ell$,
+\begin{align}
+X(\ell) &= \int \pi(\x) \mathds{1}\left(L(\x) > \ell\right) \, d\x,
+\end{align}
+
+The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
+particle with the lowest likelihood.
 

### 2017-02-23 cpp/mutual_info.py
Moved .py file

@@ -4,14 +4,14 @@ import numpy as np
 
 # Name of the problem
 name = "groupdiff"
-results1 = np.loadtxt(name + "1.txt")
-results2 = np.loadtxt(name + "2.txt")
-results3 = np.loadtxt(name + "3.txt")
+results1 = np.loadtxt("results/" + name + "1.txt")
+results2 = np.loadtxt("results/" + name + "2.txt")
+results3 = np.loadtxt("results/" + name + "3.txt")
 
 L = min([len(results1), len(results2), len(results3)])
 I = results1[0:L] + results2[0:L] - results3[0:L]
 
 plt.plot(I)
-print("I = " + str(I.mean()) + " +- " + str(I.std()/np.sqrt(L)))
+print("I = " + str(I.mean()) + " +- " + str(I.std()/np.sqrt(L)) + " nats.")
 plt.show()
 

### 2017-02-23 cpp/mutual_info.py
Horizontal line

@@ -12,6 +12,7 @@ L = min([len(results1), len(results2), len(results3)])
 I = results1[0:L] + results2[0:L] - results3[0:L]
 
 plt.plot(I)
+plt.axhline(0.0, linestyle="--", alpha=0.3)
 print("I = " + str(I.mean()) + " +- " + str(I.std()/np.sqrt(L)) + " nats.")
 plt.show()
 

### 2017-02-24 paper/ms.tex
Priors for group diff example

@@
 
-\section{Example 2: Measuring the period of an oscillating signal}
+\section{Example 2: Measuring the difference between two group means}
+
+A basic problem in statistics is the inference of the difference between
+two `group mean' parameters $\mu_1$ and $\mu_2$ from data in the form of
+samples $\boldsymbol{x} = \{x_1, ..., x_m\}$ and
+$\boldsymbol{y} = \{y_1, ..., y_n\}$
+The standard frequentist method for this
+problem is the two-sample $t$-test, and Bayesian alternatives are
+straightforward to specify.
+
+The priors are
+\begin{align}
+         \mu_1  &\sim \textnormal{Normal}(0,1)     \\
+  \ln \sigma_1  &\sim \textnormal{Normal}(0,1)     \\
+         d_\mu  &\sim \textnormal{Normal}(0,1)     \\
+ d_{\ln \sigma} &\sim \textnormal{Normal}(0,0.1^2) \\
+         \mu_2  &\leftarrow \mu_1 + d_\mu          \\
+  \ln \sigma_2  &\leftarrow \ln \sigma_1 + d_{\ln \sigma}
+\end{align}
+and the conditional prior for the data given the parameters is
+\begin{align}
+x_i &\sim \textnormal{Normal}\left(\mu_1, \sigma_1^2\right) \\
+y_i &\sim \textnormal{Normal}\left(\mu_2, \sigma_2^2\right)
+\end{align}
+
+\section{Example 3: Measuring the period of an oscillating signal}
 

### 2017-02-27 paper/figures/demo.py
More contours

@@ -6,6 +6,10 @@ import matplotlib.pyplot as plt
 Generate some figures and perhaps a movie.
 """
 
+plt.rc("font", size=14, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
+
 def density(x, y):
     """
     A 2D joint density used for the demo.
@@ -36,16 +40,20 @@ if __name__ == "__main__":
     f = density(x, y)
 
     # Generate from the pdf
-    samples = generate(100)
+    samples = generate(1)
+
+    # Euclidean distance from the sample
+    distance = np.sqrt((x - samples["x"][0])**2 + \
+                       (y - samples["y"][0])**2)
 
     # Plot the pdf and the samples
     plt.imshow(density(x, y), cmap="Blues", interpolation="nearest",
                extent=[x.min(), x.max(), y.min(), y.max()])
-    plt.xlabel("$x$", fontsize=20)
-    plt.ylabel("$y$", fontsize=20)
-    plt.hold(True)
-    plt.contour(x, y, density(x, y), 50, colors=["k"], alpha=0.1)
-    plt.plot(samples["x"], samples["y"], "r.", alpha=0.3, markersize=10)
+    plt.xlabel("$x$")
+    plt.ylabel("$y$")
+    plt.contour(x, y, density(x, y), 20, colors=["b"], alpha=0.1)
+    plt.plot(samples["x"], samples["y"], "r.", alpha=0.5, markersize=20)
+    plt.contour(x, y, distance, 20, colors=["r"], alpha=0.1)
     plt.axis([x.min(), x.max(), y.min(), y.max()])
     plt.savefig("figure.png")
     plt.show()

### 2017-02-27 paper/figures/demo.py
PDF figure

@@ -55,6 +55,6 @@ if __name__ == "__main__":
     plt.plot(samples["x"], samples["y"], "r.", alpha=0.5, markersize=20)
     plt.contour(x, y, distance, 20, colors=["r"], alpha=0.1)
     plt.axis([x.min(), x.max(), y.min(), y.max()])
-    plt.savefig("figure.png")
+    plt.savefig("figure.pdf", bbox_inches="tight")
     plt.show()
 

### 2017-02-27 paper/ms.tex
Included the figure

@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.5\linewidth]{figures/algorithm.pdf}
+\caption{To evaluate the log-probability (or density) of the blue
+probability distribution at the red point, Nested Sampling can be used,
+with the blue distribution plays the role of the ``prior'' in NS, and the
+Euclidian distance from the red point (illustrated with red contours)
+is the ``likelihood''.\label{fig:algorithm}}
+\end{figure}
+
 \subsection{Discrete vs. differential entropies}

### 2017-02-27 paper/figures/algorithm.py

@@ -55,6 +55,6 @@ if __name__ == "__main__":
     plt.plot(samples["x"], samples["y"], "r.", alpha=0.5, markersize=20)
     plt.contour(x, y, distance, 20, colors=["r"], alpha=0.1)
     plt.axis([x.min(), x.max(), y.min(), y.max()])
-    plt.savefig("figure.pdf", bbox_inches="tight")
+    plt.savefig("algorithm.pdf", bbox_inches="tight")
     plt.show()
 

### 2017-02-27 paper/ms.tex
Good writing

--- 2017-02-22 paper/ms.tex
@@
 
+In physics and astronomy, it is common to measure an oscillating signal
+at a set of times $\{t_1, ..., t_n\}$, in order to infer the amplitude,
+period, and phase of the signal.
+Here, I demonstrate the algorithm on a toy version
+of Bayesian experimental design: at what times should the signal be
+observed in order to obtain as much information as possible about the
+period? To answer this, we need to be able to calculate the mutual
+information between the unknown period and the data.
+
+I consider two possible observing strategies, both involving
+$n=101$ measurements. The `even' strategy has us observe at times
+\begin{align}
+t_i = \frac{i-1}{n-1}
+\end{align}
+for $i \in \{1, 2, ..., 101\}$, that is, the observation times are
+$0, 0.01, 0.02, ..., 0.99, 1$ ---
+evenly spaced from $t=0$ to $t=1$, including the endpoints.
+The second, `uneven' observing strategy schedules the observations
+according to
+\begin{align}
+t_i = \left(\frac{i - \frac{1}{2}}{n}\right)^3
+\end{align}
+which schedules observations close together initially, and the
+spacing increases with time.
+
+\subsection{Assumptions}
+
+A purely sinusoidal signal has the form
+\begin{align}
+y(t) &= A \sin \left(\frac{2\pi t}{T} + \phi\right)
+\end{align}
+where $A$ is the amplitude, $T$ is the period, and $\phi$ is the phase.
+Throughout this section, I parameterise the period by its logarithm,
+$\tau = \log_{10} T$. I assumed the following priors:
+\begin{align}
+\ln A   &\sim \textnormal{Normal}\left(0, 0.1^2\right)  \\
+\tau    &\sim \textnormal{Uniform}(-1, 0)  \\
+\phi    &\sim \textnormal{Uniform}(0, 2\pi)
+\end{align}
+and the following conditional prior for the data:
+\begin{align}
+d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
+\end{align}
+That is, the data is just the signal $y(t)$ observed at particular times
+$\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
+
 The multimodality of the posterior here raises an interesting issue. Is
@@
 {\todo or something like that.}
-
-The signal has the form
-\begin{align}
-y(t) &= A \sin \left(\frac{2\pi t}{T} + \phi\right)
-\end{align}
-where $A$ is the amplitude, $T$ is the period, and $\phi$ is the phase.
-I assumed the following priors:
-\begin{align}
-\log_e{A}   &\sim \textnormal{Normal}\left(0, 0.1^2\right)  \\
-\log_{10} T &\sim \textnormal{Uniform}(-1, 0)  \\
-\phi        &\sim \textnormal{Uniform}(0, 2\pi)
-\end{align}
-and the following conditional prior for the data:
-\begin{align}
-d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
-\end{align}
-That is, the data is just the signal $y(t)$ observed at particular times
-$\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
 

### 2017-02-27 paper/ms.tex
More sinusoidal writing

@@
 $\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
-
-The multimodality of the posterior here raises an interesting issue. Is
-the question we really want answered ``what is the value of $T$
-{\bf precisely}?'', to which the mutual information relates?
-Most practicing scientists would not feel particularly informed to learn
-that the vast majority of possibilities had been ruled out, if the
-posterior still consisted of several widely separated modes!
-Perhaps, in some applications, a more appropriate question is
-``what is the value of $T$ to within $\pm$ 10\%''
-{\todo or something like that.}
+The amplitude of the sinusoid is very likely to be around 10 times the
+noise level, and the period is between 0.1 and 1 times the observation
+interval.
 
@@
 \end{align}
-
-{\query The multimodality of the posterior in this problem (sometimes) might
-affect mixing sufficiently that the NS gets the incorrect answer. Be
-careful! And maybe use an example where this is unlikely, or double
-check it using brute force.}
 
@@
 H'_{\d}       &= 106.8 \pm 1.1   \\
-H'_{\tau, \d} &= 97.0 \pm 1.0
+H'_{\tau, \d} &= 97.0 \pm 1.0    \\
 I_{\tau, \d} &= 10.9 \pm 0.4
@@
 are not small enough to be certain of this.
+
+\subsection{Multimodality and precisional questions}
+
+The multimodality of the posterior here raises an interesting issue. Is
+the question we really want answered ``what is the value of $T$
+{\bf precisely}?'', to which the mutual information relates?
+Most practicing scientists would not feel particularly informed to learn
+that the vast majority of possibilities had been ruled out, if the
+posterior still consisted of several widely separated modes!
+Perhaps, in some applications, a more appropriate question is
+``what is the value of $T$ to within $\pm$ 10\%''
+{\todo or something like that.}
+{\query The multimodality of the posterior in this problem (sometimes) might
+affect mixing sufficiently that the NS gets the incorrect answer. Be
+careful! And maybe use an example where this is unlikely, or double
+check it using brute force.}
 

### 2017-02-27 paper/figures/algorithm.py
Separated rc()

@@ -1,14 +1,14 @@
+from plot_utils import rc
 import numpy as np
 import numpy.random as rng
 import matplotlib.pyplot as plt
 
+# Set default fonts etc
+rc()
+
 """
 Generate some figures and perhaps a movie.
 """
-
-plt.rc("font", size=14, family="serif", serif="Computer Sans")
-plt.rc("text", usetex=True)
-
 
 def density(x, y):
     """

### 2017-02-27 paper/figures/plot_utils.py

@@ -1,0 +1,9 @@
+import matplotlib.pyplot as plt
+
+def rc():
+    """
+    Sets plotting defaults.
+    """
+    plt.rc("font", size=14, family="serif", serif="Computer Sans")
+    plt.rc("text", usetex=True)
+

### 2017-02-27 paper/figures/algorithm.py
Fontsize 16 on the axis labels

@@ -49,8 +49,8 @@ if __name__ == "__main__":
     # Plot the pdf and the samples
     plt.imshow(density(x, y), cmap="Blues", interpolation="nearest",
                extent=[x.min(), x.max(), y.min(), y.max()])
-    plt.xlabel("$x$")
-    plt.ylabel("$y$")
+    plt.xlabel("$x$", fontsize=16)
+    plt.ylabel("$y$", fontsize=16)
     plt.contour(x, y, density(x, y), 20, colors=["b"], alpha=0.1)
     plt.plot(samples["x"], samples["y"], "r.", alpha=0.5, markersize=20)
     plt.contour(x, y, distance, 20, colors=["r"], alpha=0.1)

### 2017-02-27 paper/figures/sinewave.py
gr8 sinewave plot

@@ -1,0 +1,45 @@
+from plot_utils import rc
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+
+# Set rng seed
+rng.seed(0)
+
+# Set default fonts etc
+rc()
+
+# True signal
+def signal(t):
+    return np.sin(2 * np.pi * t / 0.3)
+
+# Number of data points
+n = 101
+
+# i-values as defined in the paper (i.e., starting from 1)
+i = np.arange(0, n) + 1
+
+# Two observing strategies
+t_even   = (i - 1) / (n - 1)
+t_uneven = ((i - 0.5) / n)**3
+
+# 'Continuous' time
+t = np.linspace(0.0, 1.0, 1001)
+
+# Data
+y_even   = signal(t_even) + 0.1*rng.randn(n)
+y_uneven = signal(t_uneven) + 0.1*rng.randn(n)
+y_smooth = signal(t)
+
+plt.plot(t, y_smooth, "k", label="True signal", alpha=0.5)
+plt.errorbar(t_even, y_even, color="orange", fmt="o", yerr=0.1,
+             label="Even data", alpha=0.3)
+plt.errorbar(t_uneven, y_uneven, color="green", fmt="o", yerr=0.1,
+             label="Uneven data", alpha=0.3)
+plt.xlabel("$t$", fontsize=16)
+plt.ylabel("$y$", fontsize=16)
+plt.ylim([-1.3, 2.4])
+plt.legend(loc="upper left")
+plt.savefig("sinewave.pdf", bbox_inches="tight")
+plt.show()
+

### 2017-02-27 paper/ms.tex
Include the sinewave figure

@@
 
+An example signal observed with the even and uneven observing schedules
+is shown in Figure~\ref{fig:sinewave}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\linewidth]{figures/sinewave.pdf}
+\caption{A signal with true parameters $A=1$, $\tau=-0.5$, and
+$\phi=0$, observed with noise standard deviation 0.1 with the
+even (gold points) and uneven (green points) observing strategies.
+\label{fig:sinewave}}
+\end{figure}
+
 \subsection{Multimodality and precisional questions}

### 2017-02-27 paper/figures/sinewave.py

@@ -11,7 +11,7 @@ rc()
 
 # True signal
 def signal(t):
-    return np.sin(2 * np.pi * t / 0.3)
+    return np.sin(2 * np.pi * t / (10.0 ** -0.5))
 
 # Number of data points
 n = 101

### 2017-02-27 paper/ms.tex
More references

@@
 \newcommand{\n}{\boldsymbol{\eta}}
+
+\newcommand{\apj}{The Astrophysical Journal}
+\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
 
@@
 The amplitude of the sinusoid is very likely to be around 10 times the
-noise level, and the period is between 0.1 and 1 times the observation
-interval.
+noise level, and the period is between 0.1 and 1 times the duration of
+the data.
 
@@
 \subsection{Multimodality and precisional questions}
+
+When inferring the period of an oscillating signal, multimodal posterior
+distributions for the period are common \citep{gregoryTrimodal, exoplanet}
 

### 2017-02-27 paper/ms.tex
Provisional results

@@
 I consider two possible observing strategies, both involving
-$n=101$ measurements. The `even' strategy has us observe at times
+$n=101$ measurements. The `even' strategy has observations at times
 \begin{align}
@@
 \end{align}
-which schedules observations close together initially, and the
-spacing increases with time.
+which schedules observations close together initially, and further apart
+as time goes on.
 
@@
 of $\tau$, $\d$, and $(\tau, \d)$.
-These results were computed using a tolerance of 0.001 for every quantity.
-The results were
+These results were computed using a tolerance of 0.001 for each of the three
+runs. The {\todo interim} results are:
 \begin{align}
-H_{\tau}    &= 7.337 \pm 0.028 \\
-H_{\d}      &= 903.8 \pm 1.1   \\
-H_{\tau,\d} &= 900.8 \pm 1.1      % Interim result only
+H_{\tau}    &= 7.307 \pm 0.027 \\
+H_{\d}      &= 905.5 \pm 5.6   \\
+H_{\tau,\d} &= 883.6 \pm 5.2
 \end{align}
 
-To convert these to differential entropies, we need to add the appropriate
-log-volume to each. These are $-6.215$, $-790.32$, and
-$-796.53$ respectively, so the differential entropies are:
+To conver these to differential entropies, I applied the log-volume
+correction factors for 1, 101, and 102 dimensions, which are
+$-6.215$, $-790.32$, and $-798.62$ respectively. This gives
+the following {\todo interim} values for the differential entropies:
 \begin{align}
-H'_{\tau}     &= 1.122 \pm 0.028 \\
-H'_{\d}       &= 115.5 \pm 1.9   \\
-H'_{\tau, \d} &= 105.7 \pm 1.7
+H'_{\tau}     &= 1.092 \pm 0.027 \label{eqn:h_tau}\\
+H'_{\d}       &= 115.2 \pm 5.6   \\
+H'_{\tau, \d} &=  93.3 \pm 5.2   \label{eqn:h_joint}
 \end{align}
@@
 uncertainties in each of these three quantities is quite large, and would lead
-to a large uncertainty on $I_{\tau, \d}$ if computed naively. To reduce this
+to a large uncertainty on $I_{\tau, \d}$ if computed from the results
+in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}. To reduce this
 uncertainty, we computed the three entropies using a common random number
@@
 {\em common random numbers} technique of variance reduction.
-The resulting mutual information is
-$I_{\tau, \d} = 10.35 \pm 0.61$ nats.
+The resulting {\todo provisional} mutual information is
+$I_{\tau, \d} = 31.7 \pm 3.8$ nats.
 

### 2017-02-27 paper/ms.tex
talking about the cost

@@
 \newcommand{\x}{\boldsymbol{\theta}}
+\newcommand{\bphi}{\boldsymbol{\phi}}
+\newcommand{\boldeta}{\boldsymbol{\eta}}
 \newcommand{\n}{\boldsymbol{\eta}}
@@
 
-The computational resources needed to compute these quantities are quite large,
-though the algorithm can be parallelised trivially. However, the problem
-being solved is not trivial. The mutual information of $\tau$ and
-$\d$ is
-\begin{align}
-I_{\tau, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
-\end{align}
-However, $\x$ is one of several parameters. The others, $A$, and $\phi$,
-were implicitly integrated out. Denoting the nuisance parameters by
-$\n$, the mutual information that has been calculated is
-\begin{align}
-I_{\x, \d} &= \int p(\x, \d)
-  \log\frac{\int p(\x, \n, \d) \, d\n }
-           {p(\x) \int p(\x', \n)p(\d | \x', \n) \, d\x' \, d\n}
-  \, d\x \, d\d.
-\end{align}
-It should not be considered surprising that this is a challenging task.
+\subsection{Uneven observations}
 
-\subsection{Uneven observations}
 The results for the uneven observing schedule are as follows:
@@
 \end{figure}
+
+\subsection{Computational cost}
+
+The computational resources needed to compute these quantities was quite large
+--- three lots of 1,000 repetitions of Nested Sampling, in all
+(though the algorithm can be parallelised trivially).
+However, it is worth reflecting on the difficulty of the calculation that
+has been done.
+
+In a Bayesian inference problem with parameters $\x$ and data $\d$, the
+mutual information is
+\begin{align}
+I_{\x, \d} &= \iint p(\x, \d)
+                        \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
+                        \, d\x \, d\d
+\end{align}
+and this measures the dependence between $\x$ and $\d$. A Monte Carlo
+strategy to evaluate this is to sample from $p(\x, \d)$ and average
+the value of the logarithm, which involves computing the log-evidence
+\begin{align}
+\ln p(\d) = \ln\left[\int p(\x)p(\d | \x) \, d\x\right]
+\end{align}
+for each possible data set. Computing $p(\d)$ for even a single
+data set has historically been considered difficult.
+
+However, if we want the mutual information between the data and
+{\em a subset of the parameters} (i.e., there are nuisance parameters
+we don't care about), things become even more tricky. Suppose we
+partition the parameters $\x$ into important parameters
+$\bphi$ and nuisance parameters $\boldeta$, such that we want to calculate
+the mutual information $I_{\bphi, \d}$. This is given by
+\begin{align}
+I_{\bphi, \d} &= \iint p(\bphi, \d)
+                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
+                        \, d\bphi \, d\d
+\end{align}
 

### 2017-02-28 paper/ms.tex
More writing

@@
 
-\documentclass[entropy,article,submit,oneauthor,pdftex,10pt,a4paper]{mdpi}
+\documentclass[entropy,article,accept,oneauthor,pdftex,10pt,a4paper]{mdpi}
 %--------------------
@@
 \end{align}
-which is the starting point for inference \citep{caticha}.
+which is the starting point for inference \citep{caticha2008lectures}.
 
-The entropy of this joint prior describes the degree to which the question
-``what is the value of the pair $(\x, \d)$?'' remains unanswered.
+The entropy of the prior $p(\x)$ describes the degree to which the question
+``what is the value of $\x$, precisely?'', while the
+entropy of the joint prior $p(\x, \d)$
+describes the degree to which the question
+``what is the value of the pair $(\x, \d)$, precisely?'' remains unanswered.
 
@@
 \end{align}
-For present purposes, it is more conveniently written as
+It can also be written in terms of the prior, likelihood, and evidence,
+as
 \begin{align}
-I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
-              \log\left[\frac{p(\d | \x)}{p(\d)}\right].
+I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x)p(\d | \x)
+              \left[\log p(\d | \x) - \log p(\d)\right],
 \end{align}
+i.e., the mutual information is the prior expected value of the
+log likelihood minus the log evidence.
 As with the conditional entropy, we see that the difficulty appears in the
@@
                         \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
-                        \, d\x \, d\d
+                        \, d\x \, d\d \label{eqn:mutual_info2}
 \end{align}
-and this measures the dependence between $\x$ and $\d$. A Monte Carlo
+and this measures the dependence between $\x$ and $\d$. A
+Monte Carlo
 strategy to evaluate this is to sample from $p(\x, \d)$ and average
@@
 \end{align}
-for each possible data set. Computing $p(\d)$ for even a single
+for each possible data set.
+Computing $p(\d)$ for even a single
 data set has historically been considered difficult.
@@
                         \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
-                        \, d\bphi \, d\d
+                        \, d\bphi \, d\d \label{eqn:mutual_info3}
+\end{align}
+which is equivalent to Equation~\ref{eqn:mutual_info2} but with
+$\x$ replaced throughout by $\bphi$. However, this is an even more
+difficult calculation than before, as $p(\d | \bphi)$, the marginal likelihood
+function with the nuisance parameters integrated out, is typically
+unavailable in closed form. If we were to marginalise out the nuisance
+parameters $\boldeta$ explicitly, this would give us
+Equation~\ref{eqn:mutual_info3} with every probability distribution
+written as an explicit integral over $\boldeta$:
+
+\begin{align}
+I_{\bphi, \d} &= \iint
+                   \int p(\bphi, \boldeta)p(\d | \phi, \boldeta) \, d\boldeta
+                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
+                        \, d\bphi \, d\d \label{eqn:mutual_info3}
 \end{align}

### 2017-02-28 paper/ms.tex
Bragging about the integral
Call data d

@@
 Consider the basic problem of inferring a quantity $\mu$ from
-100 observations $\x = \{x_1, ..., x_{100}\}$ whose probability distribution
-(conditional on $\mu$) is
+100 observations $\d = \{d_1, ..., d_{100}\}$ whose
+probability distribution (conditional on $\mu$) is
 \begin{align}
-p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
+p(d_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
 \end{align}
@@
 \begin{align}
-\mu | \x &\sim \textnormal{Normal}\left(
-                                       \frac{1}{100}\sum_{i=1}^{100} x_i,
+\mu | \d &\sim \textnormal{Normal}\left(
+                                       \frac{1}{100}\sum_{i=1}^{100} d_i,
                                        \left[\frac{10}{\sqrt{10001}}\right]^2
@@
 \end{align}
+The true value of $H_{\d}$, calculated analytically, is $146.499$.
 
@@
 The log-volume of a 100-dimensional ball of radius $10^{-3}$ is
--782.017. Therefore, the measured differential entropy of
-$p(\d)$ is $928.80 - 782.017 = 146.79 \pm 0.38$, consistent with the
-true value of $146.499$.
-
+-782.017. Therefore, the measured differential en
 \section{Example 2: Measuring the period of an oscillating signal}
@@
 I_{\bphi, \d} &= \iint
-                   \int p(\bphi, \boldeta)p(\d | \phi, \boldeta) \, d\boldeta
-                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
-                        \, d\bphi \, d\d \label{eqn:mutual_info3}
+  \left(\int p(\bphi, \boldeta)p(\d | \bphi, \boldeta) \, d\boldeta\right)
+                        \ln \left[
+  \frac{\int p(\bphi, \boldeta)p(\d | \bphi, \boldeta) \, d\boldeta}
+{p(\bphi)\int p(\bphi, \boldeta)p(\d | \bphi, \boldeta) \, d\bphi \, \d\boldeta}\right]
+                        \, d\bphi \, d\d
 \end{align}
+It should not be surprising that this is costly.
 

### 2017-02-28 paper/ms.tex
Preamble stuff

@@
 \section{Introduction}
-\noindent
+
 If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
@@
 where the sum is over all of the possible values of $x$ under consideration.
+
+The entropy quantifies the degree to which the issue
+``what is the value of $x$, precisely?'' remains unresolved. For example,
+if there is only one possible value which has probability 1, $H_x$ is
+zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
+to $\lim_{x \to 0^+} x\log x$, which is zero.}.
+If there are $N$ possibilities with equal probabilities $1/N$ each,
+then the entropy is $H_x = \log(N)$.
+
+\subsection{Notation and conventions}
+
 Throughout this paper, I use the compact
@@
 of the more cumbersome $P(X=x)$.
-This probability distribution is implicitly conditional on some prior
-information, which has been omitted from the notation.
-
-\citet{knuth} has shown that the entropy quantifies the degree to which the
-question ``which value of $x$ is true?'' remains unanswered by the prior
-information.
+Probability distribution is implicitly conditional on some prior
+information, which has been omitted from the notation unless necessary.
+All logarithms are written $\log$ and any base can be used, unless otherwise
+specified (for example by writing $\ln$ or $\log_{10}$). Any numerical
+values given for the value of specific entropies are in {\em nats}, i.e.,
+the natural logarithm was used.
 
@@
 \end{align}
-which is the starting point for inference \citep{caticha2008lectures}.
+which is the starting point for
+Bayesian inference \citep{caticha2008lectures}.
 

### 2017-03-03 paper/ms.tex
Notation
Final results for sinusoid (even)

@@
 the natural logarithm was used.
+
+I prefer to write the entropy of $x$ as $H_x$ rather than $H(x)$
+(or $H(X)$ as is often seen), because the latter can wrongly imply that
+the entropy is a property of the value of $x$ itself, rather than
+our probability distribution for it.
 
@@
 These results were computed using a tolerance of 0.001 for each of the three
-runs. The {\todo interim} results are:
+runs. The results are:
 \begin{align}
 H_{\tau}    &= 7.307 \pm 0.027 \\
-H_{\d}      &= 905.5 \pm 5.6   \\
-H_{\tau,\d} &= 883.6 \pm 5.2
+H_{\d}      &= 900.6 \pm 1.1   \\
+H_{\tau,\d} &= 882.2 \pm 1.1
 \end{align}
@@
 $-6.215$, $-790.32$, and $-798.62$ respectively. This gives
-the following {\todo interim} values for the differential entropies:
+the following values for the differential entropies:
 \begin{align}
 H'_{\tau}     &= 1.092 \pm 0.027 \label{eqn:h_tau}\\
-H'_{\d}       &= 115.2 \pm 5.6   \\
-H'_{\tau, \d} &=  93.3 \pm 5.2   \label{eqn:h_joint}
+H'_{\d}       &= 110.3 \pm 1.1   \\
+H'_{\tau, \d} &=  83.6 \pm 1.1   \label{eqn:h_joint}
 \end{align}
@@
 {\em common random numbers} technique of variance reduction.
-The resulting {\todo provisional} mutual information is
-$I_{\tau, \d} = 31.7 \pm 3.8$ nats.
+The resulting  mutual information is
+$I_{\tau, \d} = 27.85 \pm 0.83$ nats.
 

### 2017-03-05 paper/ms.tex
More results

@@
 noise level, and the period is between 0.1 and 1 times the duration of
-the data.
+the data. An example signal observed with the even and uneven observing schedules
+is shown in Figure~\ref{fig:sinewave}.
 
+\subsection{Results}
 Letting $\tau = \log_{10} T$, and treating
@@
 
-\subsection{Uneven observations}
-
-The results for the uneven observing schedule are as follows:
+The results for the uneven observing schedule were:
 \begin{align}
 H'_{\tau}     &= 1.122 \pm 0.028 \\
-H'_{\d}       &= 106.8 \pm 1.1   \\
-H'_{\tau, \d} &= 97.0 \pm 1.0    \\
-I_{\tau, \d} &= 10.9 \pm 0.4
+H'_{\d}       &= 895.5 \pm 1.5   \\
+H'_{\tau, \d} &= 874.7 \pm 1.5    \\
+I_{\tau, \d} &= 30.13 \pm 0.70
 \end{align}
-which appears to be higher than under the even observing schedule,
-though the uncertainties
-are not small enough to be certain of this.
 
-An example signal observed with the even and uneven observing schedules
-is shown in Figure~\ref{fig:sinewave}.
+The difference in mutual informations between the two observation
+schedules is $2.3 \pm 1.3$ nats, suggesting a slight advantage to the
+uneven observing schedule, in terms of the data's relevance to inference
+of the log-period of the signal.
+In a serious experimental design situation, the relevance is not the only
+consideration (or the answer to every experimental design problem would
+be to obtain more data without limit), but it is a very important one.
 

### 2017-03-05 paper/ms.tex
Moved a figure

@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\linewidth]{figures/sinewave.pdf}
+\caption{A signal with true parameters $A=1$, $\tau=-0.5$, and
+$\phi=0$, observed with noise standard deviation 0.1 with the
+even (gold points) and uneven (green points) observing strategies.
+\label{fig:sinewave}}
+\end{figure}
+
 \subsection{Results}
+
 Letting $\tau = \log_{10} T$, and treating
@@
 
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=0.7\linewidth]{figures/sinewave.pdf}
-\caption{A signal with true parameters $A=1$, $\tau=-0.5$, and
-$\phi=0$, observed with noise standard deviation 0.1 with the
-even (gold points) and uneven (green points) observing strategies.
-\label{fig:sinewave}}
-\end{figure}
-
 \subsection{Computational cost}
@@
 (though the algorithm can be parallelised trivially).
-However, it is worth reflecting on the difficulty of the calculation that
+However, it is worth reflecting on the complexity of the calculation that
 has been done.

### 2017-03-05 cpp/Rep.hpp
size_t -> unsigned int

@@ -57,7 +57,7 @@ class Rep
 
         // Output file handle, and some cache
         std::fstream fout;
-        std::vector<std::tuple<size_t, double>> cache;
+        std::vector<std::tuple<unsigned int, double>> cache;
 
         // Member function to compute all the distances
         void compute_distances();
@@ -176,7 +176,7 @@ void Rep<Particle>::iterate(bool generate_replacement)
 
     // Write its information to the output file.
     cache.push_back({iteration, distances[worst]});
-    size_t it; double d;
+    unsigned int it; double d;
     if(cache.size() >= 100)
     {
         for(size_t i=0; i<cache.size(); ++i)

### 2017-03-05 cpp/Rep.hpp
Explicitly construct tuple

@@ -175,7 +175,9 @@ void Rep<Particle>::iterate(bool generate_replacement)
             worst = i;
 
     // Write its information to the output file.
-    cache.push_back({iteration, distances[worst]});
+    cache.push_back(
+        std::tuple<unsigned int, double>{iteration, distances[worst]}
+                   );
     unsigned int it; double d;
     if(cache.size() >= 100)
     {

### 2017-03-10 paper/ms.tex
Updated results for sinusoid

@@
 \begin{align}
-H_{\tau}    &= 7.307 \pm 0.027 \\
-H_{\d}      &= 900.6 \pm 1.1   \\
-H_{\tau,\d} &= 882.2 \pm 1.1
+H_{\tau}    &= 6.243  \pm 0.024 \\
+H_{\d}      &= 712.43 \pm 0.35   \\
+H_{\tau,\d} &= 715.81 \pm 0.34
 \end{align}
 
-To conver these to differential entropies, I applied the log-volume
+To convert these to differential entropies, I applied the log-volume
 correction factors for 1, 101, and 102 dimensions, which are
@@
 \begin{align}
-H'_{\tau}     &= 1.092 \pm 0.027 \label{eqn:h_tau}\\
-H'_{\d}       &= 110.3 \pm 1.1   \\
-H'_{\tau, \d} &=  83.6 \pm 1.1   \label{eqn:h_joint}
+H'_{\tau}     &=  0.028 \pm 0.024 \label{eqn:h_tau}\\
+H'_{\d}       &= -77.89 \pm 0.35   \\
+H'_{\tau, \d} &= -82.81 \pm 0.34   \label{eqn:h_joint}
 \end{align}
@@
 The resulting  mutual information is
-$I_{\tau, \d} = 27.85 \pm 0.83$ nats.
+$I_{\tau, \d} = 4.96 \pm 0.38$ nats.
 

### 2017-03-10 paper/ms.tex
slight tweaks

@@
 % Full title of the paper (Capitalized)
-\Title{Calculating Entropies With Nested Sampling}
+\Title{Computing Entropies With Nested Sampling}
 
@@
 zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
-to $\lim_{x \to 0^+} x\log x$, which is zero.}.
+to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
 If there are $N$ possibilities with equal probabilities $1/N$ each,
@@
 of the more cumbersome $P(X=x)$.
-Probability distribution is implicitly conditional on some prior
-information, which has been omitted from the notation unless necessary.
+Probability distributions are implicitly conditional on some prior
+information, which is omitted from the notation unless necessary.
 All logarithms are written $\log$ and any base can be used, unless otherwise
@@
 The entropy of the prior $p(\x)$ describes the degree to which the question
-``what is the value of $\x$, precisely?'', while the
+``what is the value of $\x$, precisely?'' remains unanswered, while the
 entropy of the joint prior $p(\x, \d)$

### 2017-03-13 cpp/examples/AllExamples.h
A Pareto example for ABC

@@ -3,6 +3,7 @@
 
 #include "GroupDiff.h"
 #include "Normal.h"
+#include "Pareto.h"
 #include "Sinusoid.h"
 
 #endif

### 2017-03-13 cpp/examples/Pareto.cpp

@@ -1,0 +1,114 @@
+#include "Pareto.h"
+#include "Utils.h"
+
+#include <algorithm>
+
+namespace InfoNest
+{
+
+Pareto::Pareto()
+:xs(1000)
+{
+
+}
+
+void Pareto::generate(RNG& rng)
+{
+    alpha = 0.01 + 4.99*rng.rand();
+
+    for(double& x: xs)
+        x = x_min * pow(rng.rand(), -1.0 / alpha);
+}
+
+double Pareto::perturb(RNG& rng)
+{
+    double logH = 0.0;
+
+    int proposal_type = rng.rand_int(3);
+
+    if(proposal_type == 0)
+    {
+        double old_alpha = alpha;
+
+        alpha += 4.99*rng.randh();
+        wrap(alpha, 0.01, 5.0);
+
+        // Transform data to go with new alpha value
+        for(double& x: xs)
+        {
+            x = pow(x / x_min, -old_alpha);
+            x = x_min * pow(x, -1.0 / alpha);
+        }
+    }
+    else if(proposal_type == 1)
+    {
+        int which = rng.rand_int(xs.size());
+        xs[which] = pow(xs[which] / x_min, -alpha);
+        xs[which] += rng.randh();
+        wrap(xs[which], 0.0, 1.0);
+        xs[which] = x_min * pow(xs[which], -1.0 / alpha);
+    }
+    else
+    {
+        int reps = (int)pow((double)xs.size(), rng.rand());
+        int which;
+        for(int i=0; i<reps; ++i)
+        {
+            which = rng.rand_int(xs.size());
+            xs[which] = x_min * pow(rng.rand(), -1.0 / alpha);
+        }
+    }
+
+
+    return logH;
+}
+
+void Pareto::print(std::ostream& out)
+{
+    out<<alpha<<' ';
+    for(double x: xs)
+        out<<x<<' ';
+}
+
+double Pareto::parameter_distance(const Pareto& pareto1,
+                                  const Pareto& pareto2)
+{
+    double tot1 = 0.0;
+    for(size_t i=0; i<900; ++i)
+        tot1 += pareto1.xs[i];
+
+    double tot2 = 0.0;
+    for(size_t i=0; i<900; ++i)
+        tot2 += pareto2.xs[i];
+
+    return std::abs(log(tot1/900) - log(tot2/900));
+}
+
+double Pareto::data_distance(const Pareto& pareto1,
[…]

### 2017-03-13 cpp/examples/Pareto.h

@@ -1,0 +1,49 @@
+#ifndef InfoNest_Pareto
+#define InfoNest_Pareto
+
+// Includes
+#include <ostream>
+#include "../RNG.h"
+
+namespace InfoNest
+{
+
+class Pareto
+{
+    private:
+        // Lower limit
+        static constexpr double x_min = 1.0;
+
+        // Slope
+        double alpha;
+
+        // Data
+        std::vector<double> xs;
+
+    public:
+        // Do-nothing constructor
+        Pareto();
+
+        // Generate from the distribution
+        void generate(RNG& rng);
+
+        // Metropolis proposal
+        double perturb(RNG& rng);
+
+        // Printing to stream
+        void print(std::ostream& out);
+
+        // Available distance functions
+        static double parameter_distance(const Pareto& pareto1,
+                                         const Pareto& pareto2);
+        static double data_distance(const Pareto& pareto1,
+                                    const Pareto& pareto2);
+        static double joint_distance(const Pareto& pareto1,
+                                     const Pareto& pareto2);
+
+};
+
+} // namespace InfoNest
+
+#endif
+

### 2017-03-13 cpp/tests/pareto.cpp

@@ -1,0 +1,10 @@
+#include "SamplingTest.hpp"
+#include "../examples/Pareto.h"
+
+int main()
+{
+    sampling_test<InfoNest::Pareto>();
+
+    return 0;
+}
+

### 2017-03-13 paper/ms.tex
Introduction writing

@@
 then the entropy is $H_x = \log(N)$.
+If the space of possible $x$-values is continuous, the differential
+entropy
+\begin{align}
+H_{x} &= -\int p(x) \log p(x) \, dx
+\end{align}
+quantifies uncertainty by generalising the log-volume
+of the plausible region, where volume is defined with respect to $dx$.
 
@@
 Throughout this paper, I use the compact
-`overloaded' notation for probability distributions, writing $p(x)$ instead
-of the more cumbersome $P(X=x)$.
+`overloaded' notation for probability distributions favoured by
+Bayesians, writing $p(x)$ (for either a probability mass function
+or a probability density function) instead
+of the more correct but cumbersome $P(X=x)$ or $f_X(x)$.
 Probability distributions are implicitly conditional on some prior
@@
 
+In practical applications of Bayesian inference, the following operations
+are usually feasible, and are assumed to be possible at low computational
+cost:
+\begin{enumerate}
+\item Points can be generated from the prior $p(\x)$;
+\item Simulated datasets can be generated from $p(\d | \x)$ for a given value of $\x$;
+\item The likelihood, $p(\d | \x)$ can be evaluated cheaply for any $\d$ and $\x$.
+\end{enumerate}
+
+\subsection{The relevance of data}
+
 The entropy of the prior $p(\x)$ describes the degree to which the question
@@
 
-\subsection{The relevance of data}
 The degree to which the question ``what is the value of $\x$?'' is
@@
 all possible datasets which might be observed.
+
 This is difficult to compute because the expression

### 2017-03-15 paper/ms.tex
Included final results on sinusoid problem

@@
 
-The results for the uneven observing schedule were:
+The results for the uneven observing schedule, again in terms of log
+probabilities, are:
 \begin{align}
-H'_{\tau}     &= 1.122 \pm 0.028 \\
-H'_{\d}       &= 895.5 \pm 1.5   \\
-H'_{\tau, \d} &= 874.7 \pm 1.5    \\
-I_{\tau, \d} &= 30.13 \pm 0.70
+H_{\tau}     &= 6.243  \pm 0.024 \\
+H_{\d}       &= 712.77 \pm 0.35   \\
+H_{\tau, \d} &= 715.84 \pm 0.35    \\
+I_{\tau, \d} &= 5.21 \pm 0.40
 \end{align}
@@
 The difference in mutual informations between the two observation
-schedules is $2.3 \pm 1.3$ nats, suggesting a slight advantage to the
+schedules, assuming a uniform prior on this difference,
+is just $0.25 \pm 0.55$ nats, suggesting (extremely weakly)
+a slight advantage to the
 uneven observing schedule, in terms of the data's relevance to inference
-of the log-period of the signal.
+of the log-period of the signal. The situation may be different if we
+had allowed for shorter periods to be possible, as irregular observing
+schedules are known to reduce aliasing and ambiguity in the inferred period.
+When inferring the period of an oscillating signal, multimodal posterior
+distributions for the period are common \citep{gregoryTrimodal, exoplanet}.
+The multimodality of the posterior here raises an interesting issue. Is
+the question we really want answered ``what is the value of $T$
+{\bf precisely}?'', to which the mutual information relates?
+Most practicing scientists would not feel particularly informed to learn
+that the vast majority of possibilities had been ruled out, if the
+posterior still consisted of several widely separated modes!
+Perhaps, in some applications, a more appropriate question is
+``what is the value of $T$ to within $\pm$ 10\%'', or something along these
+lines.
+
 In a serious experimental design situation, the relevance is not the only
@@
 
-The computational resources needed to compute these quantities was quite large
---- three lots of 1,000 repetitions of Nested Sampling, in all
-(though the algorithm can be parallelised trivially).
+The computational resources needed to compute these quantities was quite large,
+as Nested Sampling appears in the inner loop of the algorithm.
 However, it is worth reflecting on the complexity of the calculation that
@@
 It should not be surprising that this is costly.
-
-\subsection{Multimodality and precisional questions}
-
-When inferring the period of an oscillating signal, multimodal posterior
-distributions for the period are common \citep{gregoryTrimodal, exoplanet}
-
-The multimodality of the posterior here raises an interesting issue. Is
-the question we really want answered ``what is the value of $T$
-{\bf precisely}?'', to which the mutual information relates?
-Most practicing scientists would not feel particularly informed to learn
-that the vast majority of possibilities had been ruled out, if the
-posterior still consisted of several widely separated modes!
-Perhaps, in some applications, a more appropriate question is
-``what is the value of $T$ to within $\pm$ 10\%''
-{\todo or something like that.}
-{\query The multimodality of the posterior in this problem (sometimes) might
-affect mixing sufficiently that the NS gets the incorrect answer. Be
-careful! And maybe use an example where this is unlikely, or double
-check it using brute force.}
 

### 2017-03-15 paper/ms.tex
Algorithm placeholder

@@
 \usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{algorithmicx}
-\usepackage{algpseudocode}
 \usepackage{natbib}
@@
 
+\begin{algorithm}
+\caption{Calculate $y = x^n$}
+\begin{algorithmic} 
+\REQUIRE $n \geq 0 \vee x \neq 0$
+\ENSURE $y = x^n$
+\STATE $y \leftarrow 1$
+\IF{$n < 0$}
+\STATE $X \leftarrow 1 / x$
+\STATE $N \leftarrow -n$
+\ELSE
+\STATE $X \leftarrow x$
+\STATE $N \leftarrow n$
+\ENDIF
+\WHILE{$N \neq 0$}
+\IF{$N$ is even}
+\STATE $X \leftarrow X \times X$
+\STATE $N \leftarrow N / 2$
+\ELSE[$N$ is odd]
+\STATE $y \leftarrow y \times X$
+\STATE $N \leftarrow N - 1$
+\ENDIF
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
+
 \begin{figure}[!ht]
@@
 \conflictofinterests{Conflicts of Interest}
-The authors declare no conflicts of interest.
+The authors declare no conflicts of interJason Mamford Half of the money collected from the Northern Toll road is wasted in costs.
+It cost millions of dollars for one gantry, so how much for 50 gantries?
+How much for a million transponders?
+You see, nobody will tell us anything until it is too late.
+est.
 

### 2017-03-16 paper/ms.tex
Working on the algorithm
More work on the algorithm
Fix in appendix

@@
 \usepackage[utf8]{inputenc}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+\usepackage{algorithm, mathtools}
+\usepackage[noend]{algpseudocode}
 \usepackage{amsmath}
@@
 \end{align}
-contains the marginal likelihood $p(\d)$.
+contains the marginal likelihood or `evidence'
+\begin{align}
+p(\d) &= \int p(\x) p(\d | \x) \, d\x.
+\end{align}
 
@@
 \begin{algorithm}
-\caption{Calculate $y = x^n$}
 \begin{algorithmic}
-\REQUIRE $n \geq 0 \vee x \neq 0$
-\ENSURE $y = x^n$
-\STATE $y \leftarrow 1$
-\IF{$n < 0$}
-\STATE $X \leftarrow 1 / x$
-\STATE $N \leftarrow -n$
-\ELSE
-\STATE $X \leftarrow x$
-\STATE $N \leftarrow n$
-\ENDIF
-\WHILE{$N \neq 0$}
-\IF{$N$ is even}
-\STATE $X \leftarrow X \times X$
-\STATE $N \leftarrow N / 2$
-\ELSE[$N$ is odd]
-\STATE $y \leftarrow y \times X$
-\STATE $N \leftarrow N - 1$
-\ENDIF
-\ENDWHILE
+\State {\em Set the numerical parameters:}
+\State $N \in \{1, 2, 3, ... \}$
+            \Comment{the number of Nested Sampling particles to use}
+\State $\epsilon \geq 0$
+            \Comment{the tolerance} \\
+\hrulefill
+\State $\widehat{\mathbf{h}} \leftarrow [\,]$
+            \Comment{Initialise an empty list of results}
+
+\While{more iterations desired}
+    \State Generate $\x_{\rm ref}$ from $p(\x)$
+            \Comment{Generate a reference point}
+    \State $k \leftarrow 0$
+            \Comment{Initialise counter}
+  {\color{green}
+    \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
+            \Comment{Generate initial NS particles}
+    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
+            \Comment{Calculate distance of each particle from the reference
+                     point}
+    \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
+            \Comment{Find the worst particle (greatest distance)}
+%%    \While{$d_{\rm max} > \epsilon$}
+%%        \State Blah
+%%    \EndWhile
+
+  } % \color{green}
+\State More to come...
+\EndWhile
 \end{algorithmic}
+\caption{Calculates $H_{\x} = -\int p(\x) \log p(\x) \, d\x$
+         without evaluating $\log p(\x)$.}
 \end{algorithm}
@@
 \begin{itemize}
-\item $x \in \{1, 2, 3\}$, $x \notin \{1, 2, 3\}$,
-                or anything that implies one of these
-\item $x \in \{2, 3, 4\}$, $x \notin \{2, 3, 4\}$,
-                or anything that implies one of these ...
-\item ... $x \in \{999999998, 999999999, 1000000000\}$,
-            $x \notin \{999999998, 999999999, 1000000000\}$,
-                or anything that implies one of these
+\item $x \in \{1, 2, 3\}$ and anything that implies it
+\item $x \in \{2, 3, 4\}$ and anything that implies it
+\item $x \in \{3, 4, 5\}$ and anything that implies it
+\item and so on.
 \end{itemize}

### 2017-03-16 paper/ms.tex
Removed accidental paste

@@
             \Comment{Find the worst particle (greatest distance)}
+    \State $d_{\rm max} \leftarrow d_{i^*}$
+            \Comment{Find the greatest distance}
 %%    \While{$d_{\rm max} > \epsilon$}
@@
 \conflictofinterests{Conflicts of Interest}
-The authors declare no conflicts of interJason Mamford Half of the money collected from the Northern Toll road is wasted in costs.
-It cost millions of dollars for one gantry, so how much for 50 gantries?
-How much for a million transponders?
-You see, nobody will tell us anything until it is too late.
-est.
+The authors declare no conflicts of interest.
 

### 2017-03-16 cpp/main.cpp
more numbers

@@ -19,10 +19,10 @@ int main()
     InfoNest::RNG rng2(seed2);
 
     // Define run parameters
-    constexpr double depth         = 1200.0;
+    constexpr double depth         = 1000.0;
     constexpr size_t num_reps      = 1000;
     constexpr size_t num_particles = 10;
-    constexpr size_t mcmc_steps    = 10000;
+    constexpr size_t mcmc_steps    = 5000;
 
     // Do the run.
     InfoNest::execute<TheExample>(rng1, rng2, depth, num_reps, num_particles,

### 2017-03-17 cpp/Rep.hpp
Tiebreaers

@@ -9,6 +9,7 @@
 #include <stdlib.h>
 #include <vector>
 #include "RNG.h"
+#include "Utils.h"
 
 namespace InfoNest
 {
@@ -51,6 +52,9 @@ class Rep
 
         // Distances
         std::vector<double> distances;
+
+        // Tiebreakers for distances
+        std::vector<double> tiebreakers;
 
         // Iteration counter
         unsigned int iteration;
@@ -110,6 +114,7 @@ Rep<Particle>::Rep(size_t id,
 ,rng(rng)
 ,particles(num_particles)
 ,distances(num_particles)
+,tiebreakers(num_particles)
 ,iteration(0)
 {
     // Check that the input was at least partially sane
@@ -142,6 +147,10 @@ void Rep<Particle>::initialise(RNG& temp_rng)
     // Compute all the distances
     compute_distances();
 
+    // Assign tiebreakers
+    for(double& tb: tiebreakers)
+        tb = rng.rand();
+
     // Set iteration number
     iteration = 1;
 }
@@ -171,7 +180,9 @@ void Rep<Particle>::iterate(bool last)
     // Find the worst particle.
     int worst = 0;
     for(size_t i=1; i<num_particles; ++i)
-        if(distances[i] > distances[worst])
+        if(distances[i] > distances[worst] ||
+            (distances[i] == distances[worst] &&
+             tiebreakers[i] > tiebreakers[worst]))
             worst = i;
 
     // Write its information to the output file.
@@ -211,6 +222,7 @@ void Rep<Particle>::replace(int which)
 {
     // Distance threshold
     double threshold = distances[which];
+    double tb_threshold = tiebreakers[which];
 
     // If num_particles > 1, clone a survivor.
     if(num_particles > 1)
@@ -223,11 +235,13 @@ void Rep<Particle>::replace(int which)
 
         particles[which] = particles[copy];
         distances[which] = distances[copy];
+        tiebreakers[which] = tiebreakers[copy];
     }
 
     // Stuff needed for MCMC
     Particle proposal;
     double proposal_distance;
+    double proposal_tb;
     double logA, alpha;
 
     // Count acceptances
@@ -240,17 +254,25 @@ void Rep<Particle>::replace(int which)
         proposal = particles[which];
         logA = proposal.perturb(rng);
         proposal_distance = dist(reference_particle, proposal);
+        proposal_tb = tiebreakers[which] + rng.randh();
+        wrap(proposal_tb, 0.0, 1.0);
 
         // Acceptance probability
[…]

### 2017-03-17 paper/ms.tex
Random writing in appendix
Finished algorithm
Algorithm color

@@
 \While{more iterations desired}
+    \State $k \leftarrow 0$
+            \Comment{Initialise counter}
     \State Generate $\x_{\rm ref}$ from $p(\x)$
             \Comment{Generate a reference point}
-    \State $k \leftarrow 0$
-            \Comment{Initialise counter}
-  {\color{green}
+\color{green}
     \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
@@
             \Comment{Find the greatest distance}
-%%    \While{$d_{\rm max} > \epsilon$}
-%%        \State Blah
-%%    \EndWhile
-
-  } % \color{green}
-\State More to come...
+    \While{$d_{\rm max} > \epsilon$}
+        \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
+               $p\left(\x | (d(\x) < d_{\rm max})\right)$
+            \Comment{Replace worst particle}
+        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \x_{\rm ref})$
+            \Comment{Calculate distance of new particle from reference point}
+        \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
+                \Comment{Find the worst particle (greatest distance)}
+        \State $d_{\rm max} \leftarrow d_{i^*}$
+                \Comment{Find the greatest distance}
+        \State $k \leftarrow k+1$
+            \Comment{Increment counter $k$}
     \EndWhile
+\color{black}
+    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} + [k/N]$
+            \Comment{Append latest estimate to results}
+\EndWhile
+\State $\widehat{h_{\rm final}} \leftarrow \frac{1}{\rm num\_iterations} \sum \widehat{\mathbf{h}}$
+            \Comment{Average results}
 \end{algorithmic}
 \caption{Calculates $H_{\x} = -\int p(\x) \log p(\x) \, d\x$
-         without evaluating $\log p(\x)$.}
+         without evaluating $\log p(\x)$. Technically, what it actually
+         estimates is $\int f(\x) \int f(\x')
+            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
+                        \, d\x' \, d\x$, the expected value of the
+         log-probability of a small region near $\x$.
+        The part highlighted in green is standard Nested Sampling.}
 \end{algorithm}
@@
 \begin{align}
-Q &= \int f(\x) \int f(\x')
+h &= \int f(\x) \int f(\x')
             \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
@@
 \end{itemize}
+To quantify the degree to which this question is unanswered by
+$\top$, we can write the question as a union of ideal questions:
+\begin{align}
+Q &= \left[\downarrow (x \in \{1, 2, 3\})\right] \cup
+     \left[\downarrow (x \in \{2, 3, 4\})\right] \cup ...
+\end{align}
+
+Its quantification is therefore
+\begin{align}
+H(Q) &= -\sum_{i=1}^3 p_i \log p_i - \sum_{i=2}^4 p_i \log p_i
++ \sum_{i=2}^3 p_i \log p_i
+\end{align}
 

### 2017-03-18 paper/ms.tex
Writing about NS

@@
 $N$ particles in the parameter space.
-
 Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
@@
 
-The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
-particle with the lowest likelihood.
+The $N$ particles are initially initialised from the prior $\pi(\x)$, and the particle with the lowest likelihood is found.
+{\todo Write more here}
 
@@
 \end{align}
-and the probability distribution for the sequence of $X$-values is
-obtained from
+and the probability distribution for the sequence of $X$-values of
+the discarded points is obtained from
 \begin{align}
 X_i &= \prod_{k=1}^i t_i.
+\end{align}
+The corresponding likelihood values of the discarded points are
+denoted $L_1, L_2, L_3$, and so on.
+
+Consider a particular likelihood value $\ell$ whose $X$-value
+we would like to know.
+The number of discarded points with likelihood below $\ell$, i.e.,
+the number of NS iterations taken reach a likelihood above
+$\ell$, is
+\begin{align}
+n(L_i < \ell) &= \sum_i \mathds{1}\left(L_i < \ell\right) \\
+              &= \sum_i \mathds{1}\left(X_i > X(\ell)\right)
+\end{align}
+The expected value of $n$ given the definition of the algorithm is
+\begin{align}
+\left<n(L_i < \ell)\right> &= 
+              \sum_i \left<\mathds{1}\left(X_i > X(\ell)\right)\right>
 \end{align}

### 2017-03-20 paper/ms.tex
Writing in algorithm section and caption

@@
 
+See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
+algorithm. The idea is to generate a reference particle from the distribution
+whose entropy is required, and then use Nested Sampling to measure the
+log-probability of a small region near that particle.
+
+The algorithm is written in terms of an unknown quantity $\x$ whose entropy is required. In applications, $\x$ may be parameters,
+a subset of the parameters, data, parameters and data jointly, or
+whatever.
+
 \begin{algorithm}
@@
          log-probability of a small region near $\x$.
-        The part highlighted in green is standard Nested Sampling.}
+        The part highlighted in green is standard Nested Sampling.
+        \label{alg:algorithm}}
 \end{algorithm}
@@
 
+In the sinusoidal signal example, I calculated the difference in
+mutual information between two observing strategies. It is possible
+that there exists a more direct method of obtaining this which
+is more efficient than running the two strategies separately.
+
 \section{Example 3: TBD}

### 2017-03-20 paper/ms.tex
Jared

@@
 Ewan Cameron (Oxford), David Hogg (NYU), Kevin Knuth (SUNY Albany),
-Iain Murray (Edinburgh). This work was supported by Centre for eResearch
+Iain Murray (Edinburgh), Jared Tobin (jtobin.io).
+This work was supported by Centre for eResearch
 at the University of Auckland.
@@
 
-\authorcontributions{Author Contributions}
+%\authorcontributions{Author Contributions}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-04-07 paper/ms.tex
Working on precisional question, getting unexpected result

@@
 don't need or want to know $\x$ to arbitrary precision. Using the central
-issue can also lead to counterintuitive results if you don't keep its
+issue can lead to counterintuitive results if you don't keep its specific
 definition in mind. For example, suppose $x$ could take any integer value
@@
 want to ask a question that is different from the central issue.
-For example, suppose we want to know the value of $x$ to within a tolerance
+For example, suppose $x \in \{1, ..., 10\}$ and
+we want to know the value of $x$ to within a tolerance
 of $\pm 1$. Any of the following statements would resolve the issue:
@@
 Q &= \left[\downarrow (x \in \{1, 2, 3\})\right] \cup
-     \left[\downarrow (x \in \{2, 3, 4\})\right] \cup ...
+     \left[\downarrow (x \in \{2, 3, 4\})\right] \cup
+     \left[\downarrow (x \in \{3, 4, 5\})\right] \cup ...
 \end{align}
 
-Its quantification is therefore
+The entropy of this question is
 \begin{align}
-H(Q) &= -\sum_{i=1}^3 p_i \log p_i - \sum_{i=2}^4 p_i \log p_i
-+ \sum_{i=2}^3 p_i \log p_i
+H(Q) &= f_{1,2,3}\\
+     & \quad \quad + \left(f_{2,3,4} - f_{2,3}\right) \\
+     & \quad \quad + \left(f_{3,4,5} - f_{3,4}\right) \\
+     & \quad \quad + ... \\
+     & \quad \quad + \left(f_{8,9,10} - f_{8,9}\right).
 \end{align}

### 2017-04-07 paper/precisional.py

@@ -1,0 +1,34 @@
+# Calculate the entropy of a precisional question
+# in a discrete problem.
+
+import numpy as np
+
+# Hypothesis space
+x = np.arange(0, 10)
+
+# Probabilities
+p = np.ones(len(x)) / len(x)
+
+# Precision of precisional question
+precision = 3
+
+# Entropy of central issue
+H_ci = -np.sum(p*np.log(p))
+print("H_ci = {H} nats".format(H=H_ci))
+
+H = 0.0
+
+# Downset terms
+for i in range(0, len(x) - precision + 1):
+    subset = p[i:i+precision]
+    print(subset)
+    H += -np.sum(subset * np.log(subset))
+
+# Subtract overlap terms
+for i in range(1, len(x) - precision + 1):
+    subset = p[i:i+precision-1]
+    print(subset)
+    H -= -np.sum(subset * np.log(subset))
+
+print(H)
+

### 2017-04-07 paper/precisional.py
Fixed the problem :)

@@ -20,15 +20,13 @@ H = 0.0
 
 # Downset terms
 for i in range(0, len(x) - precision + 1):
-    subset = p[i:i+precision]
-    print(subset)
-    H += -np.sum(subset * np.log(subset))
+    p_subset = p[i:i+precision].sum()
+    H += -p_subset * np.log(p_subset)
 
 # Subtract overlap terms
 for i in range(1, len(x) - precision + 1):
-    subset = p[i:i+precision-1]
-    print(subset)
-    H -= -np.sum(subset * np.log(subset))
+    p_subset = p[i:i+precision-1].sum()
+    H -= -p_subset * np.log(p_subset)
 
-print(H)
+print("H_precisional = {H} nats".format(H=H))
 

### 2017-04-08 paper/ms.tex
More promising setup

@@
 
+\subsection*{Continuous case}
+
+Consider a probability density function $f(x)$, defined on the real
+line. The probability contained in an interval
+$[x_0, x_0 + L]$, which has length $L$, is
+\begin{align}
+\int_{x_0}^{x_0 + L} f(x) \, dx &= F(x_0 + L) - F(x_0)
+\end{align}
+where $F(x)$ is the cumulative distribution function (CDF).
+This probability can be considered as a function of two variables,
+$x_0$ and $L$, which I will denote by $Q(,)$:
+\begin{align}
+Q(x_0, L) &= F(x_0 + L) - F(x_0).
+\end{align}
+The contribution of such an interval to an entropy expression
+is `minus $p\log p$', i.e.
+\begin{align}
+h(x_0, L) &= -Q(x_0, L) \log Q(x_0, L).
+\end{align}
+
+Consider the rate of change of $h$ as the interval is shifted to
+the right, but with its width $L$ held constant:
+\begin{align}
+\frac{\partial h}{\partial x_0} &= - \frac{\partial}{\partial x_0}
+  Q(x_0, L) \log Q(x_0, L)
+\end{align}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-04-08 paper/ms.tex
More working on the precisional appendix

@@
 \end{align}
-
 Consider the rate of change of $h$ as the interval is shifted to
@@
 \begin{align}
-\frac{\partial h}{\partial x_0} &= - \frac{\partial}{\partial x_0}
-  Q(x_0, L) \log Q(x_0, L)
+\frac{\partial h}{\partial x_0} &= \frac{\partial}{\partial x_0}
+    \left[-Q(x_0, L) \log Q(x_0, L)\right] \\
+    &= -\left[1 + \log Q(x_0, L)\right]\frac{\partial Q}{\partial x_0} \\
+    &= -\left[1 + \log Q(x_0, L)\right]\left[f(x_0 + L) - f(x_0)\right].
+\end{align}
+Consider also the rate of change of $h$ as the interval is
+expanded to the right, while keeping its left edge fixed:
+\begin{align}
+\frac{\partial h}{\partial L} &= \frac{\partial}{\partial L}
+    \left[-Q(x_0, L) \log Q(x_0, L)\right] \\
+    &= -\left[1 + \log Q(x_0, L)\right]\frac{\partial Q}{\partial L} \\
+    &= -\left[1 + \log Q(x_0, L)\right]f(x_0 + L).
+\end{align}
+
+The entropy of the precisional question is built up from
+$h$ terms. The extra entropy from moving from $x_0$ to
+$x_0 + \delta$, for small $\delta$, is
+\begin{align}
+\delta H &= h(x_0 + \delta, L) - h(x_0 + \delta, L - 2\delta) \\
+ &\approx \delta \frac{\partial h}{\partial x_0}
+    - \left[\delta \frac{\partial h}{\partial x_0}
+            -2\delta \frac{\partial h}{\partial L}\right] \\
+ &= 2\delta \frac{\partial h}{\partial L} \\
+ &= -2\delta\left[1 + \log Q(x_0, L)\right]f(x_0 + L)
+\end{align}
+Therefore
+\begin{align}
+\frac{dH}{dx_0} &= -2\left[1 + \log Q(x_0, L)\right]f(x_0 + L) \\
+  &= -2\left[1 + \log \left(F(x_0 + L) - F(x_0)\right)\right]f(x_0 + L)
 \end{align}

### 2017-05-05 paper/ms.tex
Final result looks great
Tidying up derivation

--- 2017-04-07 paper/ms.tex
@@
 
+\subsection*{Continuous case}
+
+Consider a probability density function $f(x)$, defined on the real
+line. The probability contained in an interval
+$[x_0, x_0 + L]$, which has length $L$, is
+\begin{align}
+\int_{x_0}^{x_0 + L} f(x) \, dx &= F(x_0 + L) - F(x_0)
+\end{align}
+where $F(x)$ is the cumulative distribution function (CDF).
+This probability can be considered as a function of two variables,
+$x_0$ and $L$, which I will denote by $P(,)$:
+\begin{align}
+P(x_0, L) &= F(x_0 + L) - F(x_0).
+\end{align}
+The contribution of such an interval to an entropy expression
+is $-P\log P$, i.e.,
+\begin{align}
+Q(x_0, L) &= -P(x_0, L) \log P(x_0, L).
+\end{align}
+Consider the rate of change of $Q$ as the interval is shifted to
+the right, but with its width $L$ held constant:
+\begin{align}
+\frac{\partial Q}{\partial x_0} &= \frac{\partial}{\partial x_0}
+    \left[-P\log P\right] \\
+    &= -\left[1 + \log P(x_0, L)\right]\frac{\partial P}{\partial x_0} \\
+    &= -\left[1 + \log P(x_0, L)\right]\left[f(x_0 + L) - f(x_0)\right].
+\end{align}
+Consider also the rate of change of $Q$ as the interval is
+expanded to the right, while keeping its left edge fixed:
+\begin{align}
+\frac{\partial Q}{\partial L} &= \frac{\partial}{\partial L}
+    \left[-P(x_0, L) \log P(x_0, L)\right] \\
+    &= -\left[1 + \log P(x_0, L)\right]\frac{\partial P}{\partial L} \\
+    &= -\left[1 + \log P(x_0, L)\right]f(x_0 + L).
+\end{align}
+
+The entropy of the precisional question is built up from
+$Q$ terms. The extra entropy from adding the interval
+$[x_0, x_0 + L]$ and removing the overlap $[x_0, x_0 + L - h]$, for small $h$, is
+\begin{align}
+\delta H &= Q(x_0, L) - Q(x_0, L - h)\\
+         &= h\frac{\partial Q}{\partial L} \\
+         &= -h\left[1 + \log P(x_0, L)\right]f(x_0 + L).
+\end{align}
+Therefore, the overall entropy is
+\begin{align}
+H &=
+      -\int_{-\infty}^\infty
+        \left[1 + \log P(x-L, L)\right]f(x)
+      \, dx
+\end{align}
+which is connected to the expected value of the log-probability.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-05-15 paper/ms.tex
Acknowledge James and Thomas

@@
 conversations about this topic: Ruth Angus (Flatiron Institute),
-Ewan Cameron (Oxford), David Hogg (NYU), Kevin Knuth (SUNY Albany),
+Ewan Cameron (Oxford), James Curran (Auckland),
+David Hogg (NYU), Kevin Knuth (SUNY Albany),
+Thomas Lumley (Auckland),
 Iain Murray (Edinburgh), Jared Tobin (jtobin.io).

### 2017-05-15 paper/prec2.py
Asymmetric density

@@ -2,16 +2,24 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-x = np.linspace(-10, 10, 50001)
+x = np.linspace(-10, 10, 200001)
 dx = x[1] - x[0]
-p = np.exp(-0.5*x**2) / np.sqrt(2*np.pi)
+p = x + 5.0
 p[np.abs(x) > 5] = 0
 L = 1.0
+p /= p.sum()*dx
+plt.plot(x, p)
+plt.show()
 
 H = 0.0
 for i in range(0, len(x)):
-    left = (x >= x[i] - L) & (x <= x[i])
+    left = (x >= x[i]) & (x <= x[i] + L)
     P = np.sum(p[left])*dx
+
+    # Alternative version uses
+    # right = (x >= x[i]) & (x <= x[i] + L)
+    # Instead of left
+
     if p[i] > 0:
         H += -p[i]*dx*(1 + np.log(P))
     if (i+1) % 1000 == 0:

### 2017-05-15 paper/prec2.py
slight amount of work

@@ -7,7 +7,7 @@ x = np.linspace(-10, 10, 50001)
 dx = x[1] - x[0]
 
 # Set up a density
-f = (x + 5.0)**3
+f = np.exp(-0.5*x**2)
 f[np.abs(x) > 5] = 0
 f /= np.trapz(f, x=x)   # Normalise
 
@@ -24,16 +24,17 @@ for i in range(0, len(x)):
     Pleft  = np.trapz(f[left],  x=x[left])
     Pright = np.trapz(f[right], x=x[right])
     if f[i] > 0.0:
-        integrand_left[i]  = f[i]*np.log(Pleft)
-        integrand_right[i] = f[i]*np.log(Pright)
+        integrand_left[i]  = -f[i]*np.log(Pleft)
+        integrand_right[i] = -f[i]*np.log(Pright)
     print(i+1, "/", len(x))
 
-# Do the integrals
-Hleft  = np.trapz(integrand_left,  x=x)
-Hright = np.trapz(integrand_right, x=x)
+# Do the integ=rals
+Hleft  = -1.0 + np.trapz(integrand_left,  x=x)
+Hright = -1.0 + np.trapz(integrand_right, x=x)
 
 # Calculate the difference
 d = np.abs((Hright - Hleft)/Hright)
+print("Hleft", Hleft)
 print("Relative difference =", d)
 
 # Plot it

### 2017-05-19 paper/ms.tex
Bit of appendix writing
slight amount of work
Sentence at the end

@@
 \begin{align}
-H &=
-      -\int_{-\infty}^\infty
+H &= -\int_{-\infty}^\infty
         \left[1 + \log P(x-L, L)\right]f(x)
-      \, dx
+      \, dx \\
+  &= -\int_{-\infty}^\infty
+        \left(1 + \log\left[F(x) - F(x-L)\right]\right)f(x)
+      \, dx.
 \end{align}
-which is connected to the expected value of the log-probability.
+Interestingly, calculating the log-probability of an interval
+to the right of $x$ gives an equivalent result:
+\begin{align}
+H' &=  -\int_{-\infty}^\infty
+        \left(1 + \log\left[F(x+L) - F(x)\right]\right)f(x)
+      \, dx \\
+   &= H.
+\end{align}
+This is not quite the same as the expected log-probability calculated by
+the version of the algorithm proposed in this paper (when the tolerance is not
+small).
 

### 2017-05-19 cpp/examples/Pareto.cpp
Vastly simpler Pareto example

@@ -8,32 +8,15 @@ namespace InfoNest
 {
 
 Pareto::Pareto()
-:us(100000)
-,top_ten(10)
+:us(N)
+,xs(N)
 {
 
 }
 
 void Pareto::generate(RNG& rng)
 {
-    // p(x_min) ~ pareto(10, 0.5)T(, 100000)
-    double log_x_min;
-    do
-    {
-        log_x_min = log(10.0) - 2.0*log(1.0 - rng.rand());
-        x_min = exp(log_x_min);
-    }while(x_min > 100000.0);
-
-    // p(alpha) ~ log-uniform(1, 5)
-    alpha = exp(log(5.0)*rng.rand());
-
-    // p(N) ~ pareto(10, 0.5)T(, 100000)
-    double logN;
-    do
-    {
-        logN = log(10.0) - 2.0*log(1.0 - rng.rand());
-        N = exp(logN);
-    }while(N > 100000.0);
+    alpha = exp(rng.randn());
 
     for(double& u: us)
         u = rng.rand();
@@ -43,104 +26,64 @@ void Pareto::generate(RNG& rng)
 
 void Pareto::generate_data()
 {
-    // All scores
-    std::vector<unsigned int> scores((size_t)N);
+    double inv_alpha = 1.0/alpha;
 
-    // Reciprocal of alpha
-    double log_x_min = log(x_min);
-    double one_over_alpha = 1.0/alpha;
-
-    // Generate scores
-    double e;
-    for(size_t i=0; i<scores.size(); ++i)
-    {
-        // Generate an exponential
-        e = -log(1.0 - us[i]);
-        scores[i] = (unsigned int)exp(log_x_min + one_over_alpha*e);
-    }
-
-    // Find top ten scores, put them in top_ten
-    std::sort(scores.begin(), scores.end());
-    size_t k=0;
-    for(size_t i=scores.size()-10; i<scores.size(); ++i)
-        top_ten[k++] = scores[i];
+    for(size_t i=0; i<N; ++i)
+       xs[i] = x_min*pow(1.0 - us[i], -inv_alpha);
 }
+
+double Pareto::sum_first_half() const
+{
+    double tot = 0.0;
+    for(size_t i=0; i<N/2; ++i)
+        tot += xs[i];
+    return tot;
+}
+
+double Pareto::sum_second_half() const
+{
+    double tot = 0.0;
+    for(size_t i=N/2; i<N; ++i)
+        tot += xs[i];
[…]

### 2017-05-19 cpp/examples/Pareto.h

@@ -12,26 +12,24 @@ namespace InfoNest
 class Pareto
 {
     private:
-        // Lower limit and slope
-        // of pareto distribution for scores
-        double x_min;
+        static constexpr double x_min = 1.0;
+
+        // Pareto slope
         double alpha;
 
-        // Number of games played (gets rounded down to
-        // an integer when needed)
-        double N;
-
-        // Source of 'random numbers' for generating the simulated data
+        // The data
+        static constexpr size_t N = 100;
         std::vector<double> us;
-
-        // Top ten scores
-        std::vector<unsigned int> top_ten;
+        std::vector<double> xs;
 
         // Function that generates the top ten scores
         void generate_data();
 
-        // Summaries of data, used to define data_distance
-        std::tuple<double, double, double> summaries() const;
+        // Sum of first half of data
+        double sum_first_half() const;
+
+        // Sum of second half of data
+        double sum_second_half() const;
 
     public:
         // Do-nothing constructor

### 2017-05-19 paper/ms.tex
Slight writing

@@
 
-Consider the basic problem of inferring a quantity $\mu$ from
+This example demonstrates how the algorithm may be used to determine the
+entropy for a quantity whose distribution can only be sampled.
+Consider the basic statistics problem of inferring a quantity $\mu$ from
 100 observations $\d = \{d_1, ..., d_{100}\}$ whose
@@
 \end{align}
-The true value of $H_{\d}$, calculated analytically, is $146.499$.
+The true value of the differential entropy $h_{\d}$ is available analytically
+in this case, and its value is $146.499$. However, in most situations, the
+entropy of a marginal distribution is not available in closed form.
 
-The result was
+From the algorithm, the expected log-probability was
 \begin{align}
-H_{\d} &= 928.80 \pm 0.38.
+?_{\d} &= 928.80 \pm 0.38.
 \end{align}
 The log-volume of a 100-dimensional ball of radius $10^{-3}$ is
--782.017. Therefore, the measured differential en
+-782.017. Therefore, the measured differential entropy was
+
 \section{Example 2: Measuring the period of an oscillating signal}

### 2017-05-21 paper/ms.tex
Bit of tidying

@@
 However, in practice, it can be difficult to compute these quantities
-for arbitrary probability distributions, particularly if some probability
+for arbitrary probability distributions, particularly if the probability
 mass functions or densities cannot be evaluated. This paper introduces a
 computational approach to evaluating entropies of probability distributions
-that can be sampled. The method is based on Nested Sampling, and is related to Approximate Bayesian Computation. I demonstrate the method
+that can only be sampled. The method is based on Nested Sampling,
+and is related to Approximate Bayesian Computation. I demonstrate the method
 on a simple gaussian example where the key quantities are known analytically,
@@
 Throughout this paper, I use the compact
-`overloaded' notation for probability distributions favoured by
-Bayesians, writing $p(x)$ (for either a probability mass function
+`overloaded' notation for probability distributions favoured by many
+Bayesian writers, writing $p(x)$ (for either a probability mass function
 or a probability density function) instead
@@
 the entropy is a property of the value of $x$ itself, rather than
-our probability distribution for it.
+the probability distribution we have assigned to it.
 
@@
 of how the data is related to the parameters
-(i.e., if the parameters where known, what data would likely be observed?). By the product rule, this yields a {\em joint prior}
+(i.e., if the parameters where known, what data would likely be observed?).
+By the product rule, this yields a {\em joint prior}
 \begin{align}
@@
 
-In practical applications of Bayesian inference, the following operations
-are usually feasible, and are assumed to be possible at low computational
-cost:
+In practical applications, the following operations
+are usually feasible and have a low computational cost:
 \begin{enumerate}
   \item Points can be generated from the prior $p(\x)$;
-\item Simulated datasets can be generated from $p(\d | \x)$ for a given value of $\x$;
-\item The likelihood, $p(\d | \x)$ can be evaluated cheaply for any $\d$ and $\x$.
+  \item Simulated datasets can be generated from $p(\d | \x)$ for any
+        given value of $\x$;
+  \item The likelihood, $p(\d | \x)$, can be evaluated cheaply for any
+        $\d$ and $\x$.
 \end{enumerate}
+Throughout this paper I assume these operations are available and inexpensive.
 
@@
 
-This is difficult to compute because the expression
+Suppose we wanted to compute $H_{\x | \d}$.
+This would be difficult because the expression
 for the posterior distribution
@@
 \end{align}
-contains the marginal likelihood or `evidence'
+contains the marginal likelihood integral:
 \begin{align}
-p(\d) &= \int p(\x) p(\d | \x) \, d\x.
+p(\d) &= \int p(\x) p(\d | \x) \, d\x
 \end{align}
+also known as the `evidence'. In this paper I use the term evidence because
+for brevity.
 
@@
 The former measures the degree to which one question answers
-another, so is a function of two questions. The latter measures the
-degree to which a {\em statement} answered a question, and is a function
-of a question and a statement.
+another (ex ante), so is a function of two questions. The latter measures the
+degree to which a {\em statement} answered a question (ex post),
+and is a function of a question and a statement.
 
@@
 As with the conditional entropy, we see that the difficulty appears in the
-form of the log of the marginal likelihood,
+form of the log evidence,
 $\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.
@@
 Nested Sampling \citep[NS][]{skilling2006nested} is an algorithm whose aim is
-to calculate the marginal likelihood
+to calculate the evidence
 \begin{align}
@@
 can be used to construct a
-version of NS that produces unbiased estimates of the marginal likelihood.
+version of NS that produces unbiased estimates of the evidence.
 However, it can also be used to construct an unbiased estimator of

### 2017-05-21 paper/ms.tex
Non-cartesian distances

@@
 
-The algorithm is written in terms of an unknown quantity $\x$ whose entropy is required. In applications, $\x$ may be parameters,
+The algorithm is written in terms of an unknown quantity $\x$
+whose entropy is required. In applications, $\x$ may be parameters,
 a subset of the parameters, data, parameters and data jointly, or
@@
             \Comment{the number of Nested Sampling particles to use}
-\State $\epsilon \geq 0$
+\State $L \geq 0$
             \Comment{the tolerance} \\
@@
             \Comment{Find the greatest distance}
-    \While{$d_{\rm max} > \epsilon$}
+    \While{$d_{\rm max} > L$}
         \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
@@
          estimates is $\int f(\x) \int f(\x')
-            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
+            \log \left[ P(d(\x', \x) < L | \x) \right]
                         \, d\x' \, d\x$, the expected value of the
@@
 h &= \int f(\x) \int f(\x')
-            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
+            \log \left[ P(d(\x', \x) < L | \x) \right]
                         \, d\x' \, d\x.
@@
 we can use the fact that density equals mass divided
-by volume. Assume $\epsilon$ is small, so that
+by volume. Assume $L$ is small, so that
 \begin{align}
-P(d(\x', \x) < \epsilon | \x)
+P(d(\x', \x) < L | \x)
     &\approx
-    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
+    f(\x) \int_{d(\x', \x) < L} \, d\x'\\
     &= f(\x) \, V
 \end{align}
-where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.
+where $V$ is the volume of the region where $d(\x', \x) < L$.
 
-If the distance function $d(\x', \x)$ is chosen to be Cartesian,
-the constraint $d(\x', \x) < \epsilon$ corresponds to a ball in the space
-of possible $\x'$ values.
-The log-volume of a ball of radius $\epsilon$ in $n$ dimensions is
+In a one-dimensional parameter space, the algorithm will calculate
 \begin{align}
-\log V(\epsilon; n) &= \frac{n}{2}\log \pi
-                        - \log \Gamma\left(\frac{n}{2} + 1\right)
-                        + n \log \epsilon,
+h(L) &= \int f(\x) \int f(\x')
+             \log \left[ \int_{x - L}^x f(x') \right]
+                        \, d\x' \, d\x.
 \end{align}
-and provides the conversion factor between the differential entropy and
-the entropy actually computed from the algorithm.
+See the Appendix for a discussion of this quantity. To obtain the differential
+entropy, compute $h(L)$ for small $L$ and then add $\log(L)$.
 

### 2017-05-21 paper/ms.tex
Bit of writing

@@
 where the sum is over all of the possible values of $x$ under consideration.
-
 The entropy quantifies the degree to which the issue
-``what is the value of $x$, precisely?'' remains unresolved. For example,
+``what is the value of $x$, precisely?'' remains unresolved
+\citep{knuth_questions}. For example,
 if there is only one possible value which has probability 1, $H_x$ is
@@
 describe uncertainty, often about unknown quantities
-(``parameters'') $\x$. Some data $\d$, initially unknown but
+(`parameters') $\x$. Some data $\d$, initially unknown but
 thought to be relevant to $\x$, is obtained.
@@
 Bayesian inference \citep{caticha2008lectures}.
-
 In practical applications, the following operations
@@
 ``what is the value of the pair $(\x, \d)$, precisely?'' remains unanswered.
-
-The degree to which the question ``what is the value of $\x$?'' is
-answered by the question ``what is the value of $\d$?'' is given by the
+The degree to which the question ``what is the value of $\x$?'' would remain
+unresolved if $\d$ were resolved is given by the
 conditional entropy
@@
 
-Suppose we wanted to compute $H_{\x | \d}$.
+Suppose we wanted to compute $H_{\x | \d}$, perhaps to compare it to
+$H_{\x}$ and quantify how much might be learned about $\x$.
 This would be difficult because the expression
@@
 \end{align}
-also known as the `evidence'. In this paper I use the term evidence because
-for brevity.
+also known as the `evidence', which tends to be computable but costly.
 
@@
 entropy of $\x$ given a particular value of $\d$, which might be written
-$H_{\x | \d=\d_0}$.
+$H_{\x | \d=\d_0}$ (in traditional notation, $H(X|Y)$ and $H(X|Y=y)$).
 The former measures the degree to which one question answers
@@
 
-For experimental design purposes, maximising either $I_{\x, \d}$ or
-$H_{\x | \d}$ will produce the same result. {\todo Mention reference priors?}
+For experimental design purposes, if we want to maximise the expected amount
+of information obtained from data, maximising either $I_{\x, \d}$ or
+$H_{\x | \d}$ will produce the same result because the prior $p(\x)$ does not
+vary with the experimental design. Reference priors {\bf cite Bernardo}
+also make use of the mutual information but in a different way.
 
@@
 \section{The algorithm}
+
+We would like to calculate the expected value of a log-probability:
 
@@
 Continuous hypothesis spaces are commonly used in practice. Suppose
-$\x$ has probability density $f(\x)$. The algorithm actually estimates
+$\x$ has probability density $f(\x)$, so the differential entropy is
+\begin{align}
+H &= -\int_{-\infty}^{\infty} f(\x) \log f(\x) \, d\x
+\end{align}
+
+What the algorithm actually estimates is
 \begin{align}
@@
 If we actually want the differential
-entropy
-\begin{align}
-h_{\x} &= -\int f(x) \log f(x) \, dx
-\end{align}
-we can use the fact that density equals mass divided
+entropy, we can use the fact that density equals mass divided
 by volume. Assume $L$ is small, so that

### 2017-05-21 paper/ms.tex
A bit of clarification in the poisson process bit

@@
 
-When $N=1$, the sequence of $\ln X$ values is a Poisson process with rate
-1. \citet{Walter2015} showed how this view of the NS sequence of points
+The sequence of $\ln X$ values produced by a Nested Sampling run can be
+considered as a Poisson process with rate equal to $N$, the number of
+particles. For example, when $N=1$, the distribution for $\ln X_{i+1}$
+given $\ln X_i$ is a shifted (and reflected) exponential with rate 1.
+\citet{Walter2015} showed how this view of the NS sequence of points
 can be used to construct a
-version of NS that produces unbiased estimates of the evidence.
+version of NS that produces unbiased (in the frequentist sense)
+estimates of the evidence.
 However, it can also be used to construct an unbiased estimator of
@@
 
-Given the definition of the NS algorithm,
-the probability distribution for the proportion $t$ of the prior mass
-retained on a given iteration is
+Consider a particular likelihood value $\ell$ whose corresponding $X$-value
+we would like to know. Since the $\ln X_i$ values have the same distribution
+as the arrival times of a Poisson process with rate $N$, the probability
+distribution for the number of points in an interval of length $t$ is
+Poisson with expected value $\lambda t$. Therefore the expected number of
+NS discarded points with likelihood below $\ell$ is $-\ln X(\ell)$:
 \begin{align}
-t_i &\sim \textnormal{Beta}(N, 1),
+\big< n(L(\x_i) \leq \ell) \big> &= -\ln X(\ell).
 \end{align}
-and the probability distribution for the sequence of $X$-values of
-the discarded points is obtained from
-\begin{align}
-X_i &= \prod_{k=1}^i t_i.
-\end{align}
-The corresponding likelihood values of the discarded points are
-denoted $L_1, L_2, L_3$, and so on.
-
-Consider a particular likelihood value $\ell$ whose $X$-value
-we would like to know.
-The number of discarded points with likelihood below $\ell$, i.e.,
-the number of NS iterations taken reach a likelihood above
-$\ell$, is
-\begin{align}
-n(L_i < \ell) &= \sum_i \mathds{1}\left(L_i < \ell\right) \\
-              &= \sum_i \mathds{1}\left(X_i > X(\ell)\right)
-\end{align}
-The expected value of $n$ given the definition of the algorithm is
-\begin{align}
-\left<n(L_i < \ell)\right> &= 
-              \sum_i \left<\mathds{1}\left(X_i > X(\ell)\right)\right>
-\end{align}
+We can therefore take $n(L(\x_i) \leq \ell)$, the number of points in the
+sequence with likelihood below $\ell$, as an unbiased estimator of
+$-\ln X(\ell)$.
 

### 2017-05-24 paper/ms.tex
Bernardo
Back to Euclidean distances

@@
 
+\subsection{Discrete vs. differential entropies}
+
+Continuous hypothesis spaces are commonly used in practice. Suppose
+$\x$ has probability density $f(\x)$. The algorithm actually estimates
+\begin{align}
+h &= \int f(\x) \int f(\x')
+            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
+                        \, d\x' \, d\x.
+\end{align}
+If we actually want the differential
+entropy
+\begin{align}
+h_{\x} &= -\int f(x) \log f(x) \, dx
+\end{align}
+we can use the fact that density equals mass divided
+by volume. Assume $\epsilon$ is small, so that
+\begin{align}
+P(d(\x', \x) < \epsilon | \x)
+    &\approx
+    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
+    &= f(\x) \, V
+\end{align}
+where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.
+
+If the distance function $d(\x', \x)$ is chosen to be Euclidean,
+the constraint $d(\x', \x) < \epsilon$ corresponds to a ball in the space
+of possible $\x'$ values.
+The log-volume of a ball of radius $\epsilon$ in $n$ dimensions is
+\begin{align}
+\log V(\epsilon; n) &= \frac{n}{2}\log \pi
+                        - \log \Gamma\left(\frac{n}{2} + 1\right)
+                        + n \log \epsilon,
+\end{align}
+and provides the conversion factor between the differential entropy and
+the entropy actually computed from the algorithm.
+
 Continuous hypothesis spaces are commonly used in practice. Suppose

### 2017-05-26 paper/ms.tex
Fixing up volume stuff

--- 2017-05-21 paper/ms.tex
@@
 \newcommand{\n}{\boldsymbol{\eta}}
+\newcommand{\depth}{\textnormal{depth}}
 
@@
 $H_{\x | \d}$ will produce the same result because the prior $p(\x)$ does not
-vary with the experimental design. Reference priors {\bf cite Bernardo}
-also make use of the mutual information but in a different way.
+vary with the experimental design. Reference priors
+\citep{bernardo2005reference} also make use of the mutual information but in a
+different way.
 
@@
 Continuous hypothesis spaces are commonly used in practice. Suppose
-$\x$ has probability density $f(\x)$, so the differential entropy is
+$\x$ has probability density $f(\x)$. The algorithm actually estimates
+the `depth', or minus the expected value of the log-probability near $x$:
 \begin{align}
-H &= -\int_{-\infty}^{\infty} f(\x) \log f(\x) \, d\x
-\end{align}
-
-What the algorithm actually estimates is
-\begin{align}
-h &= \int f(\x) \int f(\x')
+h &= -\int f(\x) \int f(\x')
             \log \left[ P(d(\x', \x) < L | \x) \right]
@@
 \end{align}
+for some tolerance $L$.
+
 If we actually want the differential
-entropy, we can use the fact that density equals mass divided
+entropy
+\begin{align}
+h_{\x} &= -\int f(x) \log f(x) \, dx
+\end{align}
+we can use the fact that density equals mass divided
 by volume. Assume $L$ is small, so that
 \begin{align}
-P(d(\x', \x) < L | \x)
+P(d(\x', \x) < \epsilon | \x)
     &\approx
@@
 
-In a one-dimensional parameter space, the algorithm will calculate
+If the distance function $d(\x', \x)$ is chosen to be Euclidean,
+the constraint $d(\x', \x) < L$ corresponds to a ball in the space
+of possible $\x'$ values.
+The log-volume of a ball of radius $L$ in $n$ dimensions is
 \begin{align}
-h(L) &= \int f(\x) \int f(\x')
-             \log \left[ \int_{x - L}^x f(x') \right]
-                        \, d\x' \, d\x.
+\log V(\epsilon; n) &= \frac{n}{2}\log \pi
+                        - \log \Gamma\left(\frac{n}{2} + 1\right)
+                        + n \log L,
 \end{align}
-See the Appendix for a discussion of this quantity. To obtain the differential
-entropy, compute $h(L)$ for small $L$ and then add $\log(L)$.
+and provides the conversion factor between the differential entropy and
+the entropy actually computed from the algorithm.
+
+If $L$ is not small, we might think we are calculating the entropy associated
+with the question `what is $x$, to within a tolerance of $\pm L$?'.
+This is not quite correct. See the Appendix for a discussion of this
+type of question.
 
@@
 \begin{align}
-?_{\d} &= 928.80 \pm 0.38.
+\depth_{\d} &= 928.80 \pm 0.38.
 \end{align}

### 2017-05-26 cpp/mutual_info.py
Just calculate log-volumes

@@ -1,35 +1,0 @@
-# A little script for mutual information calculations
-import matplotlib.pyplot as plt
-import numpy as np
-from scipy.special import gammaln
-
-# log-volume of n-ball
-def log_volume(n, e):
-    return 0.5*n*np.log(np.pi) - gammaln(0.5*n+1) + n*np.log(e)
-
-# Name of the problem
-name = "normal"
-results1 = np.loadtxt("results/" + name + "1.txt")
-results2 = np.loadtxt("results/" + name + "2.txt")
-results3 = np.loadtxt("results/" + name + "3.txt")
-
-# Parameter and data dimensions
-dim = [1, 100]
-dim.append(sum(dim))
-
-# Tolerances used
-tol = [1E-3, 1E-3, 1E-3]
-
-# log-volumes
-lvs = [ log_volume(dim[i], tol[i]) for i in range(0, 3) ]
-
-L = min([len(results1), len(results2), len(results3)])
-I =   (results1[0:L] + lvs[0]) \
-    + (results2[0:L] + lvs[1]) \
-    - (results3[0:L] + lvs[2])
-
-plt.plot(I)
-plt.axhline(0.0, linestyle="--", alpha=0.3)
-print("I = " + str(I.mean()) + " +- " + str(I.std()/np.sqrt(L)) + " nats.")
-plt.show()
-

### 2017-05-26 cpp/volume.py

@@ -1,0 +1,13 @@
+# Calculate log volume of an n-ball
+import numpy as np
+from scipy.special import gammaln
+
+# log-volume of n-ball
+def log_volume(n, L):
+    return 0.5*n*np.log(np.pi) - gammaln(0.5*n+1) + n*np.log(L)
+
+# What I use for normal example
+print(log_volume(1, 1E-3))
+print(log_volume(100, 1E-3*np.sqrt(100)))
+print(log_volume(101, 1E-3*np.sqrt(101)))
+

### 2017-05-27 paper/ms.tex
Updated results

@@
 
-From the algorithm, the expected log-probability was
+I ran the algorithm and computed the depth using a tolerance of
+$L=10^{-3}\sqrt{100}$, so that the RMS difference between
+points in the two datasets was about $10^{-3}$.
+From the algorithm, the average depth was
 \begin{align}
-\depth_{\d} &= 928.80 \pm 0.38.
+\depth_{\d} &= 698.25 \pm 0.34.
 \end{align}
-The log-volume of a 100-dimensional ball of radius $10^{-3}$ is
--782.017. Therefore, the measured differential entropy was
+The log-volume of a 100-dimensional ball of radius
+$L=10^{-3}\sqrt{100}$ is
+$-551.76$. Therefore, the differential entropy is estimated to be
+\begin{align}
+H_{\d} &\approx 146.49 \pm 0.34.
+\end{align}
+which is very close to the true value (suspiciously close, but this
+was just a fluke).
 

### 2017-05-29 paper/ms.tex
Updated results

@@
 \newcommand{\n}{\boldsymbol{\eta}}
-\newcommand{\depth}{\textnormal{depth}}
+\newcommand{\depth}{(\textnormal{depth})}
 
@@
 Letting $\tau = \log_{10} T$, and treating
-$A$ and $\phi$ are nuisance parameters, I calculated the entropies
-of $\tau$, $\d$, and $(\tau, \d)$.
-These results were computed using a tolerance of 0.001 for each of the three
-runs. The results are:
+$A$ and $\phi$ are nuisance parameters, I calculated the depths
+for $\tau$, $\d$, and $(\tau, \d)$.
+These results were computed using a tolerance of $L=0.001$
+for $\tau$, $0.001\sqrt{101}$ for $\d$, and $0.001\sqrt{102}$ for
+the joint run. The results were:
 \begin{align}
-H_{\tau}    &= 6.243  \pm 0.024 \\
-H_{\d}      &= 712.43 \pm 0.35   \\
-H_{\tau,\d} &= 715.81 \pm 0.34
+\depth_{\tau}    &= 6.216  \pm 0.025 \\
+\depth_{\d}      &= 480.13 \pm 0.32   \\
+\depth_{\tau,\d} &= 479.96 \pm 0.31
 \end{align}

### 2017-05-29 paper/ms.tex
More sinusoid results

@@
 quantifies uncertainty by generalising the log-volume
-of the plausible region, where volume is defined with respect to $dx$.
+of the plausible region, when volume is defined with respect to $dx$.
 
@@
 or a probability density function) instead
-of the more correct but cumbersome $P(X=x)$ or $f_X(x)$.
+of the more cumbersome $P(X=x)$ or $f_X(x)$ where one must always
+distinguish the `random variable' itself ($X$) from a similarly-named
+dummy variable ($x$).
 Probability distributions are implicitly conditional on some prior
@@
 
-To convert these to differential entropies, I applied the log-volume
-correction factors for 1, 101, and 102 dimensions, which are
-$-6.215$, $-790.32$, and $-798.62$ respectively. This gives
+To convert these to differential entropies, I added the log-volumes
+for 1, 101, and 102 dimensions, which are
+$-6.215$, $-557.252$, and $-562.746$ respectively. This gives
 the following values for the differential entropies:
 \begin{align}
-H'_{\tau}     &=  0.028 \pm 0.024 \label{eqn:h_tau}\\
-H'_{\d}       &= -77.89 \pm 0.35   \\
-H'_{\tau, \d} &= -82.81 \pm 0.34   \label{eqn:h_joint}
+H'_{\tau}     &=  0.001 \pm 0.025 \label{eqn:h_tau}\\
+H'_{\d}       &= -77.12 \pm 0.32   \\
+H'_{\tau, \d} &= -82.79 \pm 0.31   \label{eqn:h_joint}
 \end{align}
-The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
-uncertainties in each of these three quantities is quite large, and would lead
-to a large uncertainty on $I_{\tau, \d}$ if computed from the results
-in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}. To reduce this
-uncertainty, we computed the three entropies using a common random number
-generator for the reference pairs. This is an example of the
+The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$,
+and can be computed from the results listed
+in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}.
+However, each of these three runs were produced with an identical random
+seed for the reference points, so there is a way to reduce the uncertainty
+by calculating $H_{\tau} + H_{\d} - H_{\tau, \d}$ for each reference pair.
+This is an instance of the
 {\em common random numbers} technique of variance reduction.
 The resulting  mutual information is
-$I_{\tau, \d} = 4.96 \pm 0.38$ nats.
+$I_{\tau, \d} = 5.67 \pm 0.30$ nats.
 

### 2017-05-31 paper/ms.tex
Newer results

@@
 \begin{align}
-H_{\tau}     &= 6.243  \pm 0.024 \\
-H_{\d}       &= 712.77 \pm 0.35   \\
-H_{\tau, \d} &= 715.84 \pm 0.35    \\
-I_{\tau, \d} &= 5.21 \pm 0.40
+\depth_{\tau}      &= 6.216  \pm 0.025 \\
+\depth_{\d}        &= 479.47 \pm 0.31  \\
+\depth_{\tau, \d}  &= 480.27 \pm 0.32  \\
+I_{\tau, \d}       &= 4.69 \pm 0.32
 \end{align}
@@
 schedules, assuming a uniform prior on this difference,
-is just $0.25 \pm 0.55$ nats, suggesting (extremely weakly)
+is just $0.98 \pm 0.44$ nats, suggesting (extremely weakly)
 a slight advantage to the
-uneven observing schedule, in terms of the data's relevance to inference
+even observing schedule, in terms of the data's relevance to inference
 of the log-period of the signal. The situation may be different if we

### 2017-06-02 paper/ms.tex
Changed to more standard notation

@@
 \begin{align}
-H_{x} &= -\sum_{x} p(x) \log p(x)
+H(x) &= -\sum_{x} p(x) \log p(x)
 \end{align}
@@
 \citep{knuth_questions}. For example,
-if there is only one possible value which has probability 1, $H_x$ is
+if there is only one possible value which has probability 1, $H(x)$ is
 zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
@@
 If there are $N$ possibilities with equal probabilities $1/N$ each,
-then the entropy is $H_x = \log(N)$.
-If the space of possible $x$-values is continuous, the differential
-entropy
+then the entropy is $H(x) = \log(N)$.
+If the space of possible $x$-values is continuous so that $p(x)$ is a
+probability density function, the differential entropy
 \begin{align}
-H_{x} &= -\int p(x) \log p(x) \, dx
+H(x) &= -\int p(x) \log p(x) \, dx
 \end{align}
@@
 
-I prefer to write the entropy of $x$ as $H_x$ rather than $H(x)$
-(or $H(X)$ as is often seen), because the latter can wrongly imply that
-the entropy is a property of the value of $x$ itself, rather than
-the probability distribution we have assigned to it.
+Even though the entropy is written as $H(x)$, it is imperative that we
+remember it is not a property of the value of $x$ itself, but a property
+of the probability distribution we use to describe our ignorance of $x$.
 
@@
 \begin{align}
-H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
+H(\x | \d) &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
 \end{align}
@@
 
-Suppose we wanted to compute $H_{\x | \d}$, perhaps to compare it to
-$H_{\x}$ and quantify how much might be learned about $\x$.
+Suppose we wanted to compute $H(\x | \d)$, perhaps to compare it to
+$H(\x)$ and quantify how much might be learned about $\x$.
 This would be difficult because the expression
@@
 
-It is important to distinguish between $H_{\x | \d}$ and the
+It is important to distinguish between $H(\x | \d)$ and the
 entropy of $\x$ given a particular value of $\d$, which might be written
-$H_{\x | \d=\d_0}$ (in traditional notation, $H(X|Y)$ and $H(X|Y=y)$).
+$H(\x | \d=\d_0)$.
 The former measures the degree to which one question answers
@@
 \begin{align}
-I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
+I(\x; \d) &= \sum_{\x} \sum_{\d} p(\x, \d)
                        \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
-           &= H_{\x} + H_{\d} - H_{\x, \d}\\
-           &= H_{\d} - H_{\d | \x}\\
-           &= H_{\x} - H_{\x | \d}.
+           &= H(\x) + H(\d) - H(\x, \d)\\
+           &= H(\d) - H(\d | \x)\\
+           &= H(\x) - H(\x | \d).
 \end{align}
@@
 \begin{align}
-I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x)p(\d | \x)
+I(\x; \d) &= \sum_{\x} \sum_{\d} p(\x)p(\d | \x)
               \left[\log p(\d | \x) - \log p(\d)\right],
@@
 For experimental design purposes, if we want to maximise the expected amount
-of information obtained from data, maximising either $I_{\x, \d}$ or
-$H_{\x | \d}$ will produce the same result because the prior $p(\x)$ does not
+of information obtained from data, maximising either $I(\x; \d)$ or
+$H(\x | \d)$ will produce the same result because the prior $p(\x)$ does not
 vary with the experimental design. Reference priors
@@
 
-The $N$ particles are initially initialised from the prior $\pi(\x)$, and the particle with the lowest likelihood is found.
+The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
+particle with the lowest likelihood is found.
 {\todo Write more here}
@@
 \end{algorithmic}
-\caption{Calculates $H_{\x} = -\int p(\x) \log p(\x) \, d\x$
+\caption{Calculates $H(\x) = -\int p(\x) \log p(\x) \, d\x$
          without evaluating $\log p(\x)$. Technically, what it actually
@@
 \begin{align}
-h_{\x} &= -\int f(x) \log f(x) \, dx
+H(\x) &= -\int f(x) \log f(x) \, dx
 \end{align}
@@
 \end{align}
-The true value of the differential entropy $h_{\d}$ is available analytically
+The true value of the differential entropy $H(\d)$ is available analytically
 in this case, and its value is $146.499$. However, in most situations, the
@@
 \begin{align}
-H_{\d} &\approx 146.49 \pm 0.34.
+H(\d) &\approx 146.49 \pm 0.34.
 \end{align}
@@
 \begin{align}
-H'_{\tau}     &=  0.001 \pm 0.025 \label{eqn:h_tau}\\
-H'_{\d}       &= -77.12 \pm 0.32   \\
-H'_{\tau, \d} &= -82.79 \pm 0.31   \label{eqn:h_joint}
+H(\tau)     &=  0.001 \pm 0.025 \label{eqn:h_tau}\\
+H(\d)       &= -77.12 \pm 0.32   \\
+H(\tau, \d) &= -82.79 \pm 0.31   \label{eqn:h_joint}
 \end{align}
-The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$,
+The mutual information is equal to $H(\tau) + H(\d) - H(\tau, \d)$,
 and can be computed from the results listed
@@
 seed for the reference points, so there is a way to reduce the uncertainty
-by calculating $H_{\tau} + H_{\d} - H_{\tau, \d}$ for each reference pair.
+by calculating $H(\tau) + H(\d) - H(\tau, \d)$ for each reference pair.
 This is an instance of the
@@
 The resulting  mutual information is
-$I_{\tau, \d} = 5.67 \pm 0.30$ nats.
+$I(\tau; \d) = 5.67 \pm 0.30$ nats.
 
@@
 \depth_{\tau, \d}  &= 480.27 \pm 0.32  \\
-I_{\tau, \d}       &= 4.69 \pm 0.32
+I(\tau; \d)        &= 4.69 \pm 0.32
 \end{align}
@@
 \begin{align}
-I_{\x, \d} &= \iint p(\x, \d)
+I(\x; \d) &= \iint p(\x, \d)
                         \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
@@
 $\bphi$ and nuisance parameters $\boldeta$, such that we want to calculate
-the mutual information $I_{\bphi, \d}$. This is given by
+the mutual information $I(\bphi; \d)$. This is given by
 \begin{align}
-I_{\bphi, \d} &= \iint p(\bphi, \d)
+I(\bphi; \d) &= \iint p(\bphi, \d)
                         \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
@@
 \begin{align}
-I_{\bphi, \d} &= \iint
+I(\bphi; \d) &= \iint
   \left(\int p(\bphi, \boldeta)p(\d | \bphi, \boldeta) \, d\boldeta\right)

### 2017-06-07 paper/ms.tex
Some writing about NS

@@
 $N$ particles in the parameter space.
+The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
+particle with the lowest likelihood, $L^*_1$, is found and its details are
+recorded in the output. This worst particle is then discarded
+and replaced by a new particle
+drawn from the prior $\pi$ but subject to the constraint that its likelihood
+must be above $L^*_1$. This process is repeated, resulting in an
+increasing sequence of likelihoods
+\begin{align}
+L^*_1, L^*_2, L^*_3..., 
+\end{align}
+
 Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
@@
 \end{align}
-
-The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
-particle with the lowest likelihood is found.
-{\todo Write more here}
+\citet{skilling}'s key insight was that we can estimate the corresponding
+$X$-values of the discarded points can be estimated from the nature of the
+algorithm. Specifically, each iteration reduces the prior mass by approximately
+a factor $e^{-1/N}$. More accurately, the conditional distribution of
+$X_{i}$ given $X_{i-1}$ (and the definition of the algorithm)
+is obtainable by letting
+$t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$.
 
@@
 we want to know the value of $x$ to within a tolerance
-of $\pm 1$. Any of the following statements would resolve the issue:
+of $\pm 1$. Any of the following statements would
+resolve the issue:
 \begin{itemize}
@@
 \begin{align}
-H(Q) &= f_{1,2,3}\\
-     & \quad \quad + \left(f_{2,3,4} - f_{2,3}\right) \\
-     & \quad \quad + \left(f_{3,4,5} - f_{3,4}\right) \\
+H(Q) &= h_{1,2,3}\\
+     & \quad \quad + \left(h_{2,3,4} - h_{2,3}\right) \\
+     & \quad \quad + \left(h_{3,4,5} - h_{3,4}\right) \\
      & \quad \quad + ... \\
-     & \quad \quad + \left(f_{8,9,10} - f_{8,9}\right).
+     & \quad \quad + \left(h_{8,9,10} - h_{8,9}\right).
 \end{align}
+where $h_{x} = -P(x)\log P(x)$.
 

### 2017-06-19 paper/ms.tex
A bit more detail in the algorithm section

@@
 
-We would like to calculate the expected value of a log-probability:
+The above insight, that the number of NS discarded points with likelihood
+below $\ell$ has expected value $-\ln X(\ell)$, is the basis of the algorithm.
+If we wish to measure a log-probability $\ln X(\ell)$, we can use NS to do it.
+See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
+algorithm. The algorithm is written in terms of an unknown quantity $\x$
+whose entropy is required. In applications, $\x$ may be parameters,
+a subset of the parameters, data, some function of the data,
+parameters and data jointly, or whatever.
 
-See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
-algorithm. The idea is to generate a reference particle from the distribution
-whose entropy is required, and then use Nested Sampling to measure the
-log-probability of a small region near that particle.
-
-The algorithm is written in terms of an unknown quantity $\x$
-whose entropy is required. In applications, $\x$ may be parameters,
-a subset of the parameters, data, parameters and data jointly, or
-whatever.
+The idea is to generate a reference particle $\x_{\rm ref}$
+from the distribution $p(\x)$
+whose entropy is required. Then, a Nested Sampling run evolves a set of
+particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$, is evolved
+towards $\x_{\rm ref}$ in order to measure the log-probability near
+$\x_{\rm ref}$.
 
@@
         log-probability of a small region near $\x$.
-        The part highlighted in green is standard Nested Sampling.
+        The part highlighted in green is standard Nested Sampling
+        with quasi-prior $p(\x)$ and quasi-likelihood given by minus a
+        distance function $d(\x', \x)$.
         \label{alg:algorithm}}

### 2017-06-19 paper/ms.tex
Fixed equation

@@
 
-The sequence of $\ln X$ values produced by a Nested Sampling run can be
+The sequence of $-\ln X$ values produced by a Nested Sampling run can be
 considered as a Poisson process with rate equal to $N$, the number of
-particles. For example, when $N=1$, the distribution for $\ln X_{i+1}$
-given $\ln X_i$ is a shifted (and reflected) exponential with rate 1.
+particles. For example, when $N=1$, the distribution for $-\ln X_{i+1}$
+given $-\ln X_i$ is a shifted (and reflected) exponential with rate 1.
 \citet{Walter2015} showed how this view of the NS sequence of points
@@
 Consider a particular likelihood value $\ell$ whose corresponding $X$-value
-we would like to know. Since the $\ln X_i$ values have the same distribution
+we would like to know. Since the $-\ln X_i$ values have the same distribution
 as the arrival times of a Poisson process with rate $N$, the probability
@@
 Poisson with expected value $\lambda t$. Therefore the expected number of
-NS discarded points with likelihood below $\ell$ is $-\ln X(\ell)$:
+NS discarded points with likelihood below $\ell$ is $-N\ln X(\ell)$:
 \begin{align}
-\big< n(L(\x_i) \leq \ell) \big> &= -\ln X(\ell).
+\big< n(L(\x_i) \leq \ell) \big> &= -N\ln X(\ell).
 \end{align}
-We can therefore take $n(L(\x_i) \leq \ell)$, the number of points in the
-sequence with likelihood below $\ell$, as an unbiased estimator of
-$-\ln X(\ell)$.
+We can therefore take $n(L(\x_i) \leq \ell)/N$, the number of points in the
+sequence with likelihood below $\ell$ divided by the number of NS particles,
+as an unbiased estimator of $-\ln X(\ell)$.
 
@@
 The above insight, that the number of NS discarded points with likelihood
-below $\ell$ has expected value $-\ln X(\ell)$, is the basis of the algorithm.
-If we wish to measure a log-probability $\ln X(\ell)$, we can use NS to do it.
+below $\ell$ has expected value $-N\ln X(\ell)$, is the basis of the algorithm.
+If we wish to measure a log-probability $-\ln X(\ell)$, we can use NS to do it.
 See Algorithm~\ref{alg:algorithm} for a step-by-step description of the

### 2017-06-19 paper/ms.tex
Minor edits

@@
 
-\section{Example 1: Measuring the mean of a gaussian}
+\section{Example 1: Entropy of a prior for the data}
 
@@
 The true value of the differential entropy $H(\d)$ is available analytically
-in this case, and its value is $146.499$. However, in most situations, the
+in this case, and its value is $146.499$ nats. However, in most situations, the
 entropy of a marginal distribution is not available in closed form.

### 2017-06-20 paper/ms.tex
\xref

@@
 \newcommand{\depth}{(\textnormal{depth})}
+\newcommand{\xref}{\x_{\rm ref}}
 
@@
 
-The idea is to generate a reference particle $\x_{\rm ref}$
+The idea is to generate a reference particle $\xref$
 from the distribution $p(\x)$
@@
 particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$, is evolved
-towards $\x_{\rm ref}$ in order to measure the log-probability near
-$\x_{\rm ref}$.
+towards $\xref$ in order to measure the log-probability near
+$\xref$. Nearness is defined using a distance function
+$d(\x; \xref)$, and the number of NS iterations taken to reduce
+the distance to below some threshold $L$ provides an unbiased estimate of
+\begin{align}
+   -\log \left[ P(d(\x; \xref) < L) \right].
+\end{align}
 
@@
             \Comment{Initialise counter}
-    \State Generate $\x_{\rm ref}$ from $p(\x)$
+    \State Generate $\xref$ from $p(\x)$
             \Comment{Generate a reference point}
@@
             \Comment{Generate initial NS particles}
-    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
+    \State Calculate $d_i \leftarrow d(\x_i; \xref)$ for all $i$
             \Comment{Calculate distance of each particle from the reference
@@
             \Comment{Replace worst particle}
-        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \x_{\rm ref})$
+        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \xref)$
             \Comment{Calculate distance of new particle from reference point}
@@
          without evaluating $\log p(\x)$. Technically, what it actually
-         estimates is $\int f(\x) \int f(\x')
-            \log \left[ P(d(\x', \x) < L | \x) \right]
-                        \, d\x' \, d\x$, the expected value of the
+         estimates is $\int f(\x) \int f(\xref)
+            \log \left[ P(d(\xref, \x) < L | \x) \right]
+                        \, d\xref \, d\x$, the expected value of the
         log-probability of a small region near $\x$.
@@
         with quasi-prior $p(\x)$ and quasi-likelihood given by minus a
-        distance function $d(\x', \x)$.
+        distance function $d(\xref, \x)$.
         \label{alg:algorithm}}
@@
 \begin{align}
-h &= -\int f(\x) \int f(\x')
-            \log \left[ P(d(\x', \x) < L | \x) \right]
-                        \, d\x' \, d\x.
+h &= -\int f(\x) \int f(\xref)
+            \log \left[ P(d(\xref, \x) < L | \x) \right]
+                        \, d\xref \, d\x.
 \end{align}
@@
 \begin{align}
-P(d(\x', \x) < \epsilon | \x)
+P(d(\xref, \x) < \epsilon | \x)
     &\approx
-    f(\x) \int_{d(\x', \x) < L} \, d\x'\\
+    f(\x) \int_{d(\xref, \x) < L} \, d\xref\\
     &= f(\x) \, V
 \end{align}
-where $V$ is the volume of the region where $d(\x', \x) < L$.
+where $V$ is the volume of the region where $d(\xref, \x) < L$.
 
-If the distance function $d(\x', \x)$ is chosen to be Euclidean,
-the constraint $d(\x', \x) < L$ corresponds to a ball in the space
-of possible $\x'$ values.
+If the distance function $d(\xref, \x)$ is chosen to be Euclidean,
+the constraint $d(\xref, \x) < L$ corresponds to a ball in the space
+of possible $\xref$ values.
 The log-volume of a ball of radius $L$ in $n$ dimensions is

### 2017-06-24 paper/ms.tex
Fixed broken reference

@@
 \end{align}
-\citet{skilling}'s key insight was that we can estimate the corresponding
+\citet{skilling2006nested}'s key insight was that we can estimate the corresponding
 $X$-values of the discarded points can be estimated from the nature of the

### 2017-06-26 cpp/Rep.hpp
Should be working for conditional entropy

@@ -141,8 +141,25 @@ void Rep<Particle>::initialise(RNG& temp_rng)
     reference_particle.generate(temp_rng);
 
     // Generate the other particles
-    for(auto& p: particles)
-        p.generate(rng);
+    for(size_t k=0; k<particles.size(); ++k)
+    {
+        if(k == 0)
+            particles[k] = reference_particle;
+        else
+            particles[k] = particles[k-1];
+
+        // Do Metropolis
+        std::cout << "Generating initial particle..." << std::flush;
+        for(int i=0; i<100000; ++i)
+        {
+            auto proposal = particles[k];
+            double logA = proposal.perturb(rng);
+
+            if(rng.rand() <= exp(logA))
+                particles[k] = proposal;
+        }
+        std::cout << "done." << std::endl;
+    }
 
     // Compute all the distances
     compute_distances();

### 2017-06-26 cpp/examples/Sinusoid.cpp

@@ -87,7 +87,7 @@ double Sinusoid::perturb(RNG& rng)
 {
     double logH = 0.0;
 
-    int proposal_type = rng.rand_int(4);
+    int proposal_type = 1;//rng.rand_int(4);
 
     if(proposal_type == 0)
     {

### 2017-06-26 cpp/Execute.hpp
Allow two modes

@@ -20,9 +20,8 @@ void execute(RNG& rng0,
              size_t num_reps,
              size_t num_particles,
              size_t mcmc_steps,
-             const DistFunc<Particle>& dist_func);
-
-
+             const DistFunc<Particle>& dist_func,
+             Mode mode = Mode::standard);
 
 
 /* Implementations below */
@@ -35,7 +34,8 @@ void execute(RNG& rng0,
              size_t num_reps,
              size_t num_particles,
              size_t mcmc_steps,
-             const DistFunc<Particle>& dist_func)
+             const DistFunc<Particle>& dist_func,
+             Mode mode)
 {
     // Clear output file
     std::fstream fout("output.txt", std::ios::out);
@@ -58,7 +58,7 @@ void execute(RNG& rng0,
                                         dist_func);
 
         // Initialise and execute it.
-        rep.initialise(rng0);
+        rep.initialise(rng0, mode);
         rep.execute();
 
         // Print a message.

### 2017-06-26 cpp/Rep.hpp

@@ -15,6 +15,9 @@ namespace InfoNest
 {
 
 
+// Two modes - standard and 'conditional entropy mode'
+enum class Mode { standard, conditional_entropy };
+
 // A convenient name for a distance function
 // This is like a "template typedef"
 template <class Particle>
@@ -90,7 +93,7 @@ class Rep
         ~Rep();
 
         // Initialise all the particles
-        void initialise(RNG& temp_rng);
+        void initialise(RNG& temp_rng, Mode mode);
 
         // Execute the Rep.
         void execute();
@@ -135,7 +138,7 @@ Rep<Particle>::~Rep()
 
 
 template<class Particle>
-void Rep<Particle>::initialise(RNG& temp_rng)
+void Rep<Particle>::initialise(RNG& temp_rng, Mode mode)
 {
     // Generate the reference particle from the distribution
     reference_particle.generate(temp_rng);
@@ -143,22 +146,27 @@ void Rep<Particle>::initialise(RNG& temp_rng)
     // Generate the other particles
     for(size_t k=0; k<particles.size(); ++k)
     {
-        if(k == 0)
-            particles[k] = reference_particle;
-        else
-            particles[k] = particles[k-1];
-
-        // Do Metropolis
-        std::cout << "Generating initial particle..." << std::flush;
-        for(int i=0; i<100000; ++i)
+        if(mode == Mode::standard)
+            particles[k].generate(rng);
+        else if(mode == Mode::conditional_entropy)
         {
-            auto proposal = particles[k];
-            double logA = proposal.perturb(rng);
+            if(k == 0)
+                particles[k] = reference_particle;
+            else
+                particles[k] = particles[k-1];
 
-            if(rng.rand() <= exp(logA))
-                particles[k] = proposal;
+            // Do Metropolis
+            std::cout << "Generating initial particle..." << std::flush;
+            for(int i=0; i<100000; ++i)
+            {
+                auto proposal = particles[k];
+                double logA = proposal.perturb(rng);
+
+                if(rng.rand() <= exp(logA))
+                    particles[k] = proposal;
+            }
+            std::cout << "done." << std::endl;
         }
-        std::cout << "done." << std::endl;
     }
 
     // Compute all the distances

### 2017-06-26 cpp/Rep.hpp
Reminder of hard-coded number of MCMC steps

@@ -165,7 +165,7 @@ void Rep<Particle>::initialise(RNG& temp_rng, Mode mode)
                 if(rng.rand() <= exp(logA))
                     particles[k] = proposal;
             }
-            std::cout << "done." << std::endl;
+            std::cout << "done (1E5 Metropolis steps)." << std::endl;
         }
     }
 

### 2017-06-26 paper/ms.tex
Paragraph about conditional entropy

@@
 
+As the algorithm lets us calculate the entropy of any distribution which can
+be sampled by MCMC, there are several options. The mutual information
+can be written in the following three ways
+\begin{align}
+I(x; y) &= H(x) + H(y) - H(x, y) \\
+I(x; y) &= H(y) - H(y|x) \\
+I(x; y) &= H(x) - H(x|y).
+\end{align}
+In a Bayesian problem where $x$ is an unknown parameter and $y$ is data, the
+first and second formulas would be very costly to compute, because of the
+high dimensional probability distribution for the dataset.
+
 I consider two possible observing strategies, both involving

### 2017-06-26 paper/ms.tex
A bit of writing

@@
 vary with the experimental design. Reference priors
-\citep{bernardo2005reference} also make use of the mutual information but in a
-different way.
+\citep{bernardo2005reference} also maximise $I(\x; \d)$ but vary the prior
+in the maximisation process.
 
@@
 where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
-and $L$ is the likelihood function. It does this evolving a population of
+and $L$ is the likelihood function. While $Z$ is a simple expectation, the
+implied distribution of $L$-values tends to be very heavy-tailed, which is
+why simple Monte Carlo does not work. Equivalently, the integral is dominated
+by very small regions where $L(\x)$ is high.
+
+NS evolves a population of
 $N$ particles in the parameter space.
-The $N$ particles are initially initialised from the prior $\pi(\x)$, and the
+The particles are initially initialised from the prior $\pi(\x)$, and the
 particle with the lowest likelihood, $L^*_1$, is found and its details are
-recorded in the output. This worst particle is then discarded
+recorded as output. This worst particle is then discarded
 and replaced by a new particle
 drawn from the prior $\pi$ but subject to the constraint that its likelihood
-must be above $L^*_1$. This process is repeated, resulting in an
+must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
+and evolving it with MCMC). This process is repeated, resulting in an
 increasing sequence of likelihoods
@@
 If we wish to measure a log-probability $-\ln X(\ell)$, we can use NS to do it.
+$\pi(\x)$ and $L(\x)$ need not be the prior and likelihood respectively, but
+can be any probability distribution (over any space) and any function whose
+expected value is needed for a particular application.
+
 See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
@@
 I(x; y) &= H(y) - H(y|x) \\
-I(x; y) &= H(x) - H(x|y).
+I(x; y) &= H(x) - H(x|y) \\
+I(x; y) &= \int D_{\rm KL}\left(p(x|y) || p(x)\right) \, dy.
 \end{align}

### 2017-06-27 paper/ms.tex
Couple of lines

@@
 \begin{align}
-L^*_1, L^*_2, L^*_3..., 
+L^*_1, L^*_2, L^*_3,..., 
 \end{align}
@@
 This example demonstrates how the algorithm may be used to determine the
-entropy for a quantity whose distribution can only be sampled.
+entropy for a quantity whose distribution can only be sampled, in this case,
+a prior distribution over datasets.
+
 Consider the basic statistics problem of inferring a quantity $\mu$ from
@@
 I(x; y) &= H(x) - H(x|y) \\
-I(x; y) &= \int D_{\rm KL}\left(p(x|y) || p(x)\right) \, dy.
+I(x; y) &= \int p(y) \, D_{\rm KL}\left(p(x|y) \,||\, p(x)\right) \, dy.
 \end{align}
@@
 Thomas Lumley (Auckland),
-Iain Murray (Edinburgh), Jared Tobin (jtobin.io).
+Iain Murray (Edinburgh), Jared Tobin ({\tt jtobin.io}).
 This work was supported by Centre for eResearch

### 2017-06-28 paper/ms.tex
Wrote a paragraph

@@
 be sampled by MCMC, there are several options. The mutual information
-can be written in the following three ways
+can be written in various ways, such as the following.
 \begin{align}
-I(x; y) &= H(x) + H(y) - H(x, y) \\
-I(x; y) &= H(y) - H(y|x) \\
-I(x; y) &= H(x) - H(x|y) \\
-I(x; y) &= \int p(y) \, D_{\rm KL}\left(p(x|y) \,||\, p(x)\right) \, dy.
+I(\x; \d) &= H(\x) + H(\d) - H(\x, \d) \\
+I(\x; \d) &= H(\d) - H(\d|\x) \\
+I(\x; \d) &= H(\x) - H(\x|\d) \\
+I(\x; \d) &= \int p(\d) \, D_{\rm KL}\left(p(\x|\d) \,||\, p(\x)\right) \, d\d.
 \end{align}
-In a Bayesian problem where $x$ is an unknown parameter and $y$ is data, the
-first and second formulas would be very costly to compute, because of the
-high dimensional probability distribution for the dataset.
+In a Bayesian problem where $\x$ is an unknown parameter and $\d$ is data, the
+first and second formulas would be costly to compute, because the
+high dimensional probability distribution for the dataset would require a
+large number of NS iterations to compute $H(\d)$ and $H(\x, \d)$.
+The fourth involves the Kullback-Leibler divergence from the prior to the
+posterior, averaged over all possible datasets. This is straightforward to
+approximate with standard Nested Sampling, but the estimate for a given dataset
+may be biased. This method also would not work if we want $I(\x; \d)$ for
+a single parameter or a subset of them, rather than for all of the parameters.
+Therefore, I adopted the third formula as the best way to compute $I(\x; \d)$.
+In the example, $H(\x)$ is available analytically, and I use the algorithm to
+obtain $H(\x | \d)$.
 

### 2017-06-28 paper/ms.tex
More writing

@@
 In the example, $H(\x)$ is available analytically, and I use the algorithm to
-obtain $H(\x | \d)$.
+obtain $H(\x | \d)$ as follows:
+\begin{align}
+H(\x | \d) &= \int p(\d') H(\x | \d=\d') \, d\d'.
+\end{align}
+This is an expected value over datasets, of the entropy of the posterior given
+each dataset. To do the Nested Sampling to estimate $H(\x | \d=\d')$, I generate
+initial particles from the posterior by doing standard MCMC (targeting the
+posterior) using the $\x$-value that produced the simulated dataset $\d'$ as
+the initial condition for the MCMC.
 

### 2017-06-28 paper/ms.tex
Results

@@
 
+\subsection{Assumptions}
+
 I consider two possible observing strategies, both involving
@@
 as time goes on.
-
-\subsection{Assumptions}
 
@@
 Letting $\tau = \log_{10} T$, and treating
-$A$ and $\phi$ are nuisance parameters, I calculated the depths
-for $\tau$, $\d$, and $(\tau, \d)$.
-These results were computed using a tolerance of $L=0.001$
-for $\tau$, $0.001\sqrt{101}$ for $\d$, and $0.001\sqrt{102}$ for
-the joint run. The results were:
+$A$ and $\phi$ are nuisance parameters, I calculated the conditional entropy
+$H(\tau | \d)$ using the algorithm with a tolerance of $L = 10^{-5}$.
+The result was
 \begin{align}
-\depth_{\tau}    &= 6.216  \pm 0.025 \\
-\depth_{\d}      &= 480.13 \pm 0.32   \\
-\depth_{\tau,\d} &= 479.96 \pm 0.31
+\depth &= 5.379 \pm 0.038
 \end{align}
-
-To convert these to differential entropies, I added the log-volumes
-for 1, 101, and 102 dimensions, which are
-$-6.215$, $-557.252$, and $-562.746$ respectively. This gives
-the following values for the differential entropies:
+Converting to a differential entropy by adding $\ln (10^{-5})$, the conditional
+entropy is
 \begin{align}
-H(\tau)     &=  0.001 \pm 0.025 \label{eqn:h_tau}\\
-H(\d)       &= -77.12 \pm 0.32   \\
-H(\tau, \d) &= -82.79 \pm 0.31   \label{eqn:h_joint}
+H(\tau | \d) &= -5.441 \pm 0.038 \textnormal{ nats}.
 \end{align}
-The mutual information is equal to $H(\tau) + H(\d) - H(\tau, \d)$,
-and can be computed from the results listed
-in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}.
-However, each of these three runs were produced with an identical random
-seed for the reference points, so there is a way to reduce the uncertainty
-by calculating $H(\tau) + H(\d) - H(\tau, \d)$ for each reference pair.
-This is an instance of the
-{\em common random numbers} technique of variance reduction.
-The resulting  mutual information is
-$I(\tau; \d) = 5.67 \pm 0.30$ nats.
+Since the Uniform(0,1) prior has a differential entropy of zero, the
+mutual information is $I(\tau; \d) = H(\tau) - H(\tau | \d) = 5.441 \pm 0.038$
+nats.
 

### 2017-06-28 paper/ms.tex
Results

@@
 
-The results for the uneven observing schedule, again in terms of log
-probabilities, are:
+The results for the uneven observing schedule were
 \begin{align}
-\depth_{\tau}      &= 6.216  \pm 0.025 \\
-\depth_{\d}        &= 479.47 \pm 0.31  \\
-\depth_{\tau, \d}  &= 480.27 \pm 0.32  \\
-I(\tau; \d)        &= 4.69 \pm 0.32
+\depth       &=  5.422 \pm 0.038 \\
+H(\tau | \d) &= -5.398 \pm 0.038 \\
+I(\tau; \d   &=  5.398 \pm 0.038 \textnormal{ nats}.
 \end{align}
@@
 The difference in mutual informations between the two observation
-schedules, assuming a uniform prior on this difference,
-is just $0.98 \pm 0.44$ nats, suggesting (extremely weakly)
-a slight advantage to the
-even observing schedule, in terms of the data's relevance to inference
-of the log-period of the signal. The situation may be different if we
+schedules is trivial.
+The situation may be different if we
 had allowed for shorter periods to be possible, as irregular observing

### 2017-06-28 paper/ms.tex
A couple of minor fixes
Working through from the beginning

@@
 mass functions or densities cannot be evaluated. This paper introduces a
-computational approach to evaluating entropies of probability distributions
-that can only be sampled. The method is based on Nested Sampling,
-and is related to Approximate Bayesian Computation. I demonstrate the method
+computational approach, based on Nested Sampling, to evaluate entropies of
+probability distributions that can only be sampled. I demonstrate the method
 on a simple gaussian example where the key quantities are known analytically,
@@
 if there is only one possible value which has probability 1, $H(x)$ is
-zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
-to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
+zero. Conventionally, $0 \times \log 0$ is defined to be equal
+to $\lim_{x \to 0^+} (x\log x)$, which is zero; i.e., `possibilities' with
+probability zero do not contribute to the sum.
 If there are $N$ possibilities with equal probabilities $1/N$ each,
@@
 `overloaded' notation for probability distributions favoured by many
-Bayesian writers, writing $p(x)$ (for either a probability mass function
-or a probability density function) instead
-of the more cumbersome $P(X=x)$ or $f_X(x)$ where one must always
-distinguish the `random variable' itself ($X$) from a similarly-named
-dummy variable ($x$).
+Bayesian writers \citep{jaynes2003probability, mackay2003information},
+writing $p(x)$ for either a probability mass function
+or a probability density function, instead
+of the more cumbersome $P(X=x)$ or $f_X(x)$.
+In the compact notation, there is no distinction between the
+`random variable' itself ($X$) and a similarly-named dummy variable ($x$).
 Probability distributions are implicitly conditional on some prior
@@
 remember it is not a property of the value of $x$ itself, but a property
-of the probability distribution we use to describe our ignorance of $x$.
+of the probability distribution used to describe a state of ignorance about $x$.
+Throughout this paper, $H(x)$ is used as notation for both Shannon entropies
+(in discrete cases) and differential entropies (in continuous cases). Which
+one it is should be clear from the context of the problem at hand.
 
@@
 user to specify a prior distribution $p(\x)$ for the unknown parameters,
-along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
+along with a conditional distribution $p(\d | \x)$ describing prior knowledge
 of how the data is related to the parameters
@@
 \end{align}
-Converting to a differential entropy by adding $\ln (10^{-5})$, the conditional
+Converting to a differential entropy by adding $\ln (2 \times 10^{-5})$, the conditional
 entropy is
@@
 H(\tau | \d) &= -5.398 \pm 0.038 \\
-I(\tau; \d   &=  5.398 \pm 0.038 \textnormal{ nats}.
+I(\tau; \d)   &=  5.398 \pm 0.038 \textnormal{ nats}.
 \end{align}

### 2017-06-29 paper/ms.tex
More writing

@@
 which is the starting point for
-Bayesian inference \citep{caticha2008lectures}.
+Bayesian inference \citep{caticha2008lectures, caticha2006updating}.
 In practical applications, the following operations
@@
   \item The likelihood, $p(\d | \x)$, can be evaluated cheaply for any
-        $\d$ and $\x$.
+        $\d$ and $\x$. Usually it is the log-likelihood that is actually
+        implemented, for numerical reasons.
 \end{enumerate}
@@
 The former measures the degree to which one question answers
-another (ex ante), so is a function of two questions. The latter measures the
-degree to which a {\em statement} answered a question (ex post),
-and is a function of a question and a statement.
+another {\em ex ante}, and is a function of two questions.
+The latter measures the degree to which a {\em statement} answered a question
+{\em ex post}, and is a function of a question and a statement.
 
 \subsection{Mutual information}
+
 The mutual information is another way of describing the relevance of the
@@
 \end{align}
-It can also be written in terms of the prior, likelihood, and evidence,
-as
+The mutual information can also be written as the expected value
+(with respect to the prior over datasets $p(\d)$) of the Kullback-Leibler
+divergence from prior to posterior:
+\begin{align}
+I(\x ; \d) &= \sum_{\d} p(d) D_{\rm KL}\big(p(\x|\d) \, || \, p(\x)\big).
+\end{align}
+In terms of the prior, likelihood, and evidence,
+it is
 \begin{align}
@@
 log likelihood minus the log evidence.
-As with the conditional entropy, we see that the difficulty appears in the
-form of the log evidence,
+As with the conditional entropy, we see that the computational
+difficulty appears in the form of the log evidence,
 $\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.
@@
 \citep{bernardo2005reference} also maximise $I(\x; \d)$ but vary the prior
-in the maximisation process.
+in the maximisation process while keeping the experimental design fixed.
 
@@
 
-Nested Sampling \citep[NS][]{skilling2006nested} is an algorithm whose aim is
+Nested Sampling \citep{skilling2006nested} is an algorithm whose aim is
 to calculate the evidence
@@
 where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
-and $L$ is the likelihood function. While $Z$ is a simple expectation, the
+and $L$ is the likelihood function. While $Z$ is a simple expectation with
+respect to $\pi$, the
 implied distribution of $L$-values tends to be very heavy-tailed, which is
@@
 
-NS evolves a population of
-$N$ particles in the parameter space.
+To overcome this, NS evolves a population of $N$ particles in the parameter
+space.
 The particles are initially initialised from the prior $\pi(\x)$, and the
@@
 \begin{align}
-L^*_1, L^*_2, L^*_3,..., 
+L^*_1, L^*_2, L^*_3,..., \label{eqn:likelihood_sequence}
 \end{align}
@@
 \end{align}
+we can assign corresponding $X$-values to each discarded point in the sequence,
+and transform $Z$ to a one-dimensional integral.
 \citet{skilling2006nested}'s key insight was that we can estimate the corresponding
-$X$-values of the discarded points can be estimated from the nature of the
+$X$-values of the discarded points from the nature of the
 algorithm. Specifically, each iteration reduces the prior mass by approximately
@@
 $t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$.
+The $t_i$ variables represent the proportion of the remaining prior mass
+that is {\em retained} at each iteration, after imposing the constraint of the
+latest $L^*$ value.
 
@@
 
-The sequence of $-\ln X$ values produced by a Nested Sampling run can be
+Consider $1-t_i$, the fraction of the remaining prior mass that is {\em removed}
+at each iteration.
+The distribution for the compression factors $t_i \sim $Beta$(N,1)$
+corresponds to an exponential distribution for $\ln(1-t_i)$:
+\begin{align}
+\ln (1 - t_i) &\sim \textnormal{Exponential}(N).
+\end{align}
+
+Therefore,
+the sequence of $-\ln X$ values produced by a Nested Sampling run can be
 considered as a Poisson process with rate equal to $N$, the number of
-particles. For example, when $N=1$, the distribution for $-\ln X_{i+1}$
-given $-\ln X_i$ is a shifted (and reflected) exponential with rate 1.
+particles. This is why separate NS runs can be simply merged
+\citep{skilling2006nested}.
 \citet{Walter2015} showed how this view of the NS sequence of points
@@
 a log-probability, which is more relevant to information theoretic
-quantities.
+quantities discussed in the present paper.
 
@@
 as the arrival times of a Poisson process with rate $N$, the probability
-distribution for the number of points in an interval of length $t$ is
-Poisson with expected value $\lambda t$. Therefore the expected number of
+distribution for the number of points in an interval of length $w$ is
+Poisson with expected value $Nw$. Therefore the expected number of
 NS discarded points with likelihood below $\ell$ is $-N\ln X(\ell)$:
@@
 whose entropy is required. Then, a Nested Sampling run evolves a set of
-particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$, is evolved
+particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$,
 towards $\xref$ in order to measure the log-probability near
-$\xref$. Nearness is defined using a distance function
+$\xref$ (see Figure~\ref{fig:fig:algorithm}).
+Nearness is defined using a distance function
 $d(\x; \xref)$, and the number of NS iterations taken to reduce
@@
 \begin{align}
-   -\log \left[ P(d(\x; \xref) < L) \right].
+   \depth = -\log \left[ P(d(\x; \xref) < L) \right]
 \end{align}
+which I call the `depth'. E.g., if $N=10$ Nested Sampling particles are used,
+and it takes 100 iterations to reduce the distance to below $L$, then the
+depth is 10 nats.
+
+If we actually want the differential entropy
+\begin{align}
+H(\x) &= -\int p(\x) \log p(\x) \, d\x
+\end{align}
+we can use the fact that density equals mass divided
+by volume. Assume $L$ is small, so that
+\begin{align}
+P(d(\xref, \x) < \epsilon | \x)
+    &\approx
+    p(\x) \int_{d(\xref, \x) < L} \, d\xref\\
+    &= f(\x) \, V
+\end{align}
+where $V$ is the volume of the region where $d(\xref, \x) < L$.
+Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(L)$.
 
@@
 \end{algorithmic}
-\caption{Calculates $H(\x) = -\int p(\x) \log p(\x) \, d\x$
-         without evaluating $\log p(\x)$. Technically, what it actually
-         estimates is $\int f(\x) \int f(\xref)
+\caption{The algorithm which estimates the `depth',\\
+           \quad\quad$-\int p(\x) \int p(\xref)
             \log \left[ P(d(\xref, \x) < L | \x) \right]
-                        \, d\xref \, d\x$, the expected value of the
-        log-probability of a small region near $\x$.
+                        \, d\xref \, d\x$,\\ minus the expected value of the
+        log-probability of a small region near $\x$, which can be converted
+        to an estimate of an entropy or differential entropy.
         The part highlighted in green is standard Nested Sampling
@@
 
-\subsection{Discrete vs. differential entropies}
-
-Continuous hypothesis spaces are commonly used in practice. Suppose
-$\x$ has probability density $f(\x)$. The algorithm actually estimates
-the `depth', or minus the expected value of the log-probability near $x$:
-\begin{align}
-h &= -\int f(\x) \int f(\xref)
-            \log \left[ P(d(\xref, \x) < L | \x) \right]
-                        \, d\xref \, d\x.
-\end{align}
-for some tolerance $L$.
-
-If we actually want the differential
-entropy
-\begin{align}
-H(\x) &= -\int f(x) \log f(x) \, dx
-\end{align}
-we can use the fact that density equals mass divided
-by volume. Assume $L$ is small, so that
-\begin{align}
-P(d(\xref, \x) < \epsilon | \x)
-    &\approx
-    f(\x) \int_{d(\xref, \x) < L} \, d\xref\\
-    &= f(\x) \, V
-\end{align}
-where $V$ is the volume of the region where $d(\xref, \x) < L$.
-
 If the distance function $d(\xref, \x)$ is chosen to be Euclidean,
@@
 \end{align}
-and provides the conversion factor between the differential entropy and
-the entropy actually computed from the algorithm.
 
-If $L$ is not small, we might think we are calculating the entropy associated
+If the threshold $L$ is not small, we might think we are calculating the entropy associated
 with the question `what is $x$, to within a tolerance of $\pm L$?'.
@@
 type of question.
+
+In the following sections I demonstrate the algorithm on two problems.
 

### 2017-06-29 paper/ms.tex
A bit of appendix writing

@@
 \end{itemize}
-To quantify the degree to which this question is unanswered by
-$\top$, we can write the question as a union of ideal questions:
+According to \citet{knuth_questions}, the entropy of a question is
+computed by applying the sum rule over all statements that would answer the
+question.
+
+We can write the question of interest, $Q$, as a union of ideal questions:
 \begin{align}
@@
 \end{align}
-
 The entropy of this question is

### 2017-07-06 paper/ms.tex
Added paragraph and reference to Szabo

@@
 of the plausible region, when volume is defined with respect to $dx$.
+
+Entropies, and related quantities, tend to be analytically available only for
+a few families of probability distributions. On the numerical side, if
+$\log p(x)$ can be evaluated for any $x$, then simple Monte Carlo will suffice
+for approximating $H(x)$. On the other hand, if $p(x)$ can only be
+{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal
+distribution), then this will not work. Kernel density estimation or the like
+\citep[e.g.][]{JMLR:v15:szabo14a}
+may be effective in this case, but is unlikely to generalise well to high
+dimensions.
+
+This paper introduces a computational approach to evaluating the entropy 
+$H(x)$ of probability distributions that can only be sampled
+(using Markov Chain Monte Carlo).
 

### 2017-07-09 paper/ms.tex
How to implement precisional question

@@
 the version of the algorithm proposed in this paper (when the tolerance is not
-small).
+small). However, the algorithm can be made to estimate the entropy of the
+precisional question by redefining the distance function such that
+$d(\x; \xref)$ can only ever be below $L$ if $\x < \xref$ (or alternatively,
+if $\x > \xref$).
+
+\appendix
+\section{Software}
+
+A {\tt C++} implementation of the algorithm is available in a {\tt git}
+repository located at
+\begin{verbatim}
+https://github.com/eggplantbren/InfoNest
+\end{verbatim}
+and can be obtained using the following {\tt git} command, executed in a
+terminal:
+\begin{verbatim}
+git clone https://github.com/eggplantbren/InfoNest
+\end{verbatim}
+The following will compile the code and execute the first example from the
+paper:
+\begin{verbatim}
+cd InfoNest/cpp
+make
+./main
+\end{verbatim}
+The algorithm will run for 1000 `reps', i.e., 1000 samples of $\xref$, which is
+time consuming. Output is saved to {\tt output.txt}. At any time,
+you can execute the Python script {\tt postprocess.py} to get an estimate of
+the depth:
+\begin{verbatim}
+python postprocess.py
+\end{verbatim}
 

### 2017-07-09 paper/ms.tex
Discuss tolerance

@@
 
+By default, {\tt postprocess.py} estimates the depth with a tolerance of
+$L=10^{-3}$. This value can be changed by calling the {\tt postprocess}
+function with a different value of its argument {\tt tol}. E.g.,
+{\tt postprocess.py} can be edited so that its last line is
+{\tt postprocess(tol=0.01)} instead of {\tt postprocess()}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-09 paper/ms.tex
Pareto section

@@
 
-\subsection{Computational cost}
+\section{Example 3: Data with Pareto distribution}
+
+Consider a situation where the probability distribution for some data values
+$\x = \{x_1, x_2, ..., x_n\}$ is a Pareto distribution, conditional on
+a minimum value $x_{\rm min}=1$ and a slope $\alpha$:
+\begin{align}
+p(\x | \alpha) &= \prod_{i=1}^n \frac{1}{x_i^{\alpha + 1}}
+\end{align}
+where all of the $x_i$ are greater than $x_{\rm min} = 1$.
+If $\alpha$ is relatively low (close to zero) then this distribution is
+very heavy-tailed, and as $\alpha$ increases the probability density
+for the $x$s concentrates just above $x=1$.
+
+With heavy-tailed distributions, it can be difficult to predict future
+observations from past ones, for two reasons. First, the observations may not
+provide much information about the parameters (in this case, $\alpha$), and
+second, even conditional on the parameters, the predictive distribution for
+future data can be heavy-tailed, implying a lot of uncertainty.
+
+\section{Computational cost}
 

### 2017-07-10 paper/ms.tex
More Pareto writing

@@
 second, even conditional on the parameters, the predictive distribution for
-future data can be heavy-tailed, implying a lot of uncertainty.
+future data can be heavy-tailed, implying a lot of uncertainty. However, there
+data clearly provides {\em some} information about the future.
+
+Consider $n=100$ data values, partitioned into two halves, so
+$\boldsymbol{y} = \{x_1, ..., x_{50}\} = \{y_1, ..., y_{50}\}$
+is the first half of the data and
+$\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
+is the second half. Let $\bar{y}$ and $\bar{z}$ be the averages of the two
+subsets of the data:
+\begin{align}
+\bar{y} &= \frac{1}{50} \sum_{i=1}^{50} y_i \\
+\bar{z} &= \frac{1}{50} \sum_{i=1}^{50} z_i.
+\end{align}
+In this section we use the algorithm to answer the questions:
+i) How uncertain is $\bar{y}$?; ii) How uncertain is $\bar{z}$; and
+iii) How informative is $\bar{y}$ about $\bar{z}$? These are quantified by
+$H(\bar{y})$, $H(\bar{z})$, and $I(\bar{y}; \bar{z})$ respectively. This can
+be done despite the fact that $\alpha$ is unknown.
+
+\subsection{}
 

### 2017-07-10 paper/ms.tex
1.5 paragraphs in Pareto section

@@
 
-\subsection{}
+The prior for $\alpha$ was chosen to be lognormal such that the prior median
+of $\alpha$ is 1 and the prior standard deviation of $\ln \alpha$ is also 1.
+Due to the heavy-tailed nature of the Pareto distribution, I ran the
+algorithm in terms of $\ln \bar{y}$ and $\ln \bar{z}$. This affects the
+differential entropies but not the mutual information (which is invariant 
+under changes of parameterisation).
+
+To compute the entropy of $\ln \bar{y}$, the distance function was just the
+absolute value of the distance between the NS particles and the
+reference particle in terms of $\ln \bar{y}$.
 

### 2017-07-10 paper/ms.tex
Bit more work

@@
 Consider the basic statistics problem of inferring a quantity $\mu$ from
-100 observations $\d = \{d_1, ..., d_{100}\}$ whose
+100 observations $\boldsymbol{x} = \{x_1, ..., x_{100}\}$ whose
 probability distribution (conditional on $\mu$) is
 \begin{align}
-p(d_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
+p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
 \end{align}
@@
 \begin{align}
-\mu | \d &\sim \textnormal{Normal}\left(
-                                       \frac{1}{100}\sum_{i=1}^{100} d_i,
+\mu | \boldsymbol{x} &\sim \textnormal{Normal}\left(
+                                       \frac{1}{100}\sum_{i=1}^{100} x_i,
                                        \left[\frac{10}{\sqrt{10001}}\right]^2
@@
 \end{align}
-The true value of the differential entropy $H(\d)$ is available analytically
+The true value of the differential entropy $H(\boldsymbol{x})$ is
+available analytically
 in this case, and its value is $146.499$ nats. However, in most situations, the
 entropy of a marginal distribution is not available in closed form.
+
+To use the algorithm, a distance function $d(,)$ must be defined, to quantify
+how far the NS particles are from the reference particle. To calculate
+the entropy of the data, I used the following distance function:
+\begin{align}
+d\left(\x; \xref\right) &= \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2
+\end{align}
 
@@
 \begin{align}
-\depth_{\d} &= 698.25 \pm 0.34.
+\depth_{\boldsymbol{x}} &= 698.25 \pm 0.34.
 \end{align}
@@
 \begin{align}
-H(\d) &\approx 146.49 \pm 0.34.
+H(\boldsymbol{x}) &\approx 146.49 \pm 0.34.
 \end{align}

### 2017-07-10 paper/ms.tex
Writing about proposals

@@
 \end{align}
-which is very close to the true value (suspiciously close, but this
+which is very close to the true value (perhaps suspiciously close, but this
 was just a fluke).
+
+To carry out the sampling for this problem, Metropolis-Hastings moves that
+explore the joint prior distribution $p(\mu, \boldsymbol{x})$ had to be
+implemented. Four basic kinds of proposals were used: i) those which change
+$\mu$ while leaving $\boldsymbol{x}$ fixed; ii) those which change $\mu$ and
+shift $\boldsymbol{x}$ correspondingly; iii) those which resample a subset
+of the $x$s from $p(\boldsymbol{x}|\mu)$; and iv) those which move a single
+$x$ slightly.
 
@@
 \end{figure}
+
+To carry out the sampling for this problem, Metropolis-Hastings moves that
+explore the joint prior distribution $p(A, \tau, \phi, \boldsymbol{d})$
+had to be
+implemented. As with the first example, there were four basic kinds of
+proposals. Four basic kinds of proposals were used: i) those which change
+one of the parameters
+$A$, $\tau$ or $\phi$ while leaving $\boldsymbol{x}$ fixed
+(these are standard Metropolis moves targeting the posterior distribution);
+ii) those which change the parameters and modify the data accordingly;
+iii) those which resample a subset
+of the data; and iv) those which move a single datum slightly.
 

### 2017-07-11 paper/ms.tex
Pareto stuffs

@@
 \newcommand{\xref}{\x_{\rm ref}}
+\newcommand{\ytot}{y_{\rm tot}}
+\newcommand{\ztot}{z_{\rm tot}}
 
@@
 $\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
-is the second half. Let $\bar{y}$ and $\bar{z}$ be the averages of the two
+is the second half. Let $\ytot$ and $\ztot$ be the averages of the two
 subsets of the data:
 \begin{align}
-\bar{y} &= \frac{1}{50} \sum_{i=1}^{50} y_i \\
-\bar{z} &= \frac{1}{50} \sum_{i=1}^{50} z_i.
+\ytot &= \sum_{i=1}^{50} y_i \\
+\ztot &= \sum_{i=1}^{50} z_i.
 \end{align}
 In this section we use the algorithm to answer the questions:
-i) How uncertain is $\bar{y}$?; ii) How uncertain is $\bar{z}$; and
-iii) How informative is $\bar{y}$ about $\bar{z}$? These are quantified by
-$H(\bar{y})$, $H(\bar{z})$, and $I(\bar{y}; \bar{z})$ respectively. This can
+i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
+iii) How informative is $\ytot$ about $\ztot$? These are quantified by
+$H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
 be done despite the fact that $\alpha$ is unknown.
@@
 Due to the heavy-tailed nature of the Pareto distribution, I ran the
-algorithm in terms of $\ln \bar{y}$ and $\ln \bar{z}$. This affects the
+algorithm in terms of $\ln \ytot$ and $\ln \ztot$. This affects the
 differential entropies but not the mutual information (which is invariant 
@@
 
-To compute the entropy of $\ln \bar{y}$, the distance function was just the
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/pareto.pdf}
+\caption{The joint distribution for the log-sum of the first half of the
+`Pareto data' and the second half. The goal is to calculate the entropy of
+the marginal distributions, the entropy of the joint distribution, and
+hence the mutual information.\label{fig:pareto}}
+\end{figure}
+
+To compute the entropy of $\ln \ytot$, the distance function was just the
 absolute value of the distance between the NS particles and the
-reference particle in terms of $\ln \bar{y}$.
+reference particle in terms of $\ln \ytot$.
 

### 2017-07-11 paper/figures/pareto.py

@@ -1,0 +1,27 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import numpy.random as rng
+
+N = 100
+
+rng.seed(0)
+for i in range(0, 2000):
+    alpha = np.exp(rng.randn())
+    us = rng.rand(N)
+    xs = (1.0 - us)**(-1.0 / alpha)
+
+    first_half = xs[0:(N//2)]
+    second_half = xs[(N//2):]
+
+    plt.plot(np.log(first_half.mean()), np.log(second_half.mean()), "k.",
+             markersize=2, alpha=0.1)
+    print(i+1)
+
+plt.axis("square")
+plt.xlim([0.0, 10.0])
+plt.ylim([0.0, 10.0])
+plt.xlabel(r"$\ln(y_{\rm tot})$")
+plt.ylabel(r"$\ln(z_{\rm tot})$")
+plt.savefig("pareto.pdf", bbox_inches="tight")
+plt.show()
+

### 2017-07-11 paper/ms.tex
Cover and Thomas
Writing about Pareto results

@@
 If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
-the Shannon entropy \citep{shannon} is defined as
+the Shannon entropy \citep{shannon, cover2012elements} is defined as
 \begin{align}
@@
 $H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
-be done despite the fact that $\alpha$ is unknown.
+be done despite the fact that $\alpha$ is unknown. The joint distribution
+for $\ln \ytot$ and $\ln \ztot$, whose properties I am to obtain,
+is shown in Figure~\ref{fig:pareto}.
 
@@
 the marginal distributions, the entropy of the joint distribution, and
-hence the mutual information.\label{fig:pareto}}
+hence the mutual information. The distribution is quite heavy tailed despite
+the logarithms, and extends beyond the domain shown in this plot.
+\label{fig:pareto}}
 \end{figure}
+
+Three types of Metropolis-Hastings proposals were used for this example:
+i) A proposal which changes $\alpha$ and modifies all of the $x$s accordingly;
+ii) A proposal which resamples a subset of the $x$s from the Pareto
+distribution; and iii) A proposal which moves a single $x$-value slightly.
+A fourth proposal which moves $\alpha$ while keeping $\boldsymbol{x}$ fixed,
+i.e., a posterior-sampling move, was not implemented.
 
@@
 absolute value of the distance between the NS particles and the
-reference particle in terms of $\ln \ytot$.
+reference particle in terms of $\ln \ytot$. The result, using a tolerance
+of $L = 10^{-3}$, was:
+\begin{align}
+\depth_{\ln \ytot} &= 8.851 \pm 0.057.
+\end{align}
+Correcting for the volume factor yields a differential entropy estimate
+of $H(\ytot) = 2.366\pm 0.057$ nats.
+By the symmetry of the problem specification, $H(\ztot) = H(\ytot)$.
 

### 2017-07-11 paper/ms.tex
Taleb ref

@@
 Consider a situation where the probability distribution for some data values
-$\x = \{x_1, x_2, ..., x_n\}$ is a Pareto distribution, conditional on
+$\boldsymbol{x} = \{x_1, x_2, ..., x_n\}$ is a Pareto distribution, conditional on
 a minimum value $x_{\rm min}=1$ and a slope $\alpha$:
 \begin{align}
-p(\x | \alpha) &= \prod_{i=1}^n \frac{1}{x_i^{\alpha + 1}}
+p(\boldsymbol{x} | \alpha) &= \prod_{i=1}^n \frac{1}{x_i^{\alpha + 1}}
 \end{align}
@@
 With heavy-tailed distributions, it can be difficult to predict future
-observations from past ones, for two reasons. First, the observations may not
-provide much information about the parameters (in this case, $\alpha$), and
-second, even conditional on the parameters, the predictive distribution for
-future data can be heavy-tailed, implying a lot of uncertainty. However, there
-data clearly provides {\em some} information about the future.
+observations from past ones, a notion popularised by
+\citet{taleb2007black}. This occurs for three main reasons.
+First, the observations may not
+provide much information about the parameters (in this case, $\alpha$).
+Second, even conditional on the parameters, the predictive distribution for
+future data can be heavy-tailed, implying a lot of uncertainty.
+Finally, one may always doubt the probability assignment on which such a
+calculation is based.
+However, the data clearly provides {\em some} information about the future.
 
@@
 
+The joint entropy $H(\ln \ytot, \ln \ztot)$ was also estimated by using
+a Euclidean distance function and a tolerance of $10^{-3}$. The result was
+\begin{align}
+\depth_{(\ln \ytot, \ln \ztot)} &= 16.03 \pm 0.11.
+\end{align}
+Applying the log-volume correction yields a differential entropy estimate of
+$H(\ln \ytot, \ln \ztot) = 3.40 \pm 0.11$ nats.
+Therefore, the mutual information is
+\begin{align}
+I(\ytot; \ztot) &= H(\ytot) + H(\ztot) - H(\ytot, \ztot)\\
+                &= 1.130 \pm 0.082 \textnormal{ nats}.
+\end{align}
+The variance of this mutual information estimate was reduced by
+using a common sequence of reference points for the marginal and joint entropy
+runs, an instance of the `common random numbers'
+technique\footnote{\tt http://demonstrations.wolfram.com/TheMethodOfCommonRandomNumbersAnExample/}.
+
 \section{Computational cost}
@@
 is more efficient than running the two strategies separately.
-
-\section{Example 3: TBD}
 

### 2017-07-11 paper/ms.tex
averages -> totals
Appendix references

@@
 towards $\xref$ in order to measure the log-probability near
-$\xref$ (see Figure~\ref{fig:fig:algorithm}).
+$\xref$ (see Figure~\ref{fig:algorithm}).
 Nearness is defined using a distance function
@@
 with the question `what is $x$, to within a tolerance of $\pm L$?'.
-This is not quite correct. See the Appendix for a discussion of this
+This is not quite correct. See appendix~\ref{sec:precisional}
+for a discussion of this
 type of question.
@@
 ``what is the value of $T$ to within $\pm$ 10\%'', or something along these
-lines.
+lines. See Appendix~\ref{sec:precisional} for more information about this
+issue.
 
@@
 $\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
-is the second half. Let $\ytot$ and $\ztot$ be the averages of the two
+is the second half. Let $\ytot$ and $\ztot$ be the totals of the two
 subsets of the data:
@@
 \section{Precisional questions}
+\label{sec:precisional}
 
@@
 \section{Software}
+\label{sec:software}
 

### 2017-07-11 paper/ms.tex
Random work

@@
 probability distributions that can only be sampled. I demonstrate the method
-on a simple gaussian example where the key quantities are known analytically,
-and an experimental design example about scheduling observations in order
-to measure the period of an oscillating signal.}
+on three examples: a simple gaussian example where the key quantities are
+available analytically; (ii)
+an experimental design example about scheduling observations in order
+to measure the period of an oscillating signal; and (iii) an example based
+on a heavy-tailed prior distribution.}
 
@@
 one it is should be clear from the context of the problem at hand.
+I sometimes write general formulae are written in terms of sums (i.e., as they would
+appear in a discrete problem), and sometimes as integrals, as they would
+appear in a continuous problem.
 
@@
 
-Suppose we wanted to compute $H(\x | \d)$, perhaps to compare it to
+One might wish to compute $H(\x | \d)$, perhaps to compare it to
 $H(\x)$ and quantify how much might be learned about $\x$.
@@
 \end{align}
-contains the marginal likelihood integral:
-\begin{align}
-p(\d) &= \int p(\x) p(\d | \x) \, d\x
-\end{align}
-also known as the `evidence', which tends to be computable but costly.
+contains the marginal likelihood:
+\begin{align}
+p(\d) &= \sum_{\x} p(\x) p(\d | \x) \, d\x
+\end{align}
+also known as the `evidence', which tends to be computable but costly,
+especially when the summation is replaced by an integration over a
+high-dimensional parameter space.
 
@@
 log likelihood minus the log evidence.
-As with the conditional entropy, we see that the computational
+As with the conditional entropy, the computational
 difficulty appears in the form of the log evidence,
-$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.
-
-For experimental design purposes, if we want to maximise the expected amount
-of information obtained from data, maximising either $I(\x; \d)$ or
-$H(\x | \d)$ will produce the same result because the prior $p(\x)$ does not
+$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$, which must be evaluated
+or estimated for many possible datasets.
+
+For experimental design purposes, maximising the expected amount
+of information obtained from data is a sensible goal.
+Formally, either maximising $I(\x; \d)$ or
+minimising $H(\x | \d)$ will produce the same result because the prior $p(\x)$
+does not
 vary with the experimental design. Reference priors
@@
 
-Nested Sampling \citep{skilling2006nested} is an algorithm whose aim is
+Nested Sampling (NS), introduced by
+\citet{skilling2006nested} is an algorithm whose aim is
 to calculate the evidence
@@
 where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
-and $L$ is the likelihood function. While $Z$ is a simple expectation with
+and $L$ is the likelihood function. The original NS algorithm, and variations
+within the same family, \citep{feroz2009multinest, dns, handley2015polychord},
+have become popular tools in Bayesian data analysis
+\citep{pullen2014bayesian, exoplanet} and have also
+found use in statistical physics \citep{partay2010efficient,
+baldock2016determining, martiniani2014superposition}, which was another of
+its aims.
+
+While $Z$ is a simple expectation with
 respect to $\pi$, the
@@
 \end{align}
-we can assign corresponding $X$-values to each discarded point in the sequence,
-and transform $Z$ to a one-dimensional integral.
-\citet{skilling2006nested}'s key insight was that we can estimate the corresponding
-$X$-values of the discarded points from the nature of the
+each discarded point in the sequence can be assigned an $X$-value,
+which transforms $Z$ to a one-dimensional integral.
+\citet{skilling2006nested}'s key insight was that the
+$X$-values of the discarded points are approximately known from the nature of the
 algorithm. Specifically, each iteration reduces the prior mass by approximately
@@
 probability distribution at the red point, Nested Sampling can be used,
-with the blue distribution plays the role of the ``prior'' in NS, and the
+with the blue distribution playing the role of the ``prior'' in NS, and the
 Euclidian distance from the red point (illustrated with red contours)
-is the ``likelihood''.\label{fig:algorithm}}
+playing the role of the negative log-likelihood. Averaging over selections
+of the red point gives an estimate of the entropy of the blue distribution.
+\label{fig:algorithm}}
 \end{figure}
@@
 \end{align}
-In this section we use the algorithm to answer the questions:
+In this section I use the algorithm to answer the questions:
 i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
@@
 
-\section{Computational cost}
+\section{Computational cost, limitations, and conclusion}
 
@@
 \end{align}
-It should not be surprising that this is costly.
-
-In the sinusoidal signal example, I calculated the difference in
-mutual information between two observing strategies. It is possible
-that there exists a more direct method of obtaining this which
-is more efficient than running the two strategies separately.
+It should not be surprising that this is costly. Indeed, it seems possible
+to me that any {\em general} algorithm for computing entropies will have the
+property that NS (or a similar algorithm capable of measuring small volumes)
+appears in an inner loop, since entropies are sums of terms, each of which
+relates to the downset of a given statement --- the downset of a statement
+$S$ being the set of all statements that imply S \citep{knuth_questions}.
 
@@
 It is a pleasure to thank the following people for interesting and helpful
-conversations about this topic: Ruth Angus (Flatiron Institute),
-Ewan Cameron (Oxford), James Curran (Auckland),
+conversations about this topic, and/or comments on a draft:
+Ruth Angus (Flatiron Institute),
+Ewan Cameron (Oxford), James Curran (Auckland), Tom Elliott (Auckland),
 David Hogg (NYU), Kevin Knuth (SUNY Albany),

### 2017-07-11 paper/ms.tex
An attempt at concluding

@@
 
+The algorithm as presented will tend to become less accurate on high dimensional
+problems, as the required depth of the runs will increase correspondingly,
+since many NS iterations will be required to get a high dimensional vector
+of quantities to be close to the reference vector.
+The accuracy of the $\ln\depth$ estimates should degrade in proportion to
+$\sqrt{\depth}$, due to the Poisson-process nature of the $-\ln X$ sequence.
+
+Throughout this paper, the question of how to sample from the constrained
+distributions required by NS has not been discussed in detail. For the examples,
+I used the original \citet{skilling} suggestion of using a short MCMC run
+initialised at a surviving point. In some problems, this will not be effective,
+and this will cause the resulting estimates to be incorrect, as in standard
+Nested Sampling applications. Despite this, the algorithm should still be
+useful in a wide range of applications.
+
 \acknowledgments{Acknowledgements}

### 2017-07-11 paper/ms.tex
Editing

@@
 sequence with likelihood below $\ell$ divided by the number of NS particles,
-as an unbiased estimator of $-\ln X(\ell)$.
+as an unbiased estimator of $-\ln X(\ell)$. The standard deviation of the
+estimator is $\sqrt{-\ln X(\ell)/N}$.
 
@@
 \end{algorithmic}
-\caption{The algorithm which estimates the `depth',\\
+\caption{The algorithm which estimates the expected value of the `depth',\\
            \quad\quad$-\int p(\x) \int p(\xref)
             \log \left[ P(d(\xref, \x) < L | \x) \right]
-                        \, d\xref \, d\x$,\\ minus the expected value of the
+                        \, d\xref \, d\x$,\\ that is, minus the expected value of the
         log-probability of a small region near $\x$, which can be converted
@@
 
-In the following sections I demonstrate the algorithm on two problems.
+In the following sections I demonstrate the algorithm on three
+example problems.
 
@@
 
-I ran the algorithm and computed the depth using a tolerance of
+I ran the algorithm and computed the average depth using a tolerance of
 $L=10^{-3}\sqrt{100}$, so that the RMS difference between
 points in the two datasets was about $10^{-3}$.
-From the algorithm, the average depth was
-\begin{align}
-\depth_{\boldsymbol{x}} &= 698.25 \pm 0.34.
-\end{align}
+From the algorithm, the estimate of the expected depth was
+\begin{align}
+\left<\depth_{\boldsymbol{x}}\right> &= 698.25 \pm 0.34.
+\end{align}
+The uncertainty here is the standard error of the mean, i.e., the
+standard deviation of the depth estimates divided by square root of
+the number of NS repetitions.
+
 The log-volume of a 100-dimensional ball of radius
@@
 \begin{align}
-\depth &= 5.379 \pm 0.038
+\left<\depth\right> &= 5.379 \pm 0.038
 \end{align}
@@
 \begin{align}
-\depth       &=  5.422 \pm 0.038 \\
+\left<\depth\right>       &=  5.422 \pm 0.038 \\
 H(\tau | \d) &= -5.398 \pm 0.038 \\
@@
 \begin{align}
-\depth_{\ln \ytot} &= 8.851 \pm 0.057.
+\left<\depth_{\ln \ytot}\right> &= 8.851 \pm 0.057.
 \end{align}
@@
 \begin{align}
-\depth_{(\ln \ytot, \ln \ztot)} &= 16.03 \pm 0.11.
+\left<\depth_{(\ln \ytot, \ln \ztot)}\right> &= 16.03 \pm 0.11.
 \end{align}

### 2017-07-11 paper/ms.tex
Cleaning up notation

@@
 \begin{align}
-d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
+Y_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
 \end{align}
@@
 To carry out the sampling for this problem, Metropolis-Hastings moves that
-explore the joint prior distribution $p(A, \tau, \phi, \boldsymbol{d})$
+explore the joint prior distribution $p(A, \tau, \phi, \boldsymbol{Y})$
 had to be
@@
 one of the parameters
-$A$, $\tau$ or $\phi$ while leaving $\boldsymbol{x}$ fixed
+$A$, $\tau$ or $\phi$ while leaving the data $\boldsymbol{Y}$ fixed
 (these are standard Metropolis moves targeting the posterior distribution);
@@
 $A$ and $\phi$ are nuisance parameters, I calculated the conditional entropy
-$H(\tau | \d)$ using the algorithm with a tolerance of $L = 10^{-5}$.
+$H(\tau | \boldsymbol{Y})$ using the algorithm with a tolerance of $L = 10^{-5}$.
 The result was
@@
 \begin{align}
-H(\tau | \d) &= -5.441 \pm 0.038 \textnormal{ nats}.
+H(\tau | \boldsymbol{Y}) &= -5.441 \pm 0.038 \textnormal{ nats}.
 \end{align}
 Since the Uniform(0,1) prior has a differential entropy of zero, the
-mutual information is $I(\tau; \d) = H(\tau) - H(\tau | \d) = 5.441 \pm 0.038$
+mutual information is $I(\tau; \boldsymbol{Y}) = H(\tau) - H(\tau | \boldsymbol{Y}) = 5.441 \pm 0.038$
 nats.
@@
 \left<\depth\right>       &=  5.422 \pm 0.038 \\
-H(\tau | \d) &= -5.398 \pm 0.038 \\
-I(\tau; \d)   &=  5.398 \pm 0.038 \textnormal{ nats}.
+H(\tau | \boldsymbol{Y}) &= -5.398 \pm 0.038 \\
+I(\tau; \boldsymbol{Y})   &=  5.398 \pm 0.038 \textnormal{ nats}.
 \end{align}

### 2017-07-11 paper/ms.tex
NS parameters

@@
 In the following sections I demonstrate the algorithm on three
-example problems.
+example problems. For each of the examples, I did 1000 repetitions of NS
+(i.e., 1000 reference particles), and used 10 NS particles and 10000 MCMC
+steps per NS iteration.
 
@@
 distributions required by NS has not been discussed in detail. For the examples,
-I used the original \citet{skilling} suggestion of using a short MCMC run
+I used the original \citet{skilling2006nested} suggestion of using a short MCMC run
 initialised at a surviving point. In some problems, this will not be effective,

### 2017-07-11 paper/ms.tex
Minor edit
Tiny bit of writing in the second appendix

@@
 and this will cause the resulting estimates to be incorrect, as in standard
-Nested Sampling applications. Despite this, the algorithm should still be
+Nested Sampling applications. Despite this, the algorithm should be
 useful in a wide range of applications.
@@
 
+The numerical parameters, and the choice of which problem is being solved,
+are specified in {\tt main.cpp}. The default problem is the first example
+from the paper.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-11 cpp/main.cpp

@@ -14,6 +14,7 @@ int main()
 
     // Create random number generators
     // The first one is used to generate reference points
+    // and the second is used for the Nested Sampling
     unsigned long seed0 = 0;
     unsigned long seed1 = time(0);
     InfoNest::RNG rng0(seed0);

### 2017-07-11 prototype/algorithm.py
Removed old stuff

@@ -1,130 +1,0 @@
-# Stuff accessible from outside (?)
-__all__ = ["Run"]
-
-# Imports
-from copy import deepcopy
-import numpy as np
-import numpy.random as rng
-from particle import Particle, distance
-
-
-
-
-class Run:
-    """
-    An object of this class represents a single run of the algorithm.
-    Ultimately, you'll need to do many runs and average the results.
-    """
-
-    def __init__(self, num_particles=1, mcmc_steps=1000):
-        """
-        Constructor. Pass in the desired number of NS particles
-        and the number of mcmc steps per iteration.
-        """
-        self.num_particles = num_particles
-        self.mcmc_steps = mcmc_steps
-
-    def initialise(self):
-        """
-        Generate the reference particle and the comparison particles.
-        """
-
-        self.reference_particle = Particle.factory()
-        self.particles = [ Particle.factory()
-                                        for i in range(0, self.num_particles) ]
-        self.compute_distances()
-        self.iteration = 1
-
-
-    def compute_distances(self):
-        """
-        Compute the distance from the reference particle to each
-        comparison particle.
-        """
-        self.distances = [ distance(self.reference_particle, p)
-                                for p in self.particles ]
-        self.distances = np.array(self.distances)
-
-    def iterate(self):
-        """
-        Find the worst particle and replace it, NS style.
-        """
-
-        # Index of worst particle
-        worst = np.argmax(self.distances)
-
-        # Print iteration, and worst particle's distance
-        s = str(self.iteration) + " " + str(self.distances[worst])
-
-        f = open("output.txt", "a")
-        f.write(s + "\n")
-        f.close()
-        print(s)
-
-        # Replace worst particle.
-        self.replace(worst)
-
-        # Update iteration counter.
-        self.iteration += 1
-
-
-    def replace(self, worst):
-        """
-        Replace the worst particle.
-        """
-        # Threshold
-        threshold = self.distances[worst]
-
-        # Copy a survivor
-        if self.num_particles > 1:
-            while True:
-                copy = rng.randint(self.num_particles)
-                if copy != worst:
-                    break
-            self.particles[worst] = deepcopy(self.particles[copy])
-
-        # Do mcmc
-        for i in range(0, self.mcmc_steps):
-            # Generate proposal
-            proposal = deepcopy(self.particles[worst])
-            logA = proposal.perturb()
[…]

### 2017-07-11 prototype/particle.py

@@ -1,75 +1,0 @@
-# Stuff accessible from outside (?)
-__all__ = ["Particle"]
-
-# Imports
-import numpy as np
-import numpy.random as rng
-from utils import randh, wrap
-
-class Particle:
-    """
-    An object of this class represents a point in the relevant
-    hypothesis space.
-    """
-
-    def __init__(self):
-        """
-        Constructor does nothing.
-        """
-        pass
-
-
-    def generate(self):
-        """
-        Generate the point from its probability distribution.
-        """
-        self.x = rng.randn()
-
-
-    def perturb(self):
-        """
-        Metropolis proposal that would explore the probability distribution.
-        Return ln(a) such that acceptance probability alpha = min(1, a).
-        """
-        logA = 0.0
-        logA -= -0.5 * self.x**2
-        self.x += 1.0 * randh()
-        logA += -0.5 * self.x**2
-        return logA
-
-    def __str__(self):
-        """
-        Represent as a string.
-        """
-        return str(self.x)
-
-
-    def factory():
-        """
-        Class-level function. Returns a `generated` particle, ready for use.
-        """
-        particle = Particle()
-        particle.generate()
-        return particle
-
-
-def distance(particle1, particle2):
-    """
-    Distance between two particles.
-    """
-    return np.abs(particle1.x - particle2.x)
-
-
-
-
-if __name__ == "__main__":
-    """
-    A few tiny little tests.
-    """
-
-    particle = Particle.factory()
-
-    for i in range(0, 1000):
-        particle.perturb()
-        print(particle)
-

### 2017-07-11 prototype/postprocess.py

@@ -1,45 +1,0 @@
-__all__ = ["postprocess"]
-
-# Imports
-import numpy as np
-import matplotlib.pyplot as plt
-
-def postprocess(num_particles=1, tol=1E-3):
-    """
-    Load the output file and estimate the entropy.
-    """
-
-    # Load output file
-    output = np.loadtxt("output.txt")
-    num_particles = np.loadtxt("num_particles.txt")
-
-    # Find the starts of the runs
-    start_indices = np.nonzero(output[:,0] == 1)[0]
-
-    reps = len(start_indices)
-    counts = np.empty(reps)
-    for i in range(0, reps):
-        start_index = start_indices[i]
-        if i < reps - 1:
-            run = output[start_index : start_indices[i+1], :]
-        else:
-            run = output[start_index : , :]
-        print("Processed run {k}/{N}.".format(k=i+1, N=reps))
-        counts[i] = (run[:,1] > tol).sum()
-
-    print("mean(counts) = {mc}.".format(mc=counts.mean()))
-    print("std (counts) = {sd}.".format(sd=counts.std()))
-
-    # Estimate of the differential entropy
-    # log(probability) ~= log(2 * tol * density)
-    H = counts.mean() / num_particles + np.log(2 * tol)
-    stderr = (counts.std() / num_particles) / np.sqrt(reps)
-
-    print("H = {H} +- {stderr} nats.".format(H=H, stderr=stderr))
-
-#    plt.hist(mean_depth, 100, color=[0.2, 0.2, 0.2])
-#    plt.show()
-
-if __name__ == "__main__":
-    postprocess()
-

### 2017-07-11 prototype/utils.py

@@ -1,18 +1,0 @@
-__all__ = ["randh", "wrap"]
-
-import numpy as np
-import numpy.random as rng
-
-def randh(N=1):
-    """
-    Generate from the heavy-tailed distribution.
-    """
-    if N==1:
-	    return 10.0**(1.5 - 3*np.abs(rng.randn()/np.sqrt(-np.log(rng.rand()))))*rng.randn()
-    return 10.0**(1.5 - 3*np.abs(rng.randn(N)/np.sqrt(-np.log(rng.rand(N)))))*rng.randn(N)
-
-
-def wrap(x, a, b):
-    assert b > a
-    return (x - a)%(b - a) + a
-

### 2017-07-11 paper/ms.tex
Explain h
Fixed equations

@@
 \end{algorithmic}
-\caption{The algorithm which estimates the expected value of the `depth',\\
-           \quad\quad$-\int p(\x) \int p(\xref)
-            \log \left[ P(d(\xref, \x) < L | \x) \right]
-                        \, d\xref \, d\x$,\\ that is, minus the expected value of the
-        log-probability of a small region near $\x$, which can be converted
+\caption{The algorithm which estimates the expected value of the depth:\\
+           \quad\quad$-\int p(\xref) \int p(\x)
+            \log \left[ P(d(\x; \xref) < L | \xref) \right]
+                        \, d\x \, d\xref$,\\ that is, minus the expected value of the
+        log-probability of a small region near $\xref$, which can be converted
         to an estimate of an entropy or differential entropy.
@@
         with quasi-prior $p(\x)$ and quasi-likelihood given by minus a
-        distance function $d(\xref, \x)$.
+        distance function $d(\x; \xref)$. The final result, $\widehat{h_{\rm final}}$, is an estimate of the expected depth.
         \label{alg:algorithm}}

### 2017-07-11 paper/ms.tex
Malakar ref

@@
 % Authors (Add full first names)
-\Author{Brendon J. Brewer$^{1,}$*}
+\Author{Brendon J. Brewer}
 
@@
 \address{
-$^{1}$ Department of Statistics, The University of Auckland, Private Bag 92019,
+Department of Statistics, The University of Auckland, Private Bag 92019,
 Auckland 1142, New Zealand}
@@
 have become popular tools in Bayesian data analysis
-\citep{pullen2014bayesian, exoplanet} and have also
+\citep{knuth2015bayesian, pullen2014bayesian, exoplanet} and have also
 found use in statistical physics \citep{partay2010efficient,

### 2017-07-11 paper/ms.tex
Editing the conclusion

@@
 and this will cause the resulting estimates to be incorrect, as in standard
-Nested Sampling applications. Despite this, the algorithm should be
+Nested Sampling applications. Testing the sensitivity of the results to
+changes in the number of MCMC steps used is a simple way to increase one's
+confidence in the results, but is not a guarantee.
+Despite this weakness, I hope the algorithm will be
 useful in a wide range of applications.

### 2017-07-12 paper/ms.tex
minor edit in appendix
Half of corrections

@@
 an experimental design example about scheduling observations in order
-to measure the period of an oscillating signal; and (iii) an example based
-on a heavy-tailed prior distribution.}
+to measure the period of an oscillating signal; and (iii) predicting the
+future from the past in a heavy-tailed scenario.}
 
@@
 $H(x)$ of probability distributions that can only be sampled
-(using Markov Chain Monte Carlo).
+using Markov Chain Monte Carlo (MCMC).
 
@@
 \begin{align}
-p(\d) &= \sum_{\x} p(\x) p(\d | \x) \, d\x
+p(\d) &= \sum_{\x} p(\x) p(\d | \x)
 \end{align}
@@
 \begin{align}
-I(\x ; \d) &= \sum_{\d} p(d) D_{\rm KL}\big(p(\x|\d) \, || \, p(\x)\big).
+I(\x ; \d) &= \sum_{\d} p(\d) D_{\rm KL}\big(p(\x|\d) \, || \, p(\x)\big).
 \end{align}
@@
 difficulty appears in the form of the log evidence,
-$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$, which must be evaluated
+$\log p(\d) = \log \left(\sum_{\x} p(\x)p(\d | \x)\right)$, which must be evaluated
 or estimated for many possible datasets.
@@
 For experimental design purposes, maximising the expected amount
-of information obtained from data is a sensible goal.
+of information obtained from the data is a sensible goal.
 Formally, either maximising $I(\x; \d)$ or
@@
 \citep{bernardo2005reference} also maximise $I(\x; \d)$ but vary the prior
+$p(\x)$
 in the maximisation process while keeping the experimental design fixed.
@@
 Nested Sampling (NS), introduced by
-\citet{skilling2006nested} is an algorithm whose aim is
+\citet{skilling2006nested}, is an algorithm whose aim is
 to calculate the evidence
@@
 and $L$ is the likelihood function. The original NS algorithm, and variations
-within the same family, \citep{feroz2009multinest, dns, handley2015polychord},
+within the same family \citep{feroz2009multinest, dns, handley2015polychord},
 have become popular tools in Bayesian data analysis
@@
 To overcome this, NS evolves a population of $N$ particles in the parameter
-space.
+space towards higher likelihood regions.
 The particles are initially initialised from the prior $\pi(\x)$, and the
@@
 must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
-and evolving it with MCMC). This process is repeated, resulting in an
+and evolving it with MCMC to explore
+$\pi(\x)\mathds{1}\left(L(\x) > L^*_1\right)$).
+This process is repeated, resulting in an
 increasing sequence of likelihoods
 \begin{align}
-L^*_1, L^*_2, L^*_3,..., \label{eqn:likelihood_sequence}
-\end{align}
-
+L^*_1, L^*_2, L^*_3,... \label{eqn:likelihood_sequence}
+\end{align}
 Defining $X(\ell)$ as the amount of prior mass with likelihood greater than
@@
 Consider $1-t_i$, the fraction of the remaining prior mass that is {\em removed}
-at each iteration.
+in each NS iteration.
 The distribution for the compression factors $t_i \sim $Beta$(N,1)$
@@
 \begin{align}
-P(d(\xref, \x) < \epsilon | \x)
+P(d(\x; \xref) < L \,|\, \xref)
     &\approx
-    p(\x) \int_{d(\xref, \x) < L} \, d\xref\\
-    &= f(\x) \, V
-\end{align}
-where $V$ is the volume of the region where $d(\xref, \x) < L$.
+    p(\xref) \int_{d(\x; \xref) < L} \, d\x\\
+    &= p(\xref) \, V
+\end{align}
+where $V$ is the volume of the region where $d(\x; \xref) < L$.
 Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(L)$.
@@
         \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
-               $p\left(\x | (d(\x) < d_{\rm max})\right)$
+               $p\left(\x \,|\, d(\x) < d_{\rm max}\right)$
             \Comment{Replace worst particle}
@@
 \color{black}
-    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} + [k/N]$
+    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} ++ [k/N]$
             \Comment{Append latest estimate to results}
@@
            \quad\quad$-\int p(\xref) \int p(\x)
-            \log \left[ P(d(\x; \xref) < L | \xref) \right]
+            \log \left[ P(d(\x; \xref) < L \,|\, \xref) \right]
                         \, d\x \, d\xref$,\\ that is, minus the expected value of the
@@
 
-If the threshold $L$ is not small, we might think we are calculating the entropy associated
-with the question `what is $x$, to within a tolerance of $\pm L$?'.
-This is not quite correct. See appendix~\ref{sec:precisional}
+In one dimension,
+if the threshold $L$ is not small, we might think we are calculating the entropy associated
+with the question `what is $\theta$, to within a tolerance of $\pm L$?'.
+This is not quite correct. See Appendix~\ref{sec:precisional}
 for a discussion of this
@@
 In the following sections I demonstrate the algorithm on three
-example problems. For each of the examples, I did 1000 repetitions of NS
-(i.e., 1000 reference particles), and used 10 NS particles and 10000 MCMC
+example problems. For each of the examples, I used 1000 repetitions of NS
+(i.e., 1000 reference particles), and used 10 NS particles and 10,000 MCMC
 steps per NS iteration.
@@
 are specified in {\tt main.cpp}. The default problem is the first example
-from the paper.
+from the paper. For Example 2, since it is a conditional entropy which is
+being estimated (which requires a slight modification to the algorithm),
+an additional argument {\tt InfoNest::Mode::conditional\_entropy}
+must be passed to the {\tt InfoNest::execute} function.
 

### 2017-07-12 paper/ms.tex
Final edits

@@
 at a set of times $\{t_1, ..., t_n\}$, in order to infer the amplitude,
-period, and phase of the signal.
+period, and phase of the signal \citep{bretthorst}.
 Here, I demonstrate the algorithm on a toy version
@@
 \begin{align}
-H(\x | \d) &= \int p(\d') H(\x | \d=\d') \, d\d'.
+H(\x | \d) &= \int p(\d_{\rm ref}) H(\x | \d=\d_{\rm ref}) \, d\d_{\rm ref}.
 \end{align}
 This is an expected value over datasets, of the entropy of the posterior given
-each dataset. To do the Nested Sampling to estimate $H(\x | \d=\d')$, I generate
-initial particles from the posterior by doing standard MCMC (targeting the
-posterior) using the $\x$-value that produced the simulated dataset $\d'$ as
-the initial condition for the MCMC.
+each dataset. I ran the algorithm by generating reference particles from
+$p(\x, \d)$, i.e., generating parameter values from the prior, along with
+corresponding simulated datasets. Given each simulated dataset, I used NS
+to compute an unbiased estimate of $H(\x | \d=\d_{\rm ref})$, by evolving
+a set of particles (initially from the posterior) towards the reference particle.
+The role of the ``prior'' in the NS is actually taken by the posterior, and the
+``likelihood'' was minus a distance function defined in terms of the
+parameter(s).
+
+To generate the initial
+NS particles from the posterior, I used standard MCMC (targeting the
+posterior), initialised at the $\x$-value that produced the simulated dataset
+$\d_{\rm ref}$, which is a perfect sample from the posterior.
 
@@
 To carry out the sampling for this problem, Metropolis-Hastings moves that
-explore the joint prior distribution $p(A, \tau, \phi, \boldsymbol{Y})$
-had to be
-implemented. As with the first example, there were four basic kinds of
-proposals. Four basic kinds of proposals were used: i) those which change
-one of the parameters
-$A$, $\tau$ or $\phi$ while leaving the data $\boldsymbol{Y}$ fixed
-(these are standard Metropolis moves targeting the posterior distribution);
-ii) those which change the parameters and modify the data accordingly;
-iii) those which resample a subset
-of the data; and iv) those which move a single datum slightly.
+explore the posterior distribution $p(A, \tau, \phi | \boldsymbol{Y})$
+had to be implemented. I used heavy-tailed proposals (as in
+Section 8.3 of \citet{dnest4}) which modified one parameter at a time.
 
@@
 \end{align}
-Since the Uniform(0,1) prior has a differential entropy of zero, the
+Since the Uniform(0,1) prior for $\tau$ has a differential entropy of zero, the
 mutual information is $I(\tau; \boldsymbol{Y}) = H(\tau) - H(\tau | \boldsymbol{Y}) = 5.441 \pm 0.038$
@@
 had allowed for shorter periods to be possible, as irregular observing
-schedules are known to reduce aliasing and ambiguity in the inferred period.
+schedules are known to reduce aliasing and ambiguity in the inferred period
+\citep{bretthorst2001nonuniform}.
 When inferring the period of an oscillating signal, multimodal posterior
@@
 ``what is the value of $T$ to within $\pm$ 10\%'', or something along these
-lines. See Appendix~\ref{sec:precisional} for more information about this
+lines. See Appendix~\ref{sec:precisional} for more about this
 issue.
@@
 \begin{align}
-p(\boldsymbol{x} | \alpha) &= \prod_{i=1}^n \frac{1}{x_i^{\alpha + 1}}
+p(\boldsymbol{x} | \alpha) &= \prod_{i=1}^n \frac{\alpha}{x_i^{\alpha + 1}}
 \end{align}
@@
 calculation is based.
-However, the data clearly provides {\em some} information about the future.
+However, the data clearly provides {\em some} nonzero
+amount of information about the future.
 
@@
 $H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
-be done despite the fact that $\alpha$ is unknown. The joint distribution
-for $\ln \ytot$ and $\ln \ztot$, whose properties I am to obtain,
-is shown in Figure~\ref{fig:pareto}.
+be done despite the fact that $\alpha$ is unknown.
 
@@
 under changes of parameterisation).
+The joint distribution
+for $\ln \ytot$ and $\ln \ztot$, whose properties I aimed to obtain,
+is shown in Figure~\ref{fig:pareto}.
 
@@
 Correcting for the volume factor yields a differential entropy estimate
-of $H(\ytot) = 2.366\pm 0.057$ nats.
-By the symmetry of the problem specification, $H(\ztot) = H(\ytot)$.
+of $H(\ln\ytot) = 2.366\pm 0.057$ nats.
+By the symmetry of the problem specification, $H(\ln\ztot) = H(\ln\ytot)$.
 
@@
 \begin{align}
-I(\ytot; \ztot) &= H(\ytot) + H(\ztot) - H(\ytot, \ztot)\\
+I(\ytot; \ztot) &= H(\ln\ytot) + H(\ln\ztot) - H(\ln\ytot, \ln\ztot)\\
                 &= 1.130 \pm 0.082 \textnormal{ nats}.
@@
 relates to the downset of a given statement --- the downset of a statement
-$S$ being the set of all statements that imply S \citep{knuth_questions}.
+$S$ being the set of all statements that imply $S$ \citep{knuth_questions}.
 
@@
 
-We can write the question of interest, $Q$, as a union of ideal questions:
+We can write the question of interest, $Q$, as a union of ideal questions
+(the ideal questions are downsets of statements: the set of all statements
+that imply $S$ is the downset of $S$ and is denoted $\downarrow S$):
 \begin{align}

### 2017-07-12 paper/ms.tex
Removed leading space
Removed unneeded reference

@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{
-Department of Statistics, The University of Auckland, Private Bag 92019,
+\address{Department of Statistics, The University of Auckland, Private Bag 92019,
 Auckland 1142, New Zealand}
@@
 at a set of times $\{t_1, ..., t_n\}$, in order to infer the amplitude,
-period, and phase of the signal \citep{bretthorst}.
+period, and phase of the signal.
 Here, I demonstrate the algorithm on a toy version

### 2017-07-12 paper/ms.tex
Sign fix
Fixed errors in Poisson process part

@@
 is obtainable by letting
-$t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$.
+$t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$ (where
+$X_1 = t_1$.
 The $t_i$ variables represent the proportion of the remaining prior mass
@@
 
-Consider $1-t_i$, the fraction of the remaining prior mass that is {\em removed}
-in each NS iteration.
-The distribution for the compression factors $t_i \sim $Beta$(N,1)$
-corresponds to an exponential distribution for $\ln(1-t_i)$:
-\begin{align}
-\ln (1 - t_i) &\sim \textnormal{Exponential}(N).
-\end{align}
-
-Therefore,
+The sequence of $X$ values is defined by
+\begin{align}
+X_1 &= t_1 \\
+X_2 &= t_2X_1 \\
+X_3 &= t_3X_2
+\end{align}
+and so on, i.e. $X_i$ can be written as
+\begin{align}
+X_i = \prod_{k=1}^i t_k.
+\end{align}
+Taking the log and multiplying both sides by -1, we get
+\begin{align}
+-\ln X_i = \sum_{k=1}^i \left(-\ln t_k\right).\label{eqn:minuslogX}
+\end{align}
+
+Consider $t_i$, the fraction of the remaining prior mass that is retained
+in each NS iteration. By a change of variables,
+the distribution for these compression factors $t_i \sim $Beta$(N,1)$
+corresponds to an exponential distribution for $-\ln(t_i)$:
+\begin{align}
+-\ln t_i &\sim \textnormal{Exponential}(N).\label{eqn:exp}
+\end{align}
+
+From Equations~\ref{eqn:minuslogX} and~\ref{eqn:exp}, we see that
 the sequence of $-\ln X$ values produced by a Nested Sampling run can be

### 2017-07-12 paper/ms.tex
Bracket

@@
 $t_i \sim \textnormal{Beta}(N,1)$ and setting $X_{i+1} := t_iX_{i-1}$ (where
-$X_1 = t_1$.
+$X_1 = t_1$).
 The $t_i$ variables represent the proportion of the remaining prior mass

### 2017-07-30 paper/ms.tex
Putting q(x) in, as Skilling would like to see

@@
 ``what is the value of $x$, precisely?'' remains unresolved
-\citep{knuth_questions}. For example,
+\citep{knuth_questions}. More generally, the `Kullback-Leibler divergence'
+quantifies the degree to which a probability distribution $p(x)$ departs
+from a `base' or `prior' probability distribution $q(x)$:
+\begin{align}
+D_{\rm KL}\big(p \,||\, q\big) &=
+    \sum_{x} p(x) \log\left[\frac{p(x)}{q(x)}\right]
+\end{align}
+\citep{knuth2012foundations, caticha2006updating}.
+If $q(x)$ is taken to be uniform, this is equivalent (up to an additive
+constant and a change of sign) to the Shannon entropy $H(x)$.
+With the negative sign reinstated, Kullback-Leibler divergence is sometimes
+called relative entropy.
+
+The Shannon entropy can be interpreted straightforwardly as a measure of
+uncertainty. For example,
 if there is only one possible value which has probability 1, $H(x)$ is
@@
 quantifies uncertainty by generalising the log-volume
-of the plausible region, when volume is defined with respect to $dx$.
-
-Entropies, and related quantities, tend to be analytically available only for
+of the plausible region, where volume is defined with respect to $dx$.
+Crucially, in continuous spaces, the differential entropy is not
+invariant under changes of coordinates, so any value given for a
+differential entropy must be understood with respect to the coordinate
+system that was used.
+
+Entropies, relative entropies, and related quantities,
+tend to be analytically available only for
 a few families of probability distributions. On the numerical side, if
@@
 
-This paper introduces a computational approach to evaluating the entropy 
-$H(x)$ of probability distributions that can only be sampled
-using Markov Chain Monte Carlo (MCMC).
+This paper introduces a computational approach to evaluating the Shannon
+entropy $H(x)$ of probability distributions that can only be sampled
+using Markov Chain Monte Carlo (MCMC). With minor modifications,
+Kullback-Leibler divergences may also be obtained.
 

### 2017-07-31 paper/ms.tex
Minor clarification
Some Skilling-related corrections

@@
 With the negative sign reinstated, Kullback-Leibler divergence is sometimes
-called relative entropy.
+called relative entropy, and is more fundamental than the Shannon entropy
+\citep{knuth2012foundations}. However, this paper focuses on the Shannon
+entropy for simplicity.
 
@@
 differential entropy must be understood with respect to the coordinate
-system that was used.
+system that was used. Even in discrete cases, the Shannon entropy assumes
+that each possibility counts equally.
 
@@
 for approximating $H(x)$. On the other hand, if $p(x)$ can only be
-{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal
-distribution), then this will not work. Kernel density estimation or the like
+{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal or
+conditional distribution, or the distribution of a quantity derived from $x$),
+then this will not work. Kernel density estimation or the like
 \citep[e.g.][]{JMLR:v15:szabo14a}
@@
 entropy $H(x)$ of probability distributions that can only be sampled
-using Markov Chain Monte Carlo (MCMC). With minor modifications,
-Kullback-Leibler divergences may also be obtained.
+using Markov Chain Monte Carlo (MCMC).
 
@@
 one it is should be clear from the context of the problem at hand.
-I sometimes write general formulae are written in terms of sums (i.e., as they would
+I sometimes write general formulae in terms of sums (i.e., as they would
 appear in a discrete problem), and sometimes as integrals, as they would
@@
 of how the data is related to the parameters
-(i.e., if the parameters where known, what data would likely be observed?).
+(i.e., if the parameters were known, what data would likely be observed?).
 By the product rule, this yields a {\em joint prior}
@@
 \hrulefill
-\State $\widehat{\mathbf{h}} \leftarrow [\,]$
+\State $\widehat{\mathbf{h}} \leftarrow \{\}$
             \Comment{Initialise an empty list of results}
@@
 \color{black}
-    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} ++ [k/N]$
+    \State $\widehat{\mathbf{h}} \leftarrow \textnormal{concatenate}\big(\widehat{\mathbf{h}}, k/N\big)$
             \Comment{Append latest estimate to results}
@@
 \begin{align}
-\log V(\epsilon; n) &= \frac{n}{2}\log \pi
+\log V(L; n) &= \frac{n}{2}\log \pi
                         - \log \Gamma\left(\frac{n}{2} + 1\right)

### 2017-07-31 paper/ms.tex
More Skilling changes

@@
 differential entropy must be understood with respect to the coordinate
-system that was used. Even in discrete cases, the Shannon entropy assumes
-that each possibility counts equally.
+system that was used. Even in discrete cases, using the Shannon entropy assumes
+that each possibility contributes equally when counting possibilities ---
+if this is inappropriate, the relative entropy should be used.
 
@@
 entropy $H(x)$ of probability distributions that can only be sampled
-using Markov Chain Monte Carlo (MCMC).
+using Markov Chain Monte Carlo (MCMC). The key idea is that Nested Sampling
+can create an unbiased estimate of a {\em log}-probability, and entropies
+are averages of log-probabilities.
 
@@
 $d(\x; \xref)$, and the number of NS iterations taken to reduce
-the distance to below some threshold $L$ provides an unbiased estimate of
-\begin{align}
-   \depth = -\log \left[ P(d(\x; \xref) < L) \right]
+the distance to below some threshold $r$ (for `radius')
+provides an unbiased estimate of
+\begin{align}
+   \depth = -\log \left[ P(d(\x; \xref) < r) \right]
 \end{align}
 which I call the `depth'. E.g., if $N=10$ Nested Sampling particles are used,
-and it takes 100 iterations to reduce the distance to below $L$, then the
+and it takes 100 iterations to reduce the distance to below $r$, then the
 depth is 10 nats.
@@
 we can use the fact that density equals mass divided
-by volume. Assume $L$ is small, so that
-\begin{align}
-P(d(\x; \xref) < L \,|\, \xref)
+by volume. Assume $r$ is small, so that
+\begin{align}
+P(d(\x; \xref) < r \,|\, \xref)
     &\approx
-    p(\xref) \int_{d(\x; \xref) < L} \, d\x\\
+    p(\xref) \int_{d(\x; \xref) < r} \, d\x\\
     &= p(\xref) \, V
 \end{align}
-where $V$ is the volume of the region where $d(\x; \xref) < L$.
-Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(L)$.
+where $V$ is the volume of the region where $d(\x; \xref) < r$.
+Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(r)$.
 
@@
             \Comment{the number of Nested Sampling particles to use}
-\State $L \geq 0$
+\State $r \geq 0$
             \Comment{the tolerance} \\
@@
             \Comment{Find the greatest distance}
-    \While{$d_{\rm max} > L$}
+    \While{$d_{\rm max} > r$}
         \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
@@
            \quad\quad$-\int p(\xref) \int p(\x)
-            \log \left[ P(d(\x; \xref) < L \,|\, \xref) \right]
+            \log \left[ P(d(\x; \xref) < r \,|\, \xref) \right]
                         \, d\x \, d\xref$,\\ that is, minus the expected value of the
@@
 \caption{To evaluate the log-probability (or density) of the blue
-probability distribution at the red point, Nested Sampling can be used,
+probability distribution near the red point, Nested Sampling can be used,
 with the blue distribution playing the role of the ``prior'' in NS, and the
@@
 If the distance function $d(\xref, \x)$ is chosen to be Euclidean,
-the constraint $d(\xref, \x) < L$ corresponds to a ball in the space
+the constraint $d(\xref, \x) < r$ corresponds to a ball in the space
 of possible $\xref$ values.
-The log-volume of a ball of radius $L$ in $n$ dimensions is
-\begin{align}
-\log V(L; n) &= \frac{n}{2}\log \pi
-                        - \log \Gamma\left(\frac{n}{2} + 1\right)
-                        + n \log L,
-\end{align}
-
+The volume of such a ball, of radius $r$ in $n$ dimensions, is
+\begin{align}
+V &= \frac{(\pi r^2)^{n/2}}{(\frac{n}{2})!}.
+\end{align}
 In one dimension,
-if the threshold $L$ is not small, we might think we are calculating the entropy associated
-with the question `what is $\theta$, to within a tolerance of $\pm L$?'.
+if the threshold $r$ is not small, we might think we are calculating the entropy associated
+with the question `what is $\theta$, to within a tolerance of $\pm r$?'.
 This is not quite correct. See Appendix~\ref{sec:precisional}
@@
 I ran the algorithm and computed the average depth using a tolerance of
-$L=10^{-3}\sqrt{100}$, so that the RMS difference between
+$r=10^{-3}\sqrt{100}$, so that the RMS difference between
 points in the two datasets was about $10^{-3}$.
@@
 The log-volume of a 100-dimensional ball of radius
-$L=10^{-3}\sqrt{100}$ is
+$r=10^{-3}\sqrt{100}$ is
 $-551.76$. Therefore, the differential entropy is estimated to be
@@
 $A$ and $\phi$ are nuisance parameters, I calculated the conditional entropy
-$H(\tau | \boldsymbol{Y})$ using the algorithm with a tolerance of $L = 10^{-5}$.
+$H(\tau | \boldsymbol{Y})$ using the algorithm with a tolerance of $r = 10^{-5}$.
 The result was
@@
 reference particle in terms of $\ln \ytot$. The result, using a tolerance
-of $L = 10^{-3}$, was:
+of $r = 10^{-3}$, was:
 \begin{align}
@@
 line. The probability contained in an interval
-$[x_0, x_0 + L]$, which has length $L$, is
-\begin{align}
-\int_{x_0}^{x_0 + L} f(x) \, dx &= F(x_0 + L) - F(x_0)
+$[x_0, x_0 + r]$, which has length $r$, is
+\begin{align}
+\int_{x_0}^{x_0 + r} f(x) \, dx &= F(x_0 + r) - F(x_0)
 \end{align}
@@
 This probability can be considered as a function of two variables,
-$x_0$ and $L$, which I will denote by $P(,)$:
-\begin{align}
-P(x_0, L) &= F(x_0 + L) - F(x_0).
+$x_0$ and $r$, which I will denote by $P(,)$:
+\begin{align}
+P(x_0, r) &= F(x_0 + r) - F(x_0).
 \end{align}
@@
 \begin{align}
-Q(x_0, L) &= -P(x_0, L) \log P(x_0, L).
+Q(x_0, r) &= -P(x_0, r) \log P(x_0, r).
 \end{align}
 Consider the rate of change of $Q$ as the interval is shifted to
-the right, but with its width $L$ held constant:
+the right, but with its width $r$ held constant:
 \begin{align}
@@
     \left[-P\log P\right] \\
-    &= -\left[1 + \log P(x_0, L)\right]\frac{\partial P}{\partial x_0} \\
-    &= -\left[1 + \log P(x_0, L)\right]\left[f(x_0 + L) - f(x_0)\right].
+    &= -\left[1 + \log P(x_0, r)\right]\frac{\partial P}{\partial x_0} \\
+    &= -\left[1 + \log P(x_0, r)\right]\left[f(x_0 + r) - f(x_0)\right].
 \end{align}
@@
 \begin{align}
-\frac{\partial Q}{\partial L} &= \frac{\partial}{\partial L}
-    \left[-P(x_0, L) \log P(x_0, L)\right] \\
-    &= -\left[1 + \log P(x_0, L)\right]\frac{\partial P}{\partial L} \\
-    &= -\left[1 + \log P(x_0, L)\right]f(x_0 + L).
+\frac{\partial Q}{\partial r} &= \frac{\partial}{\partial r}
+    \left[-P(x_0, r) \log P(x_0, r)\right] \\
+    &= -\left[1 + \log P(x_0, r)\right]\frac{\partial P}{\partial r} \\
+    &= -\left[1 + \log P(x_0, r)\right]f(x_0 + r).
 \end{align}
@@
 $Q$ terms. The extra entropy from adding the interval
-$[x_0, x_0 + L]$ and removing the overlap $[x_0, x_0 + L - h]$, for small $h$, is
-\begin{align}
-\delta H &= Q(x_0, L) - Q(x_0, L - h)\\
-         &= h\frac{\partial Q}{\partial L} \\
-         &= -h\left[1 + \log P(x_0, L)\right]f(x_0 + L).
+$[x_0, x_0 + r]$ and removing the overlap $[x_0, x_0 + r - h]$, for small $h$, is
+\begin{align}
+\delta H &= Q(x_0, r) - Q(x_0, r - h)\\
+         &= h\frac{\partial Q}{\partial r} \\
+         &= -h\left[1 + \log P(x_0, r)\right]f(x_0 + r).
 \end{align}
@@
 H &= -\int_{-\infty}^\infty
-        \left[1 + \log P(x-L, L)\right]f(x)
+        \left[1 + \log P(x-r, r)\right]f(x)
       \, dx \\
   &= -\int_{-\infty}^\infty
-        \left(1 + \log\left[F(x) - F(x-L)\right]\right)f(x)
+        \left(1 + \log\left[F(x) - F(x-r)\right]\right)f(x)
       \, dx.
@@
 H' &=  -\int_{-\infty}^\infty
-        \left(1 + \log\left[F(x+L) - F(x)\right]\right)f(x)
+        \left(1 + \log\left[F(x+r) - F(x)\right]\right)f(x)
       \, dx \\
@@
 small). However, the algorithm can be made to estimate the entropy of the
-precisional question by redefining the distance function such that
-$d(\x; \xref)$ can only ever be below $L$ if $\x < \xref$ (or alternatively,
-if $\x > \xref$).
+precisional question, by redefining the distance function. For a one-dimensional
+quantity of interest, the distance function can be defined such that
+$d(x; x_{\rm ref})$ can only ever be below $r$ if $x < x_{\rm ref}$ (or alternatively,
+if $x > x_{\rm ref}$).
 
@@
 By default, {\tt postprocess.py} estimates the depth with a tolerance of
-$L=10^{-3}$. This value can be changed by calling the {\tt postprocess}
+$r=10^{-3}$. This value can be changed by calling the {\tt postprocess}
 function with a different value of its argument {\tt tol}. E.g.,

### 2017-07-31 paper/ms.tex
More Skilling comments addressed

@@
 $X_1 = t_1$).
-The $t_i$ variables represent the proportion of the remaining prior mass
+The $t_i$ quantities are the proportion of the remaining prior mass
 that is {\em retained} at each iteration, after imposing the constraint of the
@@
 \mu | \boldsymbol{x} &\sim \textnormal{Normal}\left(
-                                       \frac{1}{100}\sum_{i=1}^{100} x_i,
+                                       \frac{1}{100.01}\sum_{i=1}^{100} x_i,
                                        \left[\frac{10}{\sqrt{10001}}\right]^2
@@
 \end{align}
-Converting to a differential entropy by adding $\ln (2 \times 10^{-5})$, the conditional
+Converting to a differential entropy by adding $\ln (2r) = -10.820$, the conditional
 entropy is
@@
 changes in the number of MCMC steps used is a simple way to increase one's
-confidence in the results, but is not a guarantee.
+confidence in the results, but is not a guarantee (no general guarantee can
+exist, because any algorithm based on sampling a subset of a space might miss
+a crucial by hard-to-locate dominant spike).
 Despite this weakness, I hope the algorithm will be

### 2017-07-31 paper/ms.tex
Acknowledgement

@@
 \acknowledgments{Acknowledgements}
+I would like to thank John Skilling (MaxEnt Data Consultants) and the
+anonymous referee for their critiques.
 It is a pleasure to thank the following people for interesting and helpful

### 2017-07-31 paper/ms.tex
Mention the tradeoff in choosing the distance threshold

@@
 Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(r)$.
+Reducing $r$ towards zero ought to lead to
+a more accurate result because the probability density at a point will
+be more accurately approximated by the probability of a small region containing
+the point, divided by the volume of that region.
+On the other hand, smaller values of $r$ will lead to more Nested Sampling
+iterations required, and hence less accuracy in the depth estimates.
+Thus, there is a an awkward trade-off involved in choosing the value of $r$,
+which will need to be investigated more thoroughly in the future.
 

### 2017-07-31 paper/ms.tex
Fixed Walter reference

@@
 \citep{skilling2006nested}.
-\citet{Walter2015} showed how this view of the NS sequence of points
+\citet{Walter2017point} showed how this view of the NS sequence of points
 can be used to construct a

### 2017-08-11 paper/ms.tex
Minor clarification
revisions in blue

@@
 \definecolor{green}{rgb}{0, 0.5, 0}
-
+\definecolor{darkblue}{rgb}{0, 0, 0.7}
+
+\newcommand{\revision}{\color{darkblue}}
 \renewcommand{\d}{\boldsymbol{d}}
@@
 ``what is the value of $x$, precisely?'' remains unresolved
-\citep{knuth_questions}. More generally, the `Kullback-Leibler divergence'
+\citep{knuth_questions}. {\revision More generally, the `Kullback-Leibler divergence'
 quantifies the degree to which a probability distribution $p(x)$ departs
@@
 \citep{knuth2012foundations}. However, this paper focuses on the Shannon
-entropy for simplicity.
+entropy for simplicity, and because the resulting algorithm calculates
+the expected value of a log-probability.
 
 The Shannon entropy can be interpreted straightforwardly as a measure of
-uncertainty. For example,
+uncertainty.} For example,
 if there is only one possible value which has probability 1, $H(x)$ is
@@
 of the plausible region, where volume is defined with respect to $dx$.
-Crucially, in continuous spaces, the differential entropy is not
+{\revision Crucially, in continuous spaces, the differential entropy is not
 invariant under changes of coordinates, so any value given for a
@@
 that each possibility contributes equally when counting possibilities ---
-if this is inappropriate, the relative entropy should be used.
-
-Entropies, relative entropies, and related quantities,
+if this is inappropriate, the relative entropy should be used.}
+
+Entropies, {\revision relative entropies}, and related quantities,
 tend to be analytically available only for
@@
 for approximating $H(x)$. On the other hand, if $p(x)$ can only be
-{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal or
-conditional distribution, or the distribution of a quantity derived from $x$),
+{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal {\revision or
+conditional distribution, or the distribution of a quantity derived from $x$)},
 then this will not work. Kernel density estimation or the like
@@
 entropy $H(x)$ of probability distributions that can only be sampled
-using Markov Chain Monte Carlo (MCMC). The key idea is that Nested Sampling
+using Markov Chain Monte Carlo (MCMC). {\revision The key idea is that Nested Sampling
 can create an unbiased estimate of a {\em log}-probability, and entropies
-are averages of log-probabilities.
+are averages of log-probabilities}.
 
@@
 
+{\revision
 The sequence of $X$ values is defined by
@@
 
-From Equations~\ref{eqn:minuslogX} and~\ref{eqn:exp}, we see that
+From Equations~\ref{eqn:minuslogX} and~\ref{eqn:exp}, we see that}
 the sequence of $-\ln X$ values produced by a Nested Sampling run can be
@@
 $d(\x; \xref)$, and the number of NS iterations taken to reduce
-the distance to below some threshold $r$ (for `radius')
+the distance to below some threshold {\revision $r$ (for `radius')}
 provides an unbiased estimate of
@@
 Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(r)$.
-Reducing $r$ towards zero ought to lead to
+{\revision Reducing $r$ towards zero ought to lead to
 a more accurate result because the probability density at a point will
@@
 Thus, there is a an awkward trade-off involved in choosing the value of $r$,
-which will need to be investigated more thoroughly in the future.
+which will need to be investigated more thoroughly in the future.}
 
@@
 \color{black}
-    \State $\widehat{\mathbf{h}} \leftarrow \textnormal{concatenate}\big(\widehat{\mathbf{h}}, k/N\big)$
+    \State {\revision  $\widehat{\mathbf{h}} \leftarrow \textnormal{concatenate}\big(\widehat{\mathbf{h}}, k/N\big)$ }
             \Comment{Append latest estimate to results}
@@
 \caption{To evaluate the log-probability (or density) of the blue
-probability distribution near the red point, Nested Sampling can be used,
+probability distribution {\revision near} the red point, Nested Sampling can be used,
 with the blue distribution playing the role of the ``prior'' in NS, and the
@@
 The volume of such a ball, of radius $r$ in $n$ dimensions, is
+{\revision
 \begin{align}
 V &= \frac{(\pi r^2)^{n/2}}{(\frac{n}{2})!}.
-\end{align}
+\end{align}}
 In one dimension,
@@
 \end{align}
-Converting to a differential entropy by adding $\ln (2r) = -10.820$, the conditional
+Converting to a differential entropy by adding {\revision $\ln (2r) = -10.820$}, the conditional
 entropy is
@@
 changes in the number of MCMC steps used is a simple way to increase one's
-confidence in the results, but is not a guarantee (no general guarantee can
+confidence in the results, but is not a guarantee {\revision (no general guarantee can
 exist, because any algorithm based on sampling a subset of a space might miss
-a crucial by hard-to-locate dominant spike).
+a crucial by hard-to-locate dominant spike)}.
 Despite this weakness, I hope the algorithm will be
@@
 \acknowledgments{Acknowledgements}
-I would like to thank John Skilling (MaxEnt Data Consultants) and the
-anonymous referee for their critiques.
+{\revision I would like to thank John Skilling
+(Maximum Entropy Data Consultants) and the
+anonymous referee for their comments.}
 It is a pleasure to thank the following people for interesting and helpful
@@
 small). However, the algorithm can be made to estimate the entropy of the
-precisional question, by redefining the distance function. For a one-dimensional
+precisional question, by redefining the distance function. {\revision For a one-dimensional
 quantity of interest, the distance function can be defined such that
 $d(x; x_{\rm ref})$ can only ever be below $r$ if $x < x_{\rm ref}$ (or alternatively,
-if $x > x_{\rm ref}$).
+if $x > x_{\rm ref}$).}
 

### 2017-08-11 paper/response.tex
Working on response to referees

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{natbib}
\usepackage{microtype}

\renewcommand{\quote}{\em}

\title{Response to referees}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{\noindent Abstract}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

\section*{Response to John Skilling's comments}

\section*{Response to the other reviewer's comments}

{\quote
This is a very interesting paper that pushes state-of-the-art in both theory and practice.  As a result, there is much to learn here and it will take further effort for me to fully absorb the ideas presented in this paper.  As a result, this review is being written based on what I understand of this paper to date.  Perhaps there are places in which the author can clarify the presentation.}

Thanks for the positive comments. I have tried to clarify the presentation.

{\quote
Estimation of entropy (and other information-theoretic quantities) and estimation of the uncertainties of the estimates is a notoriously difficult problem.  I personally, would have benefited greatly from having access, or knowledge of, such a methodology years ago.  The paper should definitely be published after some minor edits.}

That's why I started thinking about this problem...

{\quote
1. The author presents a methodology for computing entropies, and other information-theoretic quantities, using Skilling's nested sampling algorithm.  It should be noted that Skilling himself advocates the estimation of the "entropy", I believe that it is the KL-divergence, as the estimate of the uncertainty in the estimated evidence.  This should be mentioned, and possibly discussed within the context of this present paper.}

{\quote
2. With regard to the algorithm to compute the depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.  Also at that step, do the square brackets refer to an array (of one element)?  Or do they refer to some kind of rounding, which I doubt.}

Sorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
$\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)$
\end{align}
which is hopefully more understandable.

{\quote
3. In Section 2, the author writes "1. Points can be generated from the prior..."  Please clarify what is meant by "points".  Does the author mean samples?}

Yes. I have changed points to samples.

The author writes:
"The former measures the degree to which one question answers another ex ante, and is a function of two questions. The latter measures the degree to which a statement answered a question ex post, and is a function of a question and a statement."
which reveals that the author has really have been thinking about questions!

The author also writes:
"The mutual information is another way of describing the relevance of the data to the parameters. its definition, and relation to other quantities"
As far as I am aware, this is a novel way of thinking about mutual information!  Very nice!
In addition, I did not know that it could be written as the expected value of KL over datasets, which is fascinating.

4. It would be useful to follow (13) with the calculation written in your notation rather than Skilling's.

5. After (13): Why does the implied distribution of L-values tend to be heavy-tailed?

6. Just before (14) and in (15): I do not understand the notation regarding the implicit likelihood boundary involving \\mathbb\{1\}

7. After (26) what is meant by "NS repetitions"?

8. After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"

9. It wasn't clear to me how the distance function (25) was chosen and why.

10. I had some questions about how the algorithm was conceived.  I had a difficult time envisioning how the blue distribution played the role of the prior, and why.  Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.

11.  It was not entirely clear to me how the algorithm was used to estimate the entropy of a Gaussian.  How was the posterior (24) arrived at.  The prior is normal.  But what about the likelihood?  And then how was the distance measure (25) computed?

There are a number of typos:

1. End of Section 1: "I sometimes write general formulae are written ..."

2. Missing citation in first sentence of Section 6

3. After (47): question mark missing in case ii)

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}

### 2017-08-11 paper/ms.tex
Working on response to referees

@@
 uncertainty.} For example,
-if there is only one possible value which has probability 1, $H(x)$ is
+if there is only one possible value which has probability one, $H(x)$ is
 zero. Conventionally, $0 \times \log 0$ is defined to be equal
@@
 \begin{enumerate}
-  \item Points can be generated from the prior $p(\x)$;
+  \item {\revision Samples} can be generated from the prior $p(\x)$;
   \item Simulated datasets can be generated from $p(\d | \x)$ for any
@@
 \color{black}
-    \State {\revision  $\widehat{\mathbf{h}} \leftarrow \textnormal{concatenate}\big(\widehat{\mathbf{h}}, k/N\big)$ }
+    \State {\revision  $\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)$ }
             \Comment{Append latest estimate to results}

### 2017-08-11 paper/response.tex
More work on the revision

@@
 
+{\quote
 The author writes:
@@
 The latter measures the degree to which a statement answered a question ex post, and is a function of a question and a statement."
-which reveals that the author has really have been thinking about questions!
+which reveals that the author has really have been thinking about questions!}
 
+I'm glad you liked it! If you found this interesting, check out the foundations
+work of Kevin Knuth, which I find fascinating.
+
+{\quote
 The author also writes:
@@
 As far as I am aware, this is a novel way of thinking about mutual information!  Very nice!
-In addition, I did not know that it could be written as the expected value of KL over datasets, which is fascinating.
+In addition, I did not know that it could be written as the expected value of KL over datasets, which is fascinating.}
+
+Thanks! In most information theory textbooks, these things are written in
+terms of abstract `random variables' $X$ and $Y$. One of my subsidiary
+goals in this paper was to discuss this stuff in terms and notation more
+familiar to Bayesians.
 

### 2017-08-11 paper/ms.tex
More work on the revision

@@
 entropy of $\x$ given a particular value of $\d$, which might be written
-$H(\x | \d=\d_0)$.
+$H(\x | \d=\d_{\rm observed})$.
 The former measures the degree to which one question answers
 another {\em ex ante}, and is a function of two questions.
-The latter measures the degree to which a {\em statement} answered a question
-{\em ex post}, and is a function of a question and a statement.
+The latter measures the degree to which {\revision a question remains unresolved
+after conditioning on a specific statement ({\em ex post})},
+and is a function of a question and a statement.
 
@@
 its aims.
+{\revision
+NS also estimates the Kullback-Leibler divergence from the prior
+$\pi(\x)$ to the posterior $\pi(x)L(x)/Z$, which Skilling calls the
+information. This is a measure of how much was learned about $\x$
+from the specific dataset, and is also useful for defining a termination
+criterion for NS.}
 

### 2017-08-11 paper/response.tex
More editing

@@
 
-4. It would be useful to follow (13) with the calculation written in your notation rather than Skilling's.
+{\quote
+4. It would be useful to follow (13) with the calculation written in your notation rather than Skilling's.}
 
-5. After (13): Why does the implied distribution of L-values tend to be heavy-tailed?
+I decided to do this the other way around, first writing it in my notation and
+then switching to Skilling's, which is more convenient when the computational
+algorithm is being discussed.
 
-6. Just before (14) and in (15): I do not understand the notation regarding the implicit likelihood boundary involving \\mathbb\{1\}
+{\quote
+5. After (13): Why does the implied distribution of $L$-values tend to be heavy-tailed?}
+
+Imagine doing simple Monte Carlo to get $Z$, by drawing from the prior and
+collecting likelihood values, which you'll want to average. A histogram of
+those likelihood values will look ridiculous, and you might have to wait
+a million years until you get enough points such that your average becomes
+accurate, because it will be dominated by huge likelihoods which are seen
+rarely. Most NS people talk about this in terms of the parameter space, but
+it also has an interpretation in terms of a `probability distribution for
+$L$' which is implied by $\pi$.
+
+{\quote
+6. Just before (14) and in (15): I do not understand the notation regarding the implicit likelihood boundary involving \\mathbb\{1\}}
+
+I've added a bit of explanation of the indicator function notation.
 

### 2017-08-11 paper/ms.tex
More editing

@@
 \citet{skilling2006nested}, is an algorithm whose aim is
-to calculate the evidence
-\begin{align}
-Z &= \int \pi(\x) L(\x) \, d\x
-\end{align}
-where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
+to calculate the evidence {\revision
+\begin{align}
+p(\x) &= \int p(\x) p(\d | \x) \, d\x,
+\end{align}
+or, in the simplified notation common when discussing computational matters,
+\begin{align}
+Z &= \int \pi(\x) L(\x) \, d\x,
+\end{align}}
+\noindent where $\x$ is the unknown parameter(s), $\pi$ is the prior distribution,
 and $L$ is the likelihood function. The original NS algorithm, and variations
@@
 respect to $\pi$, the
-implied distribution of $L$-values tends to be very heavy-tailed, which is
-why simple Monte Carlo does not work. Equivalently, the integral is dominated
-by very small regions where $L(\x)$ is high.
+{\revision distribution of $L$-values implied by $\pi$}
+tends to be very heavy-tailed, which is
+why simple Monte Carlo does not work. Equivalently, the integral 
+{\revision over $\x$} is dominated by very small regions where $L(\x)$ is high.
 
@@
 drawn from the prior $\pi$ but subject to the constraint that its likelihood
-must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
-and evolving it with MCMC to explore
-$\pi(\x)\mathds{1}\left(L(\x) > L^*_1\right)$).
+must be above $L^*_1$. {\revision This is usually achieved by cloning a surviving particle
+and evolving it with MCMC to explore the prior distribution but with
+a constraint on likelihood value, $L(\x) > L^*_1$. This `constrained prior'
+distribution has probability density proportional to
+\begin{align}
+\pi(\x)\mathds{1}\left(L(\x) > L^*_1\right)
+\end{align}
+where $\mathds{1}()$ is an indicator function which returns one if the argument
+is true and zero if it is false.}
+
 This process is repeated, resulting in an

### 2017-08-11 paper/ms.tex
typo fix

@@
 exist, because any algorithm based on sampling a subset of a space might miss
-a crucial by hard-to-locate dominant spike)}.
+a crucial but hard-to-locate dominant spike)}.
 Despite this weakness, I hope the algorithm will be

### 2017-08-11 paper/response.tex
More work on Kevin's comments

@@
 
-I'm glad you liked it! If you found this interesting, check out the foundations
-work of Kevin Knuth, which I find fascinating.
+I'm glad you liked it!
 
@@
 
-Thanks! In most information theory textbooks, these things are written in
+Thanks. In most information theory textbooks, these things are written in
 terms of abstract `random variables' $X$ and $Y$. One of my subsidiary
@@
 
-7. After (26) what is meant by "NS repetitions"?
+{\quote
+7. After (26) what is meant by "NS repetitions"?}
 
+It's the number of reference particles
+whose depth was estimated using an NS run. I've added a sentence to this effect.
+
+{\quote
 8.
-After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"
+After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"}
 
@@
 
+{\quote
 There are a number of typos:
 
-1. End of Section 1: "I sometimes write general formulae are written ..."
+1. End of Section 1: "I sometimes write general formulae are written ..."}
+Fixed.
 
-2. Missing citation in first sentence of Section 6
+{\quote
+2. Missing citation in first sentence of Section 6}
+I removed that citation.
 
-3. After (47): question mark missing in case ii)
+{\quote
+3. After (47): question mark missing in case ii)}
+Fixed.
 

### 2017-08-11 paper/ms.tex
More work on Kevin's comments

@@
 standard deviation of the depth estimates divided by square root of
-the number of NS repetitions.
+the number of NS repetitions {\revision (the number of reference particles
+whose depth was evaluated)}.
 
@@
 In this section I use the algorithm to answer the questions:
-i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
+i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$?; and
 iii) How informative is $\ytot$ about $\ztot$? These are quantified by

### 2017-08-13 paper/ms.tex
One more formula for mutual information
Working on more of Kevin's comments

@@
 
-To use the algorithm, a distance function $d(,)$ must be defined, to quantify
+To use the algorithm, a distance function $d(;)$ must be defined, to quantify
 how far the NS particles are from the reference particle. To calculate
-the entropy of the data, I used the following distance function:
-\begin{align}
-d\left(\x; \xref\right) &= \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2
-\end{align}
+the entropy of the data, I used the following distance function,
+{\revision based on
+a simple Euclidean metric in the space of datasets:
+\begin{align}
+d\left(\x; \xref\right) &= \sqrt{
+        \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2}
+\end{align}
+}
 
@@
 I(\x; \d) &= H(\x) - H(\x|\d) \\
-I(\x; \d) &= \int p(\d) \, D_{\rm KL}\left(p(\x|\d) \,||\, p(\x)\right) \, d\d.
+I(\x; \d) &= \int p(\d) \, D_{\rm KL}\left(p(\x|\d) \,||\, p(\x)\right) \, d\d\\
+I(\x; \d) &= D_{\rm KL}\left(p(\x, \d) \,||\, p(\x)p(\d)\right).
 \end{align}

### 2017-08-14 paper/response.tex
One more formula for mutual information
Working on more of Kevin's comments
More work on Kevin's comments

@@
 \begin{align}
-$\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)$
+\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
 \end{align}
@@
 
-9. It wasn't clear to me how the distance function (25) was chosen and why.
+{\quote
+9. It wasn't clear to me how the distance function (25) was chosen and why.}
 
+It is just the Euclidean distance between two data vectors, each of which
+contains 100 co-ordinates. I added a sentence to the paper to clarify this.
+
+{\quote
 10.
@@
 I had a difficult time envisioning how the blue distribution played the role of the prior, and why.
-Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.
+Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.}
 
+I know a few groups use that convention (transform the parameter space so
+the prior becomes uniform), but I never have, and there's no reason to do
+that in general, as long as it's possible to explore $\pi$ then everything
+should be fine.
+
+{\quote
 11.
@@
 But what about the likelihood?
-And then how was the distance measure (25) computed?
+And then how was the distance measure (25) computed?}
+
+I have added an extra paragraph with the intent of explaining this in more
+detail.
 

### 2017-08-14 paper/ms.tex
More work on Kevin's comments

@@
 \end{align}
-If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
+{\revision
+I assumed a Normal$(0, 10^2)$ prior for $\mu$.
+Before getting the data, the prior information is described by the joint
+distribution
+\begin{align}
+p(\mu, \boldsymbol{x}) &= p(\mu)p(\boldsymbol{x} | \mu).\label{eqn:joint_prior}
+\end{align}
+The prior over datasets is given by the marginal distribution
+\begin{align}
+p(\boldsymbol{x}) &= \int p(\mu, \boldsymbol{x}) \, d\mu.
+\end{align}
+This is the distribution whose differential entropy I seek in this section.
+Its true value is
+available analytically
+in this case, by noting that the posterior is available in closed form
+for any dataset:
 \begin{align}
@@
 \end{align}
-The true value of the differential entropy $H(\boldsymbol{x})$ is
-available analytically
-in this case, and its value is $146.499$ nats. However, in most situations, the
-entropy of a marginal distribution is not available in closed form.
+which enables the calculation of the mutual information and hence
+$H(\boldsymbol{x})$.
+The true value of $H(\boldsymbol{x})$ is $146.499$ nats.
+However, in most situations, the
+entropy of a marginal distribution such as this is not available in closed form.
 
@@
 how far the NS particles are from the reference particle. To calculate
-the entropy of the data, I used the following distance function,
-{\revision based on
-a simple Euclidean metric in the space of datasets:
+the entropy of the data, {\revision I applied the algorithm in the joint
+space of possible parameters and datasets}, i.e., distribution of reference
+particles followed Equation~\ref{eqn:joint_prior}.
+For the distance function, which helps the NS particles approach the
+reference particle, I used a
+simple Euclidean metric in the space of datasets:
 \begin{align}
 d\left(\x; \xref\right) &= \sqrt{
-        \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2}
-\end{align}
+        \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2}.
+\end{align}
+Since the distance function refers only to the data $\boldsymbol{x}$
+and not the parameter $\mu$, the sampling is effectively done in the
+space of possible datasets only --- the parameter $\mu$ can be thought of
+as merely a
+latent variable that allows $p(\boldsymbol{x})$ to be explored.
 }

### 2017-08-14 paper/response.tex
Finished Kevin's comments

@@
 \usepackage{amsmath}
+\usepackage{mathtools}
 \usepackage{dsfont}
@@
 \usepackage{microtype}
+\usepackage{amssymb}
+\usepackage{color}
 
-\renewcommand{\quote}{\em}
+\definecolor{orange}{rgb}{0.7, 0.25, 0}
+\definecolor{green}{rgb}{0, 0.5, 0}
+\definecolor{darkblue}{rgb}{0, 0, 0.7}
+
+\renewcommand{\quote}{\em \color{orange}}
 
@@
 This should be mentioned, and possibly discussed within the context of this present paper.}
+
+I have mentioned that NS computes the prior-to-posterior KL divergence, and
+that this has both an interesting interpretation and is used as a termination
+criterion. This was added to the first paragraph of Section 3.
 
@@
 
+I have tried to elaborate on this at the end of Section 7.
+
 {\quote
@@
 the prior becomes uniform), but I never have, and there's no reason to do
-that in general, as long as it's possible to explore $\pi$ then everything
-should be fine.
+that in general. As long as it's possible to explore $\pi$, then everything
+will be fine.
 
@@
 1. End of Section 1: "I sometimes write general formulae are written ..."}
+
 Fixed.
@@
 2. Missing citation in first sentence of Section 6}
+
 I removed that citation.
@@
 3. After (47): question mark missing in case ii)}
+
 Fixed.

### 2017-08-14 paper/ms.tex
Finished Kevin's comments

@@
 entropy of $\x$ given a particular value of $\d$, which might be written
-$H(\x | \d=\d_{\rm observed})$.
+$H(\x | \d=\d_{\rm obs})$ (`obs' for observed).
 The former measures the degree to which one question answers
@@
 with the blue distribution playing the role of the ``prior'' in NS, and the
-Euclidian distance from the red point (illustrated with red contours)
+Euclidean distance from the red point (illustrated with red contours)
 playing the role of the negative log-likelihood. Averaging over selections
@@
 using a common sequence of reference points for the marginal and joint entropy
-runs, an instance of the `common random numbers'
+runs. This is an instance of the `common random numbers'
 technique\footnote{\tt http://demonstrations.wolfram.com/TheMethodOfCommonRandomNumbersAnExample/}.
+{\revision In short, for those reference particles with a large depth
+in terms of the $\ln \ytot$ distance function, the depth in terms of the
+$(\ln \ytot, \ln \ztot)$ distance function was also large. Since the
+mutual information involves a difference of the averages of these depths,
+it is more accurate to compute the average difference 
+(i.e. compute the two depths using the same reference particle, to compute
+the difference between the two depths) than the difference of the averages.
+}
 

### 2017-08-14 paper/response.tex
Finalish changes

@@
 
+{\quote
+This is an interesting paper, which addresses how to compute entropy values directly, without explicit
+evaluation of the underlying probability distribution. There are a few places where I feel that presentation
+could be improved, which I list below. Nevertheless, I should note two particular points.
+
+1. Definition of entropy}
+
+I agree with Skilling that the fundamental definition of entropy involves
+a base measure $q(x)$, and that relative entropies or KL divergences are the
+appropriate objects to be working with. However, since (as he points out)
+the algorithm more naturally involves the averages of log probabilities,
+I have decided to leave most of the discussion in terms of Shannon entropies.
+However, I have added extra discussion to the introduction, about how using
+the Shannon entropy is equivalent to a uniform $q$, and cautioning that
+this should be kept in mind. I have also added some text after the definition
+of the log-volume of a ball of radius $r$, which clarifies where the implicit
+assumption of a uniform $q(x)$ comes into the procedure.
+
+{\quote
+2. Possible future work
+There is an infelicity at the heart of the author’s approach. The little target balls of radius $L$ need to be
+small in order to avoid undue smudging of the pointwise values being estimated. Yet, as L shrinks, the
+estimates lose accuracy because of the square-root-N variability of nested-sampling volumes. So the user
+will need to compromise, and it’s possible that future work may discover how to ameliorate this.}
+
+This is an excellent point, and something I have worried about. I have added
+some discussion of this in Section 4 (search for the paragraph containing
+`trade-off'). I think the current algorithm will still work well when the
+thing we want the entropy of is low-dimensional, even if it's some function of
+an underlying high-dimensional quantity.
+
+{\quote
+I suggest that applying the author’s method directly to the Kullback-Leibler form (equations 11 and
+12) may offer a better way forward.}
+
+It was not obvious to me how this would work (except perhaps from using a
+distance function defined in the coordinate system in which $q$ is uniform).
+I will think about it, and present any future work in this way if it is
+sufficiently neat.
+
+{\quote 
+Near the end of page 9, the author argues against trying that
+because “the estimate for a given dataset may be biassed”. Possibly so, though I would argue that the
+dominant uncertainty with nested sampling is statistical, with numerical biasses (of sensible quantities)
+being considerably less. Anyway, it’s not entirely clear in these nonlinear problems whether bias matters
+much. I have in mind the Bayesian evidence Z, where an unbiassed estimate would give a seriously
+biassed estimate of the logarithm log Z (which is what matters for Bayes factors), and vice versa.
+The advantage of aiming at KL is that the statistic can be computed “for free” parasitically upon
+a nested sampling run. In fact, I myself use apparent convergence of KL as the most reliable general
+termination criterion I have yet found. (Hint to author: it converges quicker if it’s evaluated as the sum
+$KL =
+X \delta L
+Z
+log
+L
+Z$
+over likelihood L, rather than using the corresponding weighted sum over prior mass shells $\delta X$ that’s
+better for getting Z and the posterior.)}
+
+It is not the bias that I was particularly worried about. I have added a
+comment about the bias potentially being negligible in many applications.
+What I was more concerned about was being able to integrate over nuisance
+parameters as well. I can't see a way to get the prior-to-posterior
+KL divergence {\em of nuisance parameters only} using standard NS, but I can
+get it using my approach.
+
+{\quote
+To the editors: I am not suggesting that there’s anything wrong with the author’s proposals. We
+advance by trying things, and even if something does become superseded — which this may not — it
+1still gives helpful perspective.}
+
+I agree. In fact, this paper has gone through a few distinct stages where I
+started seeing better ways of doing certain things. It's possible that it
+will be superseded quickly, but I do think it's an advance. I'd like to get
+people thinking about Nested Sampling in different scenarios, i.e., where
+pi and L are 'any distribution' and 'any scalar function' rather than necessarily
+being prior and likelihood. We might find applications to things other than
+data analysis and statistical mechanics by doing that.
+
+{\quote
+Minor improvements
+1. Wrong English in last sentence of 1.1 (page 2).}
+
+Fixed.
+
+{\quote
+2. “were” not “where” at beginning of 2 (page 2) — use brain not spellchecker.}
+
+Fixed.
+
+{\quote
+3. I found the notation a bit confusing. $H(\theta | d)$ looks like a function of the posterior $\theta$’s at specific
+data d, whereas it actually means the average hHi for unknown d. I agree that it’s unnecessarily fussy
+to write the formally correct H(P (· | ·)) for H as a functional of the posterior as a function of arbitrary
+arguments which readers are supposed to remember are parameters and data. However, angle brackets,
+$hH(\theta | d)i$ or $H(\theta | d)i d$ would have helped me by giving a useful clue. They might also let the author
+avoid the idiosyncratic $H(\theta | d = d 0 )$ by writing $H(\theta | d 0 )$ without the angle brackets for H at the specific
+d 0 .
+There are various overloaded definitions of H according to type of argument (see equations 8, 9, 10).
+Introducing an explicit common format
+X X
+H(argument) $\equiv$
+p($\theta$, d) log p(argument)
+would bring these together and incidentally help to stress the basic point that the paper is addressing
+averages of logarithms.}
+
+I thought about this for a while, and decided not to change the notation much.
+The reason is that my notation is closer to what is used in introductory
+information theory texts. One usually sees things such as $H(Y|X=x)$
+and $H(Y|X)$, where the latter is the expected value of the former (averaged
+over $X$). This is the standard definition of conditional entropy.
+I think my writing makes it clear what I mean.
+However, in my idiosyncratic $H(\theta | d = d_0)$ I decided to change the
+subscript 0 to a subscript obs for observed.
+
+{\quote
+4. At the top of page 5, I find it a bit nicer to initialise nested sampling with X 0 = 1 rather than
+X 1 = t 1 . That also helps with the evaluation...}
+
+Unfortunately we have competing tastes again at this point.
+I don't like to include $X_0$ because there's no zeroth iteration of Nested Sampling.
+
+{\quote
+5. I realise that L is the active constraint during iterations, but I wonder if readers might get confused
+by the use of symbol L as radius (mnemonic length?) whereas the standard usage has been L = likelihood.}
+
+Yes, $L$ was for length, and it was unintentional that I had $L$ for both
+likelihood and length. I've changed this to $r$ for radius throughout.
+
+{\quote
+6. Equation 27 is messy in form and mistaken in typography. My memorable formula for
+the volume of a hypersphere is}
+
+Fixed, and I've replaced the volume formula with the given (easier to read and
+remember) suggestion.
+
+{\quote
+7. Near the bottom of page 7 (Algorithm 1), I was confused by
+b
+h $\leftarrow$ b
+h + + [k/N ]
+Why double +? Why the space between them? Do square brackets mean integer rounding? If so, should
+it be [k]/N ?}
+
+Sorry about the confusion. I was trying to represent the operation of
+appending one new element (a depth estimate)
+to an existing list, which in Python is {\tt +}
+and in Haskell (my current flavour of the month language) is {\tt ++}. The square brackets were meant to represent
+the wrapping of a value within a list (to make a list of length one),
+since an append function is probably a function of two lists.
+All of this was too obscure. I have replaced it with
+\begin{align}
+\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
+\end{align}
+which is hopefully more understandable.
+
+{\quote
+8. In the caption of Figure 1, “around the red point” or “at the red ball” might be better than “at
+the red point”.}
+
+Changed.
+
+{\quote
+9. At equation 29,
+1/100
+should be
+1/100.01
+to conform to the 4-figure accuracy elsewhere.}
+
+That was actually a mistake (I forgot to factor in the non-uniform prior),
+rather than overzealous rounding. Thanks for noticing it.
+
+{\quote
+10. Below equation (45), I think ``$\ln(2 \times 10 ^{-5} )$'' would better be explained as ``ln(2L) = -10.820''.}
+
+Fixed. It now says $\ln (2r)$.
+
+{\quote
+11. In the last paragraph of the text (page 19), there is a correct statement about how “to increase
+one’s confidence in the results, [without getting] a guarantee”. It might be worth pointing out that there
+can never be such a guarantee for any method that relies on pointwise evaluations, because there will
+always be some risk of a dominant spike being missed.}
+
+I've added such a sentence.
+
+{\quote
+12. In the references, my search for ref. 20 gave me “Statistics and Computing, 2017, 27, 219–236”.}
+
+Somehow my bibliography file had become corrupted. I have fixed this.
+
 \section*{Response to the other reviewer's comments}
@@
 {\quote
-2.
-With regard to the algorithm to compute the depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.
+2. With regard to the algorithm to compute tSorry about the confusion. I was trying to represent the operation of
+appending one new element (a depth estimate)
+to an existing list, which in Python is {\tt +}
+and in Haskell is {\tt ++}. The square brackets were meant to represent
+the wrapping of a value within a list (to make a list of length one),
+since an append function is probably a function of two lists.
+All of this was too obscure. I have replaced it with
+\begin{align}
+\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
+\end{align}
+which is hopefully more understandable.he depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.
 Also at that step, do the square brackets refer to an array (of one element)?

### 2017-08-14 paper/ms.tex
Finalish changes

@@
 Then $H(\x) = \left< \depth \right>_{p(\x)} + \log V(r)$.
+{\revision It is important to remember that the definition of
+volume here is in terms of the coordinates $\x$, and that a change
+of coordinates (or a change of the definition of the distance function)
+would result in a different value of the
+differential entropy.}
+
 {\revision Reducing $r$ towards zero ought to lead to
@@
 Thus, there is a an awkward trade-off involved in choosing the value of $r$,
-which will need to be investigated more thoroughly in the future.}
+which will need to be investigated more thoroughly in the future.
+However, the algorithm as it currently stands ought to be useful for calculating
+the entropies of any low-to-moderate dimensional functions of an
+underlying high-dimensional space.}
 
@@
 approximate with standard Nested Sampling, but the estimate for a given dataset
-may be biased. This method also would not work if we want $I(\x; \d)$ for
+may be biased. {\revision It is quite plausible that the bias is negligible
+compared to the variance from averaging over many datasets. This is probably
+true in many situations}.
+
+However, more significantly, the above method
+would not work if we want $I(\x; \d)$ for
 a single parameter or a subset of them, rather than for all of the parameters.
+{\revision
+The prior-to-posterior KL divergence measured by standard NS is the
+amount learned about {\em all} parameters, including nuisance parameters}.
 Therefore, I adopted the third formula as the best way to compute $I(\x; \d)$.
-In the example, $H(\x)$ is available analytically, and I use the algorithm to
+In the sinusoidal example, $H(\x)$ is available analytically, and I use the algorithm to
 obtain $H(\x | \d)$ as follows:
@@
 ``likelihood'' was minus a distance function defined in terms of the
-parameter(s).
+parameter(s) of interest.
 

### 2017-08-14 paper/ms.tex
Final changes

@@
 particles. This is why separate NS runs can be simply merged
-\citep{skilling2006nested}.
+\citep{skilling2006nested, HENDERSON201784}.
 \citet{Walter2017point} showed how this view of the NS sequence of points
@@
 would result in a different value of the
-differential entropy.}
+differential entropy. This is equivalent to counting possibilities
+or volumes with a different choice of base measure $q(\x)$.}
 
@@
 as merely a
-latent variable that allows $p(\boldsymbol{x})$ to be explored.
+latent variable allowing $p(\boldsymbol{x})$ to be explored conveniently.
 }
@@
 Therefore, I adopted the third formula as the best way to compute $I(\x; \d)$.
-In the sinusoidal example, $H(\x)$ is available analytically, and I use the algorithm to
+In the sinusoidal example, the parameter of interest is the period, whereas
+the model will contain other parameters as well.
+Since the period will have its prior specified explicitly,
+$H(\x)$ will be available analytically. I then use the algorithm to
 obtain $H(\x | \d)$ as follows:
@@
 $H(\tau | \boldsymbol{Y})$ using the algorithm with a tolerance of $r = 10^{-5}$.
+{\revision The distance function was just the absolute difference between
+$\tau$ and $\tau_{\rm ref}$.}
+
 The result was
@@
 $S$ being the set of all statements that imply $S$ \citep{knuth_questions}.
+
+{\revision
+The current implementation of the algorithm (see Appendix~\ref{sec:software})
+is in C++ and has not been parallelised. Given the high computational
+costs, parallelisation should be attempted. This could follow
+the method of \citep{HENDERSON201784}, or alternatively, different
+threads could work on different reference particles simultaneously.}
 

### 2017-08-31 talk/abstract.tex
NZSA abstract

\documentclass[12pt]{article}

\usepackage{microtype}
% \documentstyle{iascars2017}

% \usepackage{iascars2017}

\pagestyle{myheadings} 
\pagenumbering{arabic}
\topmargin 0pt \headheight 23pt \headsep 24.66pt
%\topmargin 11pt \headheight 12pt \headsep 13.66pt
\parindent = 3mm

\begin{document}

\begin{flushleft}

{\LARGE\bf Computing Entropies with Nested Sampling}

\vspace{1.0cm}

Brendon J. Brewer$^1$

\begin{description}

\item $^1 \;$ Department of Statistics, The University of Auckland,
Private Bag 92019, Auckland 1142, New Zealand

\end{description}

\end{flushleft}

%  ***** ADD ENOUGH VERTICAL SPACE HERE TO ENSURE THAT THE *****
%  ***** ABSTRACT (OR MAIN TEXT) STARTS 5 CM BELOW THE TOP *****

\vspace{0.75cm}

\noindent {\bf Abstract}. The Nested Sampling algorithm, invented in the
mid-2000s by John Skilling, represented a major advance in Bayesian computation.
Whereas Markov Chain Monte Carlo (MCMC) methods are usually effective for sampling
posterior distributions, Nested Sampling also calculates the marginal likelihood
integral used for model comparison, which is a computationally demanding task.
However, there are other kinds of integrals that we might want to compute.
Specifically, the entropy, relative entropy, and mutual information,
which quantify uncertainty and relevance, are all integrals whose form is
inconvenient in most practical applications. I will present my technique,
based on Nested Sampling, for estimating these quantities for probability
distributions that are only accessible via MCMC sampling.
This includes posterior
distributions, marginal distributions, and distributions of derived quantities.
I will present an example from experimental design, where one wants to
optimise the relevance of the data for inference of a parameter.

\vskip 2mm

\noindent {\bf Keywords}.
Bayesian inference, Nested Sampling, Markov Chain Monte Carlo,
Information theory

%\section{ First-level heading}
%The C98 head 1 style leaves a half-line spacing below a
%first-level heading. There should be one blank line above
%a first-level heading.
%        
%\subsection { Second-level heading}
%There should also be one blank line above a second- or
%third-level heading (but no extra space below them).
%
%Do not intent the first paragraph following a heading.
%Second and subsequent paragraphs are indented by one Tab
%character (= 3 mm). If footnotes are used, they should be
%placed at the foot of the page\footnote{ Footnotes are separated
%from the text by a blank line and a printed line of length 3.5 cm.
%They should be printed in 9-point Times Roman in single line spacing.}.
%        
%\subsubsection { Third-level heading}
%Please specify references using the conventions
%illustrated below. Each should begin on a new line, and
%second and subsequent lines should be on the same page
%indented by 3 mm.

\subsection*{References}

\begin{description}

\item Brewer, B. J. (2017).
\textit{Computing Entropies with Nested Sampling.}
Entropy, 19, 422.

\item Skilling, J. (2006).
\textit{Nested Sampling for General Bayesian Computation.}
Bayesian analysis, 1(4), 833-859.

\end{description}

\end{document}

### 2017-10-15 talk/ms.tex
Working on talk

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{palatino}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{multimedia}

\usetheme{Warsaw}
\usecolortheme{crane}

% www.sharelatex.com/learn/Beamer

\title{Computing Entropies with Nested Sampling}
\author{Brendon J. Brewer}
\institute{Department of Statistics\\
The University of Auckland}
\date{{\tt \color{blue} https://www.stat.auckland.ac.nz/\~{ }brewer/}}

\begin{document}

\frame{\titlepage}

% New slide
\begin{frame}
\frametitle{What is entropy?}

Firstly, I'm talking about information theory, not thermodynamics (though the
two are connected).

\end{frame}

% New slide
\begin{frame}
\frametitle{What is entropy?}

Firstly, I'm talking about information theory, not thermodynamics (though the
two are connected).

\end{frame}

% New slide
\begin{frame}
\frametitle{References I.}

On the connection between Shannon entropy and thermodynamic entropy,
see: \vspace{2em}

{\footnotesize Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
American Journal of Physics 33, no. 5 (1965): 391-398. \\
}

\end{frame}

% New slide
\begin{frame}
\frametitle{References II.}

On the connection between Shannon entropy and thermodynamic entropy,
see: \vspace{2em}

{\footnotesize Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
American Journal of Physics 33, no. 5 (1965): 391-398. \\

Brewer, Brendon J. ``Unscrambling the Second Law of Thermodynamics''
{\color{blue} \tt
http://quillette.com/2016/01/28/unscrambling-the-second-law-of-thermodynamics/}
}

\end{frame}

\end{document}

### 2017-10-16 talk/ms.tex
A figure

@@
 \begin{frame}
-\frametitle{What is entropy?}
+\frametitle{Shannon entropy}
 
-Firstly, I'm talking about information theory, not thermodynamics (though the
-two are connected).
+Consider a discrete probability distribution with probabilities
+$\boldsymbol{p} = \{p_i\}$. The Shannon entropy is
+\begin{align}
+H(\boldsymbol{p}) &= -\sum_i p_i \log p_i
+\end{align}
+
+It is a real-valued property of the distribution.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Relative entropy}
+
+Consider two discrete probability distributions with probabilities
+$\boldsymbol{p} = \{p_i\}$ and $\boldsymbol{q} = \{q_i\}$.
+The relative entropy is
+\begin{align}
+H(\boldsymbol{p}; \boldsymbol{q}) &= -\sum_i p_i \log\left(\frac{p_i}{q_i}\right)
+\end{align}
+Without the minus sign,
+it's the `Kullback-Leibler divergence', and is more fundamental than the
+Shannon entropy. With uniform $\boldsymbol{q}$, it reduces to the Shannon
+entropy (up to an additive constant).
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Entropy quantifies uncertainty}
+If there are just $N$ equally likely possibilities,
+i.e., $p_i = 1/N$, then $H = \log N$. \vspace{0.5em}
+
+\begin{center}
+\includegraphics[width=0.6\textwidth]{entropy1.pdf}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Statements and questions}
+Consider a problem with three mutually exclusive, exhaustive possibilities
+$a$, $b$, and $c$. The lattice of statements is
 

### 2017-10-16 talk/figures.py

@@ -1,0 +1,25 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Set fonts
+plt.rc("font", size=20, family="Serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
+# Set up a discrete probability distribution
+x = np.arange(1, 4)
+p = np.ones(len(x)) / len(x)
+
+# Make the plot
+plt.figure(figsize=(8, 6.5))
+plt.bar(x, p, color="lightgray", width=0.6)
+plt.ylim()
+plt.gca().set_xticks([1.0, 2.0, 3.0])
+plt.gca().set_yticks([0.0, 0.1, 0.2, 0.3, 0.4])
+plt.ylim([0.0, 0.4])
+plt.xlabel("$x$")
+plt.ylabel("Probability")
+plt.title("$H \\approx 1.0986$ nats")
+plt.savefig("entropy1.pdf", bbox_inches="tight")
+print("Created entropy1.pdf")
+
+

### 2017-10-16 talk/ms.tex
More writing, lattice image

@@
 \usepackage{multimedia}
+\usepackage{hyperref}
 
@@
 The University of Auckland}
-\date{{\tt \color{blue} https://www.stat.auckland.ac.nz/\~{ }brewer/}}
+\date{\color{blue}\url{https://www.stat.auckland.ac.nz/~brewer/}}
 
@@
 \begin{frame}
+\frametitle{Entropy quantifies uncertainty}
+If there are just $N$ equally likely possibilities,
+i.e., $p_i = 1/N$, then $H = \log N$. \vspace{0.5em}
+
+\begin{center}
+\includegraphics[width=0.6\textwidth]{entropy2.pdf}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Entropy quantifies uncertainty}
+If there are just $N$ equally likely possibilities,
+i.e., $p_i = 1/N$, then $H = \log N$. \vspace{0.5em}
+
+\begin{center}
+\includegraphics[width=0.6\textwidth]{entropy3.pdf}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{What about densities?}
+We get `differential entropy'
+
+\begin{align}
+H = - \int_{\textnormal{all }x} f(x) \log f(x) \, dx
+\end{align}
+
+This generalises log-volume, as defined {\bf with respect to} $dx$.
+
+\begin{alertblock}{Important}
+Differential entropy is {\em coordinate-system dependent}.
+\end{alertblock}
+
+\end{frame}
+
+% New slide
+\begin{frame}
 \frametitle{Statements and questions}
 Consider a problem with three mutually exclusive, exhaustive possibilities
-$a$, $b$, and $c$. The lattice of statements is
+\texttt{a}, \texttt{b}, and \texttt{c}.
+
+\begin{center}
+  \includegraphics[width=0.3\textwidth]{lattice1.pdf}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Some entropies in Bayesian statistics}
+Writing them in terms of parameters $\boldsymbol{\theta}$ and data
+$d$, for Bayesian purposes.
+
+\begin{enumerate}
+\item Entropy of the prior for the parameters $H(\boldsymbol{\theta})$
+\item Entropy of the conditional prior for the data $H(\boldsymbol{\theta} | d)$
+\item Entropy of the posterior $H(\boldsymbol{\theta} | d)$
+\item Entropy of the prior for the data $H(d)$
+\end{enumerate}
+
+Note: {\em conditional entropies} such as (2) and (3)
+are defined using an expectation over the quantity conditioned upon.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Connections}
+
+Entropy of the joint prior:
+\begin{align}
+H(\boldsymbol{\theta}, d) &= H(\boldsymbol{\theta}) + H(d | \boldsymbol{\theta}) \\
+                          &= H(d) + H(\boldsymbol{\theta} | d).
+\end{align}
+
+Mutual information:
+\begin{align}
+I(\boldsymbol{\theta}; d) &= H(\boldsymbol{\theta}) - H(\boldsymbol{\theta} | d).
+\end{align}
+This quantities dependence --- or more fundamentally,
+relevance, or the potential for learning.
 
@@
 
-{\footnotesize Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
-American Journal of Physics 33, no. 5 (1965): 391-398. \\
-}
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{References II.}
-
-On the connection between Shannon entropy and thermodynamic entropy,
-see: \vspace{2em}
-
-{\footnotesize Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
+{\tiny Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
 American Journal of Physics 33, no. 5 (1965): 391-398. \\
@@
 Brewer, Brendon J. ``Unscrambling the Second Law of Thermodynamics''
-{\color{blue} \tt
-http://quillette.com/2016/01/28/unscrambling-the-second-law-of-thermodynamics/}
+{\color{blue}
+  \url{https://quillette.com/2016/01/28/unscrambling-the-second-law-of-thermodynamics}
 }
+} % tiny
 

### 2017-10-16 talk/figures.py
Committing before going home

@@ -18,8 +18,39 @@ plt.gca().set_yticks([0.0, 0.1, 0.2, 0.3, 0.4])
 plt.ylim([0.0, 0.4])
 plt.xlabel("$x$")
 plt.ylabel("Probability")
-plt.title("$H \\approx 1.0986$ nats")
+plt.title("$H = 1.0986$ nats")
 plt.savefig("entropy1.pdf", bbox_inches="tight")
 print("Created entropy1.pdf")
+
+# Set up a discrete probability distribution
+x = np.arange(5, 26)
+p = np.ones(len(x)) / len(x)
+
+# Make the plot
+plt.figure(figsize=(8, 6.5))
+plt.bar(x, p, color="lightgray", width=0.6)
+plt.ylim([0.0, 0.055])
+plt.xlabel("$x$")
+plt.ylabel("Probability")
+plt.title("$H = 3.0445$ nats")
+plt.savefig("entropy2.pdf", bbox_inches="tight")
+print("Created entropy2.pdf")
+
+
+# Set up a discrete probability distribution
+x = np.hstack([np.arange(5, 21), np.arange(41, 46)])
+p = np.ones(len(x)) / len(x)
+
+# Make the plot
+plt.figure(figsize=(8, 6.5))
+plt.bar(x, p, color="lightgray", width=0.6)
+plt.ylim([0.0, 0.055])
+plt.xlabel("$x$")
+plt.ylabel("Probability")
+plt.title("$H = 3.0445$ nats")
+plt.savefig("entropy3.pdf", bbox_inches="tight")
+print("Created entropy3.pdf")
+
+
 
 

### 2017-10-16 talk/ms.tex
Committing before going home
Edited one character :)
Moved some things around

@@
 \begin{frame}
-\frametitle{Statements and questions}
-Consider a problem with three mutually exclusive, exhaustive possibilities
-\texttt{a}, \texttt{b}, and \texttt{c}.
+\frametitle{Some entropies in Bayesian statistics}
+Written them in terms of parameters $\boldsymbol{\theta}$ and data
+$d$, for Bayesian purposes.
 
-\begin{center}
-  \includegraphics[width=0.3\textwidth]{lattice1.pdf}
-\end{center}
+\begin{enumerate}
+\item<2-> Entropy of the prior for the parameters $H(\boldsymbol{\theta})$
+\item<3-> Entropy of the conditional prior for the data $H(\boldsymbol{\theta} | d)$
+\item<4-> Entropy of the posterior $H(\boldsymbol{\theta} | d)$
+\item<5-> Entropy of the prior for the data $H(d)$
+\end{enumerate}
 
@@
 \frametitle{Some entropies in Bayesian statistics}
-Writing them in terms of parameters $\boldsymbol{\theta}$ and data
-$d$, for Bayesian purposes.
 
-\begin{enumerate}
-\item Entropy of the prior for the parameters $H(\boldsymbol{\theta})$
-\item Entropy of the conditional prior for the data $H(\boldsymbol{\theta} | d)$
-\item Entropy of the posterior $H(\boldsymbol{\theta} | d)$
-\item Entropy of the prior for the data $H(d)$
-\end{enumerate}
-
-Note: {\em conditional entropies} such as (2) and (3)
+\begin{block}{Remark}
+{\em Conditional entropies} such as (2) and (3)
 are defined using an expectation over the quantity conditioned upon.
+\end{block}
 
@@
 \end{align}
-This quantities dependence --- or more fundamentally,
+This quantifies dependence --- or more fundamentally,
 relevance, or the potential for learning.
+
+{\tiny There are many other ways of expressing $I$.}
 
@@
 
+% New slide
+\begin{frame}
+\frametitle{Statements and questions}
+Consider a problem with three mutually exclusive, exhaustive possibilities
+\texttt{a}, \texttt{b}, and \texttt{c}.
+
+\begin{center}
+  \includegraphics[width=0.3\textwidth]{lattice1.pdf}
+\end{center}
+
+\end{frame}
+
 \end{document}

### 2017-10-16 talk/ms.tex
Another slide
Removed bold

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{dsfont}
@@
 \frametitle{Some entropies in Bayesian statistics}
-Written them in terms of parameters $\boldsymbol{\theta}$ and data
+Written them in terms of parameters $\theta$ and data
 $d$, for Bayesian purposes.
@@
 \begin{enumerate}
-\item<2-> Entropy of the prior for the parameters $H(\boldsymbol{\theta})$
-\item<3-> Entropy of the conditional prior for the data $H(\boldsymbol{\theta} | d)$
-\item<4-> Entropy of the posterior $H(\boldsymbol{\theta} | d)$
+\item<2-> Entropy of the prior for the parameters $H(\theta)$
+\item<3-> Entropy of the conditional prior for the data $H(\theta | d)$
+\item<4-> Entropy of the posterior $H(\theta | d)$
 \item<5-> Entropy of the prior for the data $H(d)$
@@
 \begin{align}
-H(\boldsymbol{\theta}, d) &= H(\boldsymbol{\theta}) + H(d | \boldsymbol{\theta}) \\
-                          &= H(d) + H(\boldsymbol{\theta} | d).
+H(\theta, d) &= H(\theta) + H(d | \theta) \\
+                          &= H(d) + H(\theta | d).
 \end{align}
@@
 \begin{align}
-I(\boldsymbol{\theta}; d) &= H(\boldsymbol{\theta}) - H(\boldsymbol{\theta} | d).
+I(\theta; d) &= H(\theta) - H(\theta | d).
 \end{align}
@@
 
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Pre-data considerations}
+
+We might want to know {\em how relevant the data is to the parameters},
+before learning the data. We might want to optimise that quantity for
+experimental design. But it's nasty, especially if there's nuisance
+parameters.
 \end{frame}

### 2017-10-16 talk/ms.tex
Hard integrals slide
Pasted a slide from NS talk

@@
 \begin{frame}
+\frametitle{Hard integrals}
+
+E.g.
+\begin{align}
+H(\theta|d) &= -\int p(d) \int p(\theta|d) \log p(\theta|d) \, d\theta \, dd \\
+            &= -\int p(d) \int p(\theta|d)
+                    \log\left[\frac{p(\theta)p(d|\theta)}{p(d)}\right] \, d\theta \, dd
+\end{align}
+
+But $p(d)$, sitting there inside a logarithm, is already supposed to be
+a hard integral (the marginal likelihood / evidence)...
+
+\begin{align}
+p(d) &= \int p(\theta) p(d|\theta) \, d\theta
+\end{align}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Nested Sampling uses {\em constrained priors}}
+Instead of using a sequence of distributions
+defined by $\pi(\theta)L(\theta)^{1/T}$, use a sequence defined by
+
+\begin{align*}
+p(\theta; \ell) &\propto \left\{
+    \begin{array}{lr}
+        \pi(\theta), & L(\theta) > \ell\\
+        0,           & \textnormal{otherwise}.
+    \end{array}
+\right.
+\end{align*}
+and have $\ell$ increase at a controlled rate, so that enclosed prior mass
+shrinks {\em geometrically}.
+
+\end{frame}
+
+% New slide
+\begin{frame}
 \frametitle{References I.}

### 2017-10-17 talk/ms.tex
lots of writing

--- 2017-10-16 talk/ms.tex
@@
 \begin{frame}
+\frametitle{Interpretation of conditional entropies}
+
+How uncertain would the question ``what's the value of $\theta$,
+precisely?'' be if the question ``what's the value of $d$, precisely''
+were to be resolved?
+
+\end{frame}
+
+% New slide
+\begin{frame}
 \frametitle{Connections}
@@
 parameters.
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Hard integrals}
+
+E.g.
+\begin{align}
+H(\theta|d) &= -\int p(d) \int p(\theta|d) \log p(\theta|d) \, d\theta \, dd \\
+            &= -\int p(d) \int p(\theta|d)
+                    \log\left[\frac{p(\theta)p(d|\theta)}{p(d)}\right] \, d\theta \, dd
+\end{align}
+
+But $p(d)$, sitting there inside a logarithm, is already supposed to be
+a hard integral (the marginal likelihood / evidence)...
+
+\begin{align}
+p(d) &= \int p(\theta) p(d|\theta) \, d\theta
+\end{align}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Marginal Likelihood Integral}
+Nested Sampling was invented in order to do this hard integral
+\begin{align}
+p(d) &= \int p(\theta) p(d|\theta) \, d\theta
+\end{align}
+
+or
+\begin{align}
+Z &= \int \pi(\theta) L(\theta) \, d\theta
+\end{align}
+where $\pi=$ prior, $L=$ likelihood.
+It's just an expectation. Why is it hard?
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Simple Monte Carlo fails}
+
+\begin{align}
+Z &= \int \pi(\theta) L(\theta) \, d\theta \\
+  &\approx \frac{1}{N} \sum_{i=1}^N L(\theta_i)
+\end{align}
+with $\theta_i \sim \pi$ will probably miss the
+tiny regions where $L$ is high.
+
+Equivalently, $\pi$ implies a {\em very heavy-tailed}
+distribution of $L$-values, and simple Monte Carlo fails.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Nested Sampling}
+Nested Sampling takes the original problem and constructs
+a 1D problem from it.
+
+\begin{align}
+Z = \int_0^1 L(X) dX
+\end{align}
+where
+\begin{align}
+X(\ell) &= \int_{L(\theta) > \ell} \pi(\theta) \, d\theta
+\end{align}
+
+\begin{block}{The meaning of $X$}
+$X(\ell)$ is the amount of prior mass whose likelihood exceeds
+$\ell$. As $\ell$ increases, $X$ decreases.
+\end{block}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Nested Sampling}
+
+\begin{center}
+\includegraphics[width=0.8\textwidth]{skilling.png}
+
+Figure from Skilling (2006).
+\end{center}
+
+Since $X(\ell)$ is the CDF of $L$-values implied by $\pi$,
+points $\theta \sim \pi$ have a uniform distribution over $X$.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Nested Sampling}
+The idea is to generate a sequence of points with increasing
+likelihoods, such that we can estimate their $X$ values. Since
+we know their $L$-values, we can then do the integral 
+numerically.
+
+\begin{align}
+Z = \int_0^1 L(X) dX
+\end{align}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Nested Sampling algorithm}
+
+\begin{enumerate}
+  \item<2-> Generate $N$ points from $\pi$
+  \item<3-> Find the worst one (lowest likelihood $L^*$), save it
+  \item<4-> Estimate its $X$ using Beta$(N,1)$ distribution\footnote{From order statistics.}
+  \item<5-> Record worst particle and its $X$-value, then discard it.
+  \item<6-> Replace that point with a new one from $\pi$ but with
+        likelihood above $L^*$.
+  \item<7-> Repeat steps 2--5 indefinitely.
+\end{enumerate}
+
 \end{frame}

### 2017-10-17 talk/ms.tex
Some stuff

@@
 \end{enumerate}
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{The sequence of $X$-values}
+The sequence of $X$ values, if you transform them, have
+a Poisson process distribution with rate $N$.
+
+\begin{align}
+-\ln(X_1) &\sim \textnormal{Exponential}(N) \\
+-\ln(X_2) &\sim -\ln(X_1) + \textnormal{Exponential}(N) \\
+-\ln(X_3) &\sim -\ln(X_2) + \textnormal{Exponential}(N) 
+\end{align}
+
+(forgive the notational abuse)
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Poisson process view of NS}
+
+The number of NS iterations taken
+to enter a small region (defined by a likelihood threshold) 
+as an unbiased estimator of the log-probability
+of that region!\vspace{0.7em}
+
+Also, $\pi(\theta)$ can be any distribution (needn't be
+a prior) and $L(\theta)$ any scalar function. Opportunities here...
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{My algorithm}
+To compute $H(\theta) = -\int f(\theta) \log f(\theta) \, d\theta$
+when $f$ can be sampled but not evaluated:
+
+\begin{enumerate}
+  \item<2-> Generate a `reference point' $\theta_{\rm ref}$ from $f$
+  \item<3-> Do a Nested Sampling run with $f$ as ``prior'' and
+            minus the distance to $\theta_{\rm ref}$ as
+            ``likelihood''.
+  \item<4-> Measure how many NS iterations were needed to make the
+            distance to $\theta_{\rm ref}$ really small, and divide
+            by $N$. That gives an
+            unbiased estimate of the
+            log-prob near $\theta_{\rm ref}$.
+  \item<5-> Repeat steps 1--3 many times.
+  \item<6-> Average the estimated log-probs, then apply corrections
+            to convert to density.
+\end{enumerate}
 

### 2017-10-17 talk/movie.py

@@ -68,3 +68,12 @@ plt.title("Distance from the reference point")
 plt.savefig("movie/image3.png", bbox_inches="tight")
 print("Saved movie/image3.png")
 
+
+# Add Nested Sampling points
+ns_points = generate(10)
+plt.plot(ns_points["x"], ns_points["y"], "*", color="green",
+         markersize=10, alpha=0.8)
+plt.title("Nested Sampling live points")
+plt.savefig("movie/image4.png", bbox_inches="tight")
+print("Saved movie/image4.png")
+

### 2017-10-17 talk/movie.py
Movie :)

@@ -73,7 +73,31 @@ print("Saved movie/image3.png")
 ns_points = generate(10)
 plt.plot(ns_points["x"], ns_points["y"], "*", color="green",
          markersize=10, alpha=0.8)
-plt.title("Nested Sampling live points")
+plt.title("Nested Sampling")
 plt.savefig("movie/image4.png", bbox_inches="tight")
 print("Saved movie/image4.png")
+
+# NS loop
+for i in range(0, 60):
+
+    # Find worst point
+    dist = np.sqrt((ns_points["x"] - xref)**2 + (ns_points["y"] - yref)**2)
+    worst = np.nonzero(dist == np.max(dist))[0]
+    plt.plot(ns_points["x"][worst], ns_points["y"][worst], "^", color="black",
+             markersize=10, alpha=0.8)
+    plt.savefig("movie/image{k}.png".format(k=i+5), bbox_inches="tight")
+    print("Saved movie/image{k}.png".format(k=i+5))
+
+    # Replace worst point
+    threshold = dist[worst]
+    while True:
+        new = generate()
+        dist_new = np.sqrt((new["x"] - xref)**2 + (new["y"] - yref)**2)
+        if dist_new < threshold:
+            ns_points["x"][worst] = new["x"]
+            ns_points["y"][worst] = new["y"]
+
+            plt.plot(new["x"], new["y"], "*", color="green",
+                        markersize=10, alpha=0.8)
+            break
 

### 2017-10-17 talk/movie.py
Multiple reference points

@@ -40,64 +40,70 @@ y = y[::-1]
 p = density(x, y)
 p /= p.sum()
 
-
-# First image
-plt.figure(figsize=(8, 7))
-plt.imshow(p, extent=[x.min(), x.max(), y.min(), y.max()],
-           cmap="Blues")
-plt.xlabel(r"$x$")
-plt.ylabel(r"$y$")
-plt.title("A joint pdf $f(x, y)$")
-plt.savefig("movie/image1.png", bbox_inches="tight")
-print("Saved movie/image1.png")
+k = 1
 
 
-# Second image
-point = generate()
-xref, yref = point["x"], point["y"]
-plt.plot(xref, yref, "o", color="r", alpha=0.5, markersize=12)
-plt.title("A reference point drawn from $f(x,y)$")
-plt.savefig("movie/image2.png", bbox_inches="tight")
-print("Saved movie/image2.png")
+for rep in range(10):
 
+    # First image
+    plt.figure(figsize=(8, 7))
+    plt.imshow(p, extent=[x.min(), x.max(), y.min(), y.max()],
+               cmap="Blues")
+    plt.xlabel(r"$x$")
+    plt.ylabel(r"$y$")
+    plt.title("A joint pdf $f(x, y)$")
+    plt.savefig("movie/image{k}.png".format(k=k), bbox_inches="tight")
+    print("Saved movie/image{k}.png".format(k=k))
+    k += 1
 
-# Add contours of the distance function
-dist = np.sqrt((x - xref)**2 + (y - yref)**2)
-plt.contour(x, y, dist, 50, alpha=0.2, colors="red", linestyles="--")
-plt.title("Distance from the reference point")
-plt.savefig("movie/image3.png", bbox_inches="tight")
-print("Saved movie/image3.png")
+    # Second image
+    point = generate()
+    xref, yref = point["x"], point["y"]
+    plt.plot(xref, yref, "o", color="r", alpha=0.5, markersize=12)
+    plt.title("A reference point drawn from $f(x,y)$")
+    plt.savefig("movie/image{k}.png".format(k=k), bbox_inches="tight")
+    print("Saved movie/image{k}.png".format(k=k))
+    k += 1
 
+    # Add contours of the distance function
+    dist = np.sqrt((x - xref)**2 + (y - yref)**2)
+    plt.contour(x, y, dist, 50, alpha=0.2, colors="red", linestyles="--")
+    plt.title("Distance from the reference point")
[…]

### 2017-10-17 talk/movie.py
Bugfix

@@ -46,6 +46,7 @@ k = 1
 for rep in range(10):
 
     # First image
+    plt.clf()
     plt.figure(figsize=(8, 7))
     plt.imshow(p, extent=[x.min(), x.max(), y.min(), y.max()],
                cmap="Blues")
@@ -90,8 +91,8 @@ for rep in range(10):
         worst = np.nonzero(dist == np.max(dist))[0]
         plt.plot(ns_points["x"][worst], ns_points["y"][worst], "^", color="black",
                  markersize=10, alpha=0.8)
-        plt.savefig("movie/image{k}.png".format(k=i+5), bbox_inches="tight")
-        print("Saved movie/image{k}.png".format(k=i+5))
+        plt.savefig("movie/image{k}.png".format(k=k), bbox_inches="tight")
+        print("Saved movie/image{k}.png".format(k=k))
 
         # Replace worst point
         threshold = dist[worst]

### 2017-10-17 talk/ms.tex
Neil Lawrence, also references
Include figures from movie
Hard integrals
Finalising

@@
 two are connected).
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Information Theory}
+
+The fundamental theorem of information theory
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Information Theory}
+
+The fundamental theorem of information theory
+\begin{block}{Theorem}
+If you take the log of a probability, it seems like you understand profound
+truths.
+\end{block}
 
@@
 \end{center}
+
+Heuristic: standard deviation quantifies uncertainty `horizontally',
+entropy does it `vertically'.
 
@@
 \end{align}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Hard integrals}
+Hard integrals with nuisance parameters
+$\boldsymbol{\eta}$, interesting parameter(s)
+$\boldsymbol{\phi}$
+\vspace{0.5em}
+
+\begin{center}
+\includegraphics[width=0.9\textwidth]{hard-integrals.png}
+\end{center}
 
@@
 \end{enumerate}
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{My algorithm}
+
+\begin{center}
+\includegraphics[width=0.7\textwidth]{movie/image000001.png}
+\end{center}
 
@@
 \begin{frame}
-\frametitle{References I.}
+\frametitle{My algorithm}
 
-On the connection between Shannon entropy and thermodynamic entropy,
-see: \vspace{2em}
-
-{\tiny Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
-American Journal of Physics 33, no. 5 (1965): 391-398. \\
-
-Brewer, Brendon J. ``Unscrambling the Second Law of Thermodynamics''
-{\color{blue}
-  \url{https://quillette.com/2016/01/28/unscrambling-the-second-law-of-thermodynamics}
-}
-} % tiny
+\begin{center}
+\includegraphics[width=0.7\textwidth]{movie/image000021.png}
+\end{center}
 
@@
 \begin{frame}
-\frametitle{Statements and questions}
-Consider a problem with three mutually exclusive, exhaustive possibilities
-\texttt{a}, \texttt{b}, and \texttt{c}.
+\frametitle{My algorithm}
 
 \begin{center}
-  \includegraphics[width=0.3\textwidth]{lattice1.pdf}
+\includegraphics[width=0.7\textwidth]{movie/image000041.png}
 \end{center}
@@
 
+% New slide
+\begin{frame}
+\frametitle{Movie}
+
+Play movie.mkv
+
+(It's also on YouTube)
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Toy experimental design example}
+Two observing strategies --- even and uneven.
+Which is better for measuring a period?
+
+\begin{center}
+\includegraphics[width=0.8\textwidth]{../paper/figures/sinewave.pdf}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Specifics}
+Let $\tau = \log_{10}(\textnormal{period})$.\vspace{0.5em}
+
+I knew $H(\tau)$ because I chose
+the prior. I used the algorithm to estimate $H(\tau | d)$,
+so marginal posteriors
+were the distributions whose entropies I estimated\footnote{If you only
+care about one parameter, define your distance function in terms of that
+parameter only!}.\vspace{0.5em}
+
+I then computed the mutual information
+\vspace{0.5em}
+
+\begin{align}
+I(\tau; d) &= H(\tau) - H(\tau | d)
+\end{align}
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Result}
+
+\begin{align}
+I_{\rm even}   &= 5.441 \pm 0.038 \textnormal{ nats}\\
+I_{\rm uneven} &= 5.398 \pm 0.038 \textnormal{ nats}
+\end{align}
+
+i.e., any difference is trivial.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{In practice...}
+
+When a period is short relative to observations, you can get a multimodal
+posterior pdf\footnote{e.g., see Larry Bretthorst's work connecting
+the posterior pdf to the periodogram.},
+and `learn a lot' by ruling out most of the space, but still
+having many peaks.\vspace{0.5em}
+
+I did not investigate this aspect of the problem --- I assumed long periods.
+
+\end{frame}
+
+% New slide
+\begin{frame}
+\frametitle{Paper/software/endorsement}
+
+{\color{blue}\url{http://www.mdpi.com/1099-4300/19/8/422}
+\url{https://github.com/eggplantbren/InfoNest}}
+\vspace{0.5cm}
+
+\begin{center}
+\includegraphics[width=0.7\textwidth]{lawrence.png}
+\end{center}
+
+\end{frame}
+
+%% New slide
+%\begin{frame}
+%\frametitle{References I.}
+
+%On the connection between Shannon entropy and thermodynamic entropy,
+%see: \vspace{2em}
+
+%{\tiny Jaynes, Edwin T. ``Gibbs vs Boltzmann entropies.''
+%American Journal of Physics 33, no. 5 (1965): 391-398. \\
+
+%Brewer, Brendon J. ``Unscrambling the Second Law of Thermodynamics''
+%{\color{blue}
+%  \url{https://quillette.com/2016/01/28/unscrambling-the-second-law-of-thermodynamics}
+%}
+%} % tiny
+
+%\end{frame}
+
+%% New slide
+%\begin{frame}
+%\frametitle{Statements and questions}
+%Consider a problem with three mutually exclusive, exhaustive possibilities
+%\texttt{a}, \texttt{b}, and \texttt{c}.
+
+%\begin{center}
+%  \includegraphics[width=0.3\textwidth]{lattice1.pdf}
+%\end{center}
+
+%\end{frame}
+
 \end{document}

### 2017-10-17 talk/ms.tex
Minor fixes

@@
 \item<2-> Entropy of the prior for the parameters $H(\theta)$
-\item<3-> Entropy of the conditional prior for the data $H(\theta | d)$
+\item<3-> Entropy of the conditional prior for the data $H(d | \theta)$
 \item<4-> Entropy of the posterior $H(\theta | d)$
@@
 {\em Conditional entropies} such as (2) and (3)
-are defined using an expectation over the quantity conditioned upon.
+are defined using an expectation over the second argument
+(the thing conditioned on).
 \end{block}

### 2017-10-17 talk/ms.tex
Thanks

@@
 \frametitle{Some entropies in Bayesian statistics}
-Written them in terms of parameters $\theta$ and data
+Written in terms of parameters $\theta$ and data
 $d$, for Bayesian purposes.
@@
 before learning the data. We might want to optimise that quantity for
-experimental design. But it's nasty, especially if there's nuisance
+experimental design. But it's nasty, especially if there are nuisance
 parameters.
@@
 to enter a small region (defined by a likelihood threshold) 
-as an unbiased estimator of the log-probability
+is an unbiased estimator of the log-probability
 of that region!\vspace{0.7em}
@@
 
+% New slide
+\begin{frame}
+\frametitle{Thanks}
+
+Ruth Angus (Flatiron Institute),
+Ewan Cameron (Oxford), James Curran (Auckland), Tom Elliott (Auckland),
+David Hogg (NYU), Kevin Knuth (SUNY Albany),
+Thomas Lumley (Auckland),
+Iain Murray (Edinburgh), John Skilling
+(Maximum Entropy Data Consultants), Jared Tobin ({\tt jtobin.io}).
+
+\end{frame}
+
 %% New slide

### 2017-10-18 talk/movie.py
Remember ffmpeg command

@@ -127,4 +127,5 @@ for rep in range(10):
                 k += 1
                 break
 
-
+# The ffmpeg command
+# ffmpeg -framerate 12 -i image%06d.png -c:v libx264 -r 12 movie.mkv

### 2017-12-08 talk/ms-abridged.tex
Attempt at shorter talk

--- 2017-10-17 talk/ms.tex
@@
 
-\usetheme{Warsaw}
+\usetheme{Copenhagen}
 \usecolortheme{crane}
@@
 The University of Auckland}
-\date{\color{blue}\url{https://www.stat.auckland.ac.nz/~brewer/}}
+\date{\color{blue}\bf \url{https://www.stat.auckland.ac.nz/~brewer/}}
 
@@
 \begin{frame}
-\frametitle{What is entropy?}
-
-Firstly, I'm talking about information theory, not thermodynamics (though the
-two are connected).
+\frametitle{The fundamental theorem of information theory}
 
@@
 \begin{frame}
-\frametitle{Information Theory}
+\frametitle{The fundamental theorem of information theory}
 
-The fundamental theorem of information theory
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Information Theory}
-
-The fundamental theorem of information theory
 \begin{block}{Theorem}
@@
 It is a real-valued property of the distribution.
+There is also relative entropy / KL divergence,
+which is a function of two distributions, and is more fundamental.
 
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Relative entropy}
-
-Consider two discrete probability distributions with probabilities
-$\boldsymbol{p} = \{p_i\}$ and $\boldsymbol{q} = \{q_i\}$.
-The relative entropy is
-\begin{align}
-H(\boldsymbol{p}; \boldsymbol{q}) &= -\sum_i p_i \log\left(\frac{p_i}{q_i}\right)
-\end{align}
-Without the minus sign,
-it's the `Kullback-Leibler divergence', and is more fundamental than the
-Shannon entropy. With uniform $\boldsymbol{q}$, it reduces to the Shannon
-entropy (up to an additive constant).
 \end{frame}
@@
 
-Heuristic: standard deviation quantifies uncertainty `horizontally',
-entropy does it `vertically'.
-
 \end{frame}
@@
 This generalises log-volume, as defined {\bf with respect to} $dx$.
-
+\pause
 \begin{alertblock}{Important}
@@
 \begin{frame}
+\frametitle{My algorithm}
+
+\begin{center}
+\includegraphics[width=0.7\textwidth]{movie/image000061.png}
+\end{center}
+
+\end{frame}
+
+% New slide
+\begin{frame}
 \frametitle{Movie}

### 2017-12-08 talk/ms-abridged.tex
Removed some more stuff

@@
 \begin{frame}
-\frametitle{Interpretation of conditional entropies}
-
-How uncertain would the question ``what's the value of $\theta$,
-precisely?'' be if the question ``what's the value of $d$, precisely''
-were to be resolved?
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Connections}
-
-Entropy of the joint prior:
-\begin{align}
-H(\theta, d) &= H(\theta) + H(d | \theta) \\
-                          &= H(d) + H(\theta | d).
-\end{align}
-
-Mutual information:
-\begin{align}
-I(\theta; d) &= H(\theta) - H(\theta | d).
-\end{align}
-This quantifies dependence --- or more fundamentally,
-relevance, or the potential for learning.
-
-{\tiny There are many other ways of expressing $I$.}
-
-\end{frame}
-
-% New slide
-\begin{frame}
 \frametitle{Pre-data considerations}
@@
 
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Hard integrals}
-Hard integrals with nuisance parameters
-$\boldsymbol{\eta}$, interesting parameter(s)
-$\boldsymbol{\phi}$
-\vspace{0.5em}
-
-\begin{center}
-\includegraphics[width=0.9\textwidth]{hard-integrals.png}
-\end{center}
+It's even worse with nuisance parameters.
 
@@
 where $\pi=$ prior, $L=$ likelihood.
-It's just an expectation. Why is it hard?
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Simple Monte Carlo fails}
-
-\begin{align}
-Z &= \int \pi(\theta) L(\theta) \, d\theta \\
-  &\approx \frac{1}{N} \sum_{i=1}^N L(\theta_i)
-\end{align}
-with $\theta_i \sim \pi$ will probably miss the
-tiny regions where $L$ is high.
-
-Equivalently, $\pi$ implies a {\em very heavy-tailed}
-distribution of $L$-values, and simple Monte Carlo fails.
+This is hard because $\pi \implies$ a heavy-tailed
+distribution of $L$-values.
 
@@
 
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Nested Sampling}
-The idea is to generate a sequence of points with increasing
-likelihoods, such that we can estimate their $X$ values. Since
-we know their $L$-values, we can then do the integral 
-numerically.
-
-\begin{align}
-Z = \int_0^1 L(X) dX
-\end{align}
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Nested Sampling algorithm}
-
-\begin{enumerate}
-  \item<2-> Generate $N$ points from $\pi$
-  \item<3-> Find the worst one (lowest likelihood $L^*$), save it
-  \item<4-> Estimate its $X$ using Beta$(N,1)$ distribution\footnote{From order statistics.}
-  \item<5-> Record worst particle and its $X$-value, then discard it.
-  \item<6-> Replace that point with a new one from $\pi$ but with
-        likelihood above $L^*$.
-  \item<7-> Repeat steps 2--5 indefinitely.
-\end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{In practice...}
-
-When a period is short relative to observations, you can get a multimodal
-posterior pdf\footnote{e.g., see Larry Bretthorst's work connecting
-the posterior pdf to the periodogram.},
-and `learn a lot' by ruling out most of the space, but still
-having many peaks.\vspace{0.5em}
-
-I did not investigate this aspect of the problem --- I assumed long periods.
-
-\end{frame}
-
-% New slide
-\begin{frame}
 \frametitle{Paper/software/endorsement}
 
-{\color{blue}\url{http://www.mdpi.com/1099-4300/19/8/422}
+{\color{blue}\bf \url{http://www.mdpi.com/1099-4300/19/8/422}
 \url{https://github.com/eggplantbren/InfoNest}}

### 2017-12-10 talk/ms-abridged.tex
Two slides into one
Fewer slides

@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{entropy1.pdf}
-\end{center}
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Entropy quantifies uncertainty}
-If there are just $N$ equally likely possibilities,
-i.e., $p_i = 1/N$, then $H = \log N$. \vspace{0.5em}
-
-\begin{center}
-\includegraphics[width=0.6\textwidth]{entropy2.pdf}
-\end{center}
-
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Entropy quantifies uncertainty}
-If there are just $N$ equally likely possibilities,
-i.e., $p_i = 1/N$, then $H = \log N$. \vspace{0.5em}
-
-\begin{center}
 \includegraphics[width=0.6\textwidth]{entropy3.pdf}
@@
 Written in terms of parameters $\theta$ and data
-$d$, for Bayesian purposes.
+$d$, for Bayesian purposes. These are relevant to
+experimental design...
 
@@
 
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Some entropies in Bayesian statistics}
+\pause
+\pause
+\pause
+\pause
+\pause
 
@@
 
-\end{frame}
-
-% New slide
-\begin{frame}
-\frametitle{Pre-data considerations}
-
-We might want to know {\em how relevant the data is to the parameters},
-before learning the data. We might want to optimise that quantity for
-experimental design. But it's nasty, especially if there are nuisance
-parameters.
 \end{frame}
@@
 
-% New slide
-\begin{frame}
-\frametitle{Poisson process view of NS}
+%% New slide
+%\begin{frame}
+%\frametitle{Poisson process view of NS}
 
-The number of NS iterations taken
-to enter a small region (defined by a likelihood threshold) 
-is an unbiased estimator of the log-probability
-of that region!\vspace{0.7em}
+%The number of NS iterations taken
+%to enter a small region (defined by a likelihood threshold) 
+%is an unbiased estimator of the log-probability
+%of that region!\vspace{0.7em}
 
-Also, $\pi(\theta)$ can be any distribution (needn't be
-a prior) and $L(\theta)$ any scalar function. Opportunities here...
+%Also, $\pi(\theta)$ can be any distribution (needn't be
+%a prior) and $L(\theta)$ any scalar function. Opportunities here...
 
-\end{frame}
+%\end{frame}
 

### 2019-01-16 talk/ms.tex
Extra image before movie
Minor updates

@@
 
-{\color{blue}\url{http://www.mdpi.com/1099-4300/19/8/422}
-\url{https://github.com/eggplantbren/InfoNest}}
+{\color{blue}
+\url{http://www.mdpi.com/1099-4300/19/8/422} \\
+
+\url{https://github.com/eggplantbren/InfoNest} \\
+
+\url{https://github.com/eggplantbren/PosteriorEntropy}
+}
 \vspace{0.5cm}
@@
 Ewan Cameron (Oxford), James Curran (Auckland), Tom Elliott (Auckland),
+Dan Foreman-Mackey (Flatiron Institute),
 David Hogg (NYU), Kevin Knuth (SUNY Albany),
